### 2021-02-11 paper.Rmd
basic template placed

---
author:
  - name: Stuart Lee
    affiliation: Monash University
    # use this syntax to add text on several lines
    address: |
      | Department of Econometrics and Business Statistics,
      | Monash University
    email: stuart.a.lee@monash.edu
  - name: Ursula Laa 
    affiliation: Monash University
    address: |
      | Department of Econometrics and Business Statistics,
      | Monash University
  - name: Dianne Cook
    address: |
      | Department of Econometrics and Business Statistics,
      | Monash University
    affiliation: Monash University
title:
  formatted: "An Overview of Tours for Dynamic Visualisation of High-dimensional Data"
  type: "Advanced Review"
abstract: > 
  Abstract goes here.
keywords:
  formatted: ["tours", "data visualisation"]
output: 
  bookdown::pdf_document2:
    template: template/wires_template_rmd.tex
    keep_tex: true
    toc: false
    citation_package: natbib
bibliography: references.bib
biblio-style: apalike
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.show = 'hold',
  dpi = 300 #,
  # out.width = "\\textwidth",
  # fig.path = here::here('img/')
)
```

# Introduction

- Purpose/why and what is high-d data
- Original tours
- Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
- Alternatives (Stuart)

# Tours for high dimensional visualization

- Bases notation
- Display
- Choosing targets, inc pp indexes
- Geodesic interpolation
- Slices vs projections
- Transformations on projections

# Ways of interacting

- Stop/start/reverse
- Manual (Nick)
- Linked brushing (Earo)
- Spin-and-brush

# Software

- tourr
- liminal
- python

# Application

- Model vis (from Hadley, removing the blidnfold)
- Physics (Ursula)
- Bioinformatics (Stuart)
- High-d shapes/distributions

# Discussion

# References {-}

### 2021-02-11 template/wires_template_guideline.tex
basic template placed

\documentclass[10pt]{article}

\usepackage{fullpage}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage[section]{placeins}
\usepackage{xcolor}
\usepackage{breakcites}
\usepackage{lineno}
\usepackage{hyphenat}

\renewcommand{\familydefault}{\sfdefault}

\PassOptionsToPackage{hyphens}{url}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\usepackage{etoolbox}
\makeatletter
%\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
%  \errmessage{\noexpand\@combinedblfloats could not be patched}%
%}%
\makeatother

\usepackage{natbib}

\renewenvironment{abstract}
  {{\bfseries\noindent{\abstractname}\par\nobreak}\footnotesize}
  {\bigskip}

\titlespacing{\section}{0pt}{*3}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.5}
\titlespacing{\subsubsection}{0pt}{*1.5}{0pt}

\usepackage{authblk}

\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{tabulary}
\usepackage{booktabs,array,multirow}
\usepackage{amsfonts,amsmath,amssymb}
\providecommand\citet{\cite}
\providecommand\citep{\cite}
\providecommand\citealt{\cite}
% You can conditionalize code for latexml or normal latex using this.
\newif\iflatexml\latexmlfalse
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}%

\AtBeginDocument{\DeclareGraphicsExtensions{.pdf,.PDF,.eps,.EPS,.png,.PNG,.tif,.TIF,.jpg,.JPG,.jpeg,.JPEG}}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{float}

\usepackage[margin=1.5in]{geometry}

\begin{document}

\title{Wiley Interdisciplinary Reviews (WIREs)}

\author[1]{Alberto Pepe}%
\affil[1]{Authorea Team}%

\vspace{-1em}

  \date{}

\begingroup
\let\center\flushleft
\let\endcenter\endflushleft
\maketitle
\endgroup

\selectlanguage{english}
\begin{abstract}
The abstract should be a concise (less than 250 words) description of
the article and its implications. It should include all keywords
associated with your article, as keywords increase its discoverability.
Please do not include generic phrases such as ``This article discusses
\ldots{}'' or ``Here we review,'' or references to other articles. Note:
You will be required to copy this abstract into the submission system
when uploading your article.%
\end{abstract}%

\sloppy

\textbf{Remember that you are writing for an interdisciplinary audience.
Please be sure to discuss interdisciplinary themes, issues, debates,
etc. where appropriate.} Note that the WIREs are forums for review
articles, rather than primary literature describing the results of
original research.

\section*{Article Title}

{\label{350486}}

The title should not exceed 20 words. Please be original and try to
include keywords, especially before a colon if applicable, as they will
increase the discoverability of your
article.~\href{https://authorservices.wiley.com/author-resources/Journal-Authors/Prepare/writing-for-seo.html}{Tips
on Search Engine Optimization}

\section*{Article Type}

{\label{925764}}

The \href{http://wires.wiley.com/go/forauthors\#ArticleTypes}{Article
Type} denotes the intended level of readership for your article. Please
select one of the below article type options. An Editor may have
mentioned a specific Article Type in your invitation letter; if so,
please let them know if you think a different Article Type better suits
your topic.~

\begin{itemize}
\tightlist
\item
  Opinion
\item
  Primer
\item
  Overview
\item
  Advanced Review
\item
  Focus Article
\item
  Software Focus
\end{itemize}

\section*{Authors}

{\label{290010}}

List each person's full name, ORCID iD, affiliation, email address, and
any conflicts of interest. Please use an asterisk (*) to indicate the
corresponding author.

The preferred (but optional) format for author names is First Name,
Middle Initial, Last Name.~~

The submitting author is required to provide
an~\href{https://authorservices.wiley.com/author-resources/Journal-Authors/Submission/orcid.html}{ORCID
iD}, and all other authors are encouraged to do so.~

Wiley requires that all authors disclose any potential conflicts of
interest. Any interest or relationship, financial or otherwise, that
might be perceived as influencing an author's objectivity is considered
a potential conflict of interest. The existence of a conflict of
interest does not preclude publication.

\section*{Abstract}

{\label{468816}}

The abstract should be a concise (less than 250 words) description of
the article and its implications. It should include all keywords
associated with your article, as keywords increase its discoverability
(\href{https://authorservices.wiley.com/author-resources/Journal-Authors/Prepare/writing-for-seo.html}{Tips
on Search Engine Optimization}). Please try not to include generic
phrases such as ``This article discusses \ldots{}'' or ``Here we
review,'' or references to other articles. Note: You will be required to
copy this abstract into the submission system when uploading your
article.

Optional: If you would like to submit your abstract in an additional
language,~\href{http://wires.wiley.com/go/forauthors\#Resources}{read
more}.

\section*{Graphical/Visual Abstract and
Caption}

{\label{750712}}

Include an attractive full color image to go under the text abstract and
in the online Table of Contents.~\textbf{You will also need to upload
this as a separate file during submission.~}It may be a figure or panel
from the article or may be specifically designed as a visual summary.
While original images are preferred, if you need to look for a
thematically appropriate stock image, you can go
to~\href{http://pixabay.com/}{pixabay.com}~(not affiliated with Wiley)
to find a free stock image with a CC0 license. Another option you have
is to utilize professional illustrators with
Wiley's~\href{https://wileyeditingservices.com/en/article-preparation/graphical-abstract-design}{Graphical
Abstract Design service}.

Size: The minimum resolution is 300 dpi. Please keep the image as simple
as possible because it will be displayed in multiple sizes. Multiple
panels and text other than labels are strongly discouraged.

Caption: This is a narrative sentence to convey the article's essence
and wider implications to a non-specialist audience. The maximum length
is 50 words, but consider using 280 characters or less to facilitate
social media sharing, which can increase the discoverability of your
article.

\par\null

\section*{1. Introduction}

{\label{252565}}

Introduce your topic in \textasciitilde{}2 paragraphs,
\textasciitilde{}750 words.

While Wiley does consider articles on preprint servers (ArXiv, bioRxiv,
psyArXiv, SocArXiv, engrXiv, etc.) for submission to primary research
journals, preprint articles should not be cited in WIREs manuscripts as
review articles should discuss and draw conclusions only from
peer-reviewed research. Remember that original research/unpublished work
should also not be included as it has not yet been peer-reviewed and
could put the work in jeopardy of getting published in the primary
press.

Citations are automatically generated by Authorea. Select~\textbf{cite}
to find and cite bibliographic resources. The citations will
automatically be generated for you in APA format, the style used by most
WIREs titles. If you are writing for~\emph{WIREs Computational Molecular
Science} (WCMS), you will need to use the Vancouver reference and
citation style, so before exporting click Export-\textgreater{} Options
and select a Vancouver export style.

A sample citation:~~\hyperref[csl:1]{(Murphy et al., 2019)}

\section*{2. First Level Heading}

{\label{350277}}

Begin the main body of the text here, using a maximum of 3 levels of
headings (style as shown below and numbered). Try to create headings
that:

\begin{itemize}
\tightlist
\item
  help the reader find information quickly;~
\item
  are descriptive yet specific;
\item
  are compatible in phrasing and style; and~
\item
  are concise (less than 50 characters)
\end{itemize}

\subsection*{2.1 Second level heading}

{\label{733281}}

Text

\subsubsection*{2.1.1 Third level heading}

{\label{824595}}

Text

\par\null

\section*{Figures and Tables}

{\label{432859}}

Figures and tables should be numbered separately, in the order in which
they appear in the manuscript. Please embed them in the correct places
in the text to facilitate peer review. Permission to \textbf{reuse or
adapt} previously published materials must be submitted before article
acceptance.
\href{http://wires.wiley.com/go/forauthors\#Resources}{Resources}

Production quality figure files with captions are still to be submitted
separately.
\href{https://authorservices.wiley.com/asset/photos/electronic_artwork_guidelines.pdf}{Figure
preparation and formatting}

\textbf{Captions should stand alone} and be informative outside of the
context of the article. This will help educators who may want to use a
PowerPoint slide of your figure. Explain any abbreviations or symbols
that appear in the figure and make sure to include \textbf{credit lines}
for any previously published materials.

\section*{Conclusion}

{\label{880788}}

Sum up the key conclusions of your review, highlighting the most
promising scientific developments, directions for future research,
applications, etc. The conclusion should be \textasciitilde{}2
paragraphs, \textasciitilde{}750 words total.

\section*{Funding Information}

{\label{974317}}

You will be required to enter your funding information into the
submission system so that we can apply proper IDs to your funders and
help you comply with any funder mandates.

\section*{Research Resources}

{\label{808103}}

List sources of non-monetary support such as supercomputing time at a
recognized facility, special collections or specimens, or access to
equipment or services.
\href{https://orcid.org/organizations/research-orgs/resources}{Research
Resources} can also be added to ORCiD profiles. For biomedical
researchers, the \href{https://scicrunch.org/resources}{Resource
Identification Portal} supports NIH's new guidelines for Rigor and
Transparency in biomedical publications.

\section*{Acknowledgments}

{\label{749861}}

List contributions from individuals who do not meet the criteria for
authorship (for example, to recognize people who provided technical
help, collation of data, writing assistance, acquisition of funding, or
a department chairperson who provided general support), \textbf{with
permission} from the individual. Thanks to anonymous reviewers are not
appropriate.

\section*{Notes}

{\label{390481}}

Authors writing from a humanities or social sciences perspective may use
notes if a \textbf{comment or additional information} is needed to
expand on a citation. (Notes only containing citations should be
converted to references. Conversely, any references containing comments,
such as ``For an excellent summary of\ldots{},'' should be converted to
notes.) Notes should be indicated by \textbf{superscript letters}, both
in the text and in the notes list. Citations within notes should be
included in the reference section, as indicated below.

\section*{Further Reading}

{\label{153582}}

For readers who may want more information on concepts in your article,
provide full references and/or links to additional recommended resources
(books, articles, websites, videos, datasets, etc.) that are not
included in the reference section. Please do not include links to
non-academic sites, such as Wikipedia, or to impermanent websites.

\section*{\texorpdfstring{{Note About
References}}{Note About References}}

{\label{514168}}

References are automatically generated by Authorea.
Select~\textbf{cite~}to find and cite bibliographic resources. The
bibliography will automatically be generated for you in APA format, the
style used by most WIREs titles. If you are writing for~\emph{WIREs
Computational Molecular Science}~ (WCMS), you will need to use
the~Vancouver reference style, so before exporting click
Export-\textgreater{} Options and select a Vancouver export style.~

\selectlanguage{english}
\FloatBarrier
\section*{References}\sloppy
\phantomsection
\label{csl:1}{Understanding institutions for water allocation and exchange: Insights from dynamic agent-based modeling}. (2019). \textit{Wiley Interdisciplinary Reviews: Water}, \textit{6}(6). \url{https://doi.org/10.1002/wat2.1384}

\end{document}

### 2021-02-16 paper.Rmd
assigning tasks

@@
   - name: Ursula Laa 
-    affiliation: Monash University
+    affiliation: 
     address: |
-      | Department of Econometrics and Business Statistics,
-      | Monash University
+      | Department of  ,
+      |  University
   - name: Dianne Cook
@@
     affiliation: Monash University
+    email: dicook@monash.edu
 title:
@@
 
+<!--
+Writing style: third person (it was done, ...)
+-->
+
 # Introduction
 
-- Purpose/why and what is high-d data
-- Original tours
+- Purpose/why and what is high-d data (Di)
+- Original tours (Di)
 - Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
@@
 
-- Bases notation
-- Display
-- Choosing targets, inc pp indexes
-- Geodesic interpolation
-- Slices vs projections
-- Transformations on projections
+- Bases notation (Sherry)
+- Display (Sherry)
+- Choosing targets, inc pp indexes (Sherry)
+- Geodesic interpolation (Ursula)
+- Slices vs projections (Ursula)
+- Transformations on projections (Ursula)
 
@@
 - Linked brushing (Earo)
-- Spin-and-brush
+- Spin-and-brush  (Ursula)
 
@@
 
-- tourr
-- liminal
-- python
+- tourr (Di) and spinifex (Nick)
+- liminal (Stuart)
+- python (Di to follow up with Csaba)
 
@@
 
-- Model vis (from Hadley, removing the blidnfold)
+- Model vis examples eg random forest (from Hadley, removing the blindfold) (Natalia)
 - Physics (Ursula)
 - Bioinformatics (Stuart)
-- High-d shapes/distributions
+- High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)
 

### 2021-02-22 paper.Rmd
fleshing out intro

@@
   formatted: "An Overview of Tours for Dynamic Visualisation of High-dimensional Data"
-  type: "Advanced Review"
 abstract: > 
@@
 
-- Purpose/why and what is high-d data (Di)
+<!-- Purpose/why and what is high-d data (Di)-->
+
+Data commonly arrives with more than two measured variables, which makes it more complicated to plot on a page.
+With multiple variables, especially if there is some association between variables, this would be called multivariate or high-dimensional data.
+When the variables are all numeric, or quantitative, visualisation often relies on some form of dimension reduction.
+This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936], or nonlinear if one considers techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbour Embedding (t-SNE) [@Van_der_Maaten2008-qa].
+For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+Interactive and dynamic displays can provide information beyond what can be achieved in a static display.
+The grand tour shows a smooth sequence of interpolated low dimensional projections, and allows the viewer to extrapolate from the low-dimensional shapes to the multidimensional distribution.
+It is particularly useful for detecting clusters, outliers and non-linear dependence.
+
 - Original tours (Di)

### 2021-02-22 paper.Rmd
more intro

@@
 When the variables are all numeric, or quantitative, visualisation often relies on some form of dimension reduction.
-This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936], or nonlinear if one considers techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbour Embedding (t-SNE) [@Van_der_Maaten2008-qa].
-For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
-Interactive and dynamic displays can provide information beyond what can be achieved in a static display.
-The grand tour shows a smooth sequence of interpolated low dimensional projections, and allows the viewer to extrapolate from the low-dimensional shapes to the multidimensional distribution.
-It is particularly useful for detecting clusters, outliers and non-linear dependence.
+This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936].
+It is also common to reduce dimension nonlinearly with techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbour Embedding (t-SNE) [@Van_der_Maaten2008-qa].
+
+The area of visualising high-dimensions is very large, so the focus of this review will be on linear projections, in particular as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+An overview of the technique, and recent developments is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
+
+A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
+It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (see for e.g.
+@Swayne2003-cc or @Tierney1991).
+To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
+It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
 

### 2021-02-22 paper.Rmd
more intro
author list info update

@@
       |  University
+  - name: Natalia da Silva
+    address: |
+      | Instituto de Estadística (IESTA), 
+      | Universidad de la República
+    affiliation: Universidad de la República
+    email: natalia@iesta.edu.uy
   - name: Dianne Cook
@@
 
-- Original tours (Di)
+<!-- Original tours (Di)
 - Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
+-->
 - Alternatives (Stuart)

### 2021-02-23 paper.Rmd
Nick added to author list.

@@
     email: natalia@iesta.edu.uy
+  - name: Nicholas Spyrison
+    address: |
+      | Faculty of Information and Technology 
+      | Monash University
+    affiliation: Monash University
+    email: nicholas.spyrison@monash.edu
   - name: Dianne Cook

### 2021-02-23 paper.Rmd
remove natbib for the moment

@@
     toc: false
-    citation_package: natbib
 bibliography: references.bib

### 2021-02-23 paper.Rmd
Affiliation added for Ursula

@@
   - name: Ursula Laa 
-    affiliation: 
+    affiliation: University of Natural Resources and Life Sciences
     address:  |
-      | Department of  ,
-      |  University
+      | Institute of Statistics,
+      | University of Natural Resources and Life Sciences
+    email:  ursula.laa@boku.ac.at
   - name: Natalia da Silva

### 2021-02-24 paper.Rmd
small tweaks to intro

@@
 
-The area of visualising high-dimensions is very large, so the focus of this review will be on linear projections, in particular as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
-An overview of the technique, and recent developments is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
+The focus of this review will be on linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+The reason being that it is not feasible to adequately the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.
+An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
 

### 2021-02-24 paper.Rmd
notes to self

@@
 - Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
+(picture of the cluster data from Stuart's thesis)
 -->
 - Alternatives (Stuart)
+
+(parallel coordinates, sploms, what tours are not)
 

### 2021-02-28 paper.Rmd
added wang

@@
     email: dicook@monash.edu
+  - name: Earo Wang
+    address: |
+      | Department of Statistics,
+      | The University of Auckland
+    affiliation: The University of Auckland
+    email: earo.wang@auckland.ac.nz
 title:

### 2021-02-28 data/clusters.R
example for opening paragraph

@@ -1,0 +1,52 @@
+# adapted from Holmes and Ngyuen, 2019
+# Simple Gaussian 10-d
+library(ggplot2)
+library(tourr)
+library(Rtsne)
+library(dplyr)
+library(readr)
+
+# 5 gaussian clusters with equal covariance
+set.seed(63968)
+m <- 500
+d <- 5
+p <- 10
+k <- 5
+sigma <- 1
+muMax <- 25
+mu <- matrix(runif(k*d, max = muMax), ncol = d)
+X <- lapply(seq_len(k), function(i) {
+  ix <- rep(1, m) %*% t(mu[i, ]) + matrix(sigma*rnorm(m*d), ncol = d)
+  as.data.frame(ix)
+})
+X_tbl <- dplyr::bind_rows(X, .id = "cluster")
+X <- as.matrix(X_tbl[,-1])
+Y <- X  %*% matrix(rnorm(d*p), ncol = p)
+colnames(Y) <- sprintf("dim%02d", 1:p)
+tour_tbl <- dplyr::bind_cols(as.data.frame(Y), 
+                             cluster = factor(X_tbl[["cluster"]]))
+
+
+tsne_y <- Rtsne::Rtsne(Y)
+embed_tbl <- tidy_tsne(tsne_y, 
+                       list(cluster = factor(X_tbl[["cluster"]])))
+colnames(embed_tbl) <- c("tsneX", "tsneY", "cluster")
+
+
+# multichallenge data set
+url <- "http://ifs.tuwien.ac.at/dm/download/multiChallenge-matrix.txt"
+multi <- read_tsv(url, col_names = FALSE) %>% 
+  mutate(
+    index = rep(seq_len(200), 5),
+    key = rep(1:5, each = 200L)
+  ) %>%
+  filter(key %in% c(1, 4)) %>% 
+  mutate(group = rep(c("hierachical", "flat"), each = 200L)) %>% 
+  select(group, X1:X10)
+
+
+set.seed(143280)
+tsne_multi <- Rtsne::Rtsne(multi[, -1])
+embed_multi <- tidy_tsne(tsne_multi, list(group = multi[["group"]]))
+colnames(embed_multi) <- c("tsneX", "tsneY", "group")
+

### 2021-03-01 paper.Rmd
add alternatives

@@
 It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
+While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualising structure in high dimensional data.
+An early technique proposed for assessing pairwise relationships between variables is the scatter plot matrix (SPLOM) [@Tukey1983-fj] <!-- there might be others here(?) -->.
+In this graphic, the canvas is split into a matrix where each off-diagonal element is a scatter plot of variable pairs, while the diagonal elements are either left blank or represent univariate structure.
+The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables.
+
+Similarly, parallel coordinates plots (PCP) can be used to explore correlation and co-linearities in multivariate data [@Inselberg1985-kf].
+In PCPs, variables are arranged along the (vertical) horizontal axis with vertical (horizontal) axes drawn representing the range of values for each variable.
+Points (observations) of the data are represented by drawing lines between each variables axis.
+If there is positive correlation between two variables the lines will be parallel, while for negative correlation the lines will cross.
+By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings is revealed, however the ordering of variables is important to reveal these features.
+<!-- anything else?
+Andrew's curves etc.?
+-->
 
@@
 (picture of the cluster data from Stuart's thesis)
+- Alternatives (Stuart) 
 -->
-- Alternatives (Stuart)
-
-(parallel coordinates, sploms, what tours are not)
 
@@
 - Bioinformatics (Stuart)
+
 - High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)

### 2021-03-01 paper.Rmd
paragraph on liminal

@@
 - liminal (Stuart)
+
+The liminal R package uses the tour to explore the quality of non-linear dimensionality reduction algorithms.
+The interface consists of two side by side views consisting of a scatter plot displaying a reduced form of the data, and an interactive tour.
+Controls such as play/stop/restart are implemented allowing a user to pause on interesting projections and return them to their R session for further analysis.
+Linked brushing is implemented on both views - if a users brush on the scatter plot view they can see if and how an algorithm like t-SNE has distorted distances in higher dimensional space, while if they brush on the tour view, the tour is paused and structure like cluster separation can be ascertained.
+
 - python (Di to follow up with Csaba)

### 2021-03-01 paper.Rmd
-

@@
 - Bioinformatics (Stuart)
-
 - High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)

### 2021-03-01 paper.Rmd
add sherry

@@
     email: earo.wang@auckland.ac.nz
+  - name: H. Sherry Zhang
+    address: |
+      | Department of Econometrics and Business Statistics,
+      | Monash University
+    affiliation: Monash University
+    email: huize.zhang@monash.edu
 title:

### 2021-03-01 paper.Rmd
dot points

@@
 - Bases notation (Sherry)
+
+$d$-dimensional projection in $p$-dimensional space
+
+Data: $X_{{n \times p}$, 
+Basis: $A_{p \times d}$, commonly 1d or 2d projection, orthonormality, more displays in the next section - can display a matrix A, Link to geometry properties.
+projection: $Y = X \cdot A$
+
 - Display (Sherry)
+
+[@Wickham2011-uz]
+
+1D: `display_hist`: histogram
+2D: `display_xy`, recent development `display_sage()` - will be detailed in the last section @Transformations on projections (Ursula)
+
 - Choosing targets, inc pp indexes (Sherry)
+
+grand tour, random, may miss interesting, take long to rotate to that specific angle, guided tour, along with projection pursuit, by generating target bases, interestingness
+
+interesting, index function, $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$, normality, theoretical result: [@Diaconis1984-mv], aslo peter huber projection pursuit
+
+index: 
+  - initial idea from Friedman, orthonormal function expansion (Do we need to review these old index contruction?)
+  - holes, central mass, PDA, LDA
+  - scagnostic indices
+  - skewness & kurtosis index loperfido (moment)
+  - chi-square @posse for detect spiral
+
 - Geodesic interpolation (Ursula)

### 2021-03-01 paper.Rmd
LaTeX fix (extra open `{`)

@@
 
-Data: $X_{{n \times p}$, 
+Data: $X_{n \times p}$, 
 Basis: $A_{p \times d}$, commonly 1d or 2d projection, orthonormality, more displays in the next section - can display a matrix A, Link to geometry properties.

### 2021-03-02 paper.Rmd
spinifex section init and bib.

@@
 
-- tourr (Di) and spinifex (Nick)
+- tourr (Di) 
+- spinifex (Nick)
+
+The R package spinifex [@spyrison_spinifex_2020] has 3 primary features: it produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package, renders (manual or other) tours as animations exportable to static .gif or interactive .html widgets, lastly it offers interactive shiny application that offers an graphical user interface to quickly sample tour features.
+
 - liminal (Stuart)

### 2021-03-03 paper.Rmd
add application sketch

@@
 - Bioinformatics (Stuart)
+
+In single cell RNA sequencing, scientists are interested in identifying novel cell types and understanding the relationships between cells or their developmental trajectory.
+To achieve this they perform cluster analysis on a counts matrix or principal components and embed the results via t-SNE and label points according the cluster label.
+One of the main advantages of t-SNE is the avoidance of over-plotting so clusters can be clearly identified on a scatter plot, however, this can come at the cost of interpretability as global distances are distorted.
+In @Laa2020-uv, we used radial transformations of the tour projections as an alternative to t-SNE that better preserves global structure while still retaining cluster topology.
+
 - High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)

### 2021-03-03 paper.Rmd
more notes

@@
 <!-- anything else?
-Andrew's curves etc.?
+Andrew's curves etc., heat?
 -->
@@
 - Alternatives (Stuart) 
+  - cluster heatmaps 
+  - data pipelines (?)
+  - point to other reviews
+- What do we mean by multivariate data visualisation?
+- multivariate categorical vis, tree maps etc.
+- subplots trellis literature
 -->
@@
 1D: `display_hist`: histogram
-2D: `display_xy`, recent development `display_sage()` - will be detailed in the last section @Transformations on projections (Ursula)
+2D: `display_xy`, recent development `display_sage()` - will be detailed in the last section Transformations on projections (Ursula)
 

### 2021-03-03 paper.Rmd
mapped out places for plots

--- 2021-03-03 paper.Rmd
@@
 It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
+
+> figure 1 illustrating tour views vs nonlinear dimension reduction for Stuart's cluster example
+
 While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualising structure in high dimensional data.
@@
 
+> figure 2 diagram illustrating a projection
+
 - Display (Sherry)
@@
 2D: `display_xy`, recent development `display_sage()` - will be detailed in the last section @Transformations on projections (Ursula)
+
+> figure 3 diagram illustrating different displays for different d
 
@@
 
+> figure 4 diagram illustrating different types of target generation
+
 - Geodesic interpolation (Ursula)
+
+> figure 5 diagram illustrating geodesic interpolation
+
 - Slices vs projections (Ursula)
+
+> figure 6 diagram illustrating projection vs slice
+
 - Transformations on projections (Ursula)
+
+> figure 7 diagram of burning sage transformation
 
@@
 - Stop/start/reverse
+
+> pull something from ggobi book?
+> figure 8 illustration
+
 - Manual (Nick)
+
+> something from nick's paper
+> figure 9 illustration
+
 - Linked brushing (Earo)
+
+> something from ggobi book
+> > figure 10: plots showing example
+
 - Spin-and-brush  (Ursula)
+
+> something from ggobi book
+> figure 11: plots showing example
 
@@
 - Model vis examples eg random forest (from Hadley, removing the blindfold) (Natalia)
+
+> figure 12 tour plot(s) of vote matrix and tour plot(s) of variables, with some annotation of cases?
+
 - Physics (Ursula)
+
+> figure 13
+
 - Bioinformatics (Stuart)
@@
 
+> figure 14
+
 - High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)
+
+> figure 15
+
+- neural networks (from Scheidegger - Stuart)
+
+> figure 16
 

### 2021-03-03 paper.Rmd
umap example

@@
 
+- umap (Mingwei Li paper - Stuart)
+
+> figure 17
+
 # Discussion

### 2021-03-06 paper.Rmd
paragraph on bases notation

@@
 
-$d$-dimensional projection in $p$-dimensional space
-
-Data: $X_{n \times p}$, 
-Basis: $A_{p \times d}$, commonly 1d or 2d projection, orthonormality, more displays in the next section - can display a matrix A, Link to geometry properties.
-projection: $Y = X \cdot A$
+Some notations are defined for describing $d$-dimensional projections in $p$-dimensional space.
+Let $X_{n \times p}$ be the original data whose projections are of our interest.
+A projection basis, $A_{p \times d}$, characterises the direction from which the high-dimensional data are projected.
+A constraint that projection bases needs to satisfy is the orthonormality condition (which requires each axis to be unit vector and perpendicular to each other).
+This gives the basis space a geometrical structure of a Stiefel manifold.
+The projection of the original data on a basis hence can be defined as $Y = X \cdot A$.
+The projection is usually in 1D or 2D for the purpose of visual display.
 

### 2021-03-07 paper.Rmd
paragraph on pp index

@@
 
+[grand/ guided / manual tour or 1D/ 2D display in tourr???]
+
+grand tour, random, may miss interesting, take long to rotate to that specific angle, guided tour, along with projection pursuit, by generating target bases, interestingness
+
 > figure 3 diagram illustrating different displays for different d
@@
 
-grand tour, random, may miss interesting, take long to rotate to that specific angle, guided tour, along with projection pursuit, by generating target bases, interestingness
-
-interesting, index function, $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$, normality, theoretical result: [@Diaconis1984-mv], aslo peter huber projection pursuit
-
-index: 
-  - initial idea from Friedman, orthonormal function expansion (Do we need to review these old index contruction?)
-  - holes, central mass, PDA, LDA
-  - scagnostic indices
-  - skewness & kurtosis index loperfido (moment)
-  - chi-square @posse for detect spiral
+To define whether a projection is interesting or not in the guided tour, an index function is used.
+An index function defines a surjection function from the projection $Y$ in $ \mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
+For the high-dimensional data, an interesting projection is the one that deviates from normality, as discussed in @Diaconis1984-mv and @Huber1985-zq [expand more here].
+Numerous index functions have been developed since the initial index proposed by @Friedman1974-ck: <!-- entropy-based index?
+--> holes and central mass index [@Cook2008-nl] computes the deviation of the projected data from a normal distribution; @Posse1995-ih proposed a chi-square index for detecting the spiral structure in the data;  kurtosis and skewness indices are raised in @Loperfido2020-xj and @Loperfido2018-ga to detect outliers in financial time series; and most recently, scagnostic indices [@Laa2020-zj] provides eight measures for summarising information between pairs of variables in scatter plot matrices.
 

### 2021-03-09 paper.Rmd
added bullet points for brushing and linking from "ggobi"
-code for figure of radial manual tour. -fix other space issue in \mathbb -NS TODO: add fig caption and w/h, and write up

@@
 To define whether a projection is interesting or not in the guided tour, an index function is used.
-An index function defines a surjection function from the projection $Y$ in $ \mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
+An index function defines a surjection function from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
 For the high-dimensional data, an interesting projection is the one that deviates from normality, as discussed in @Diaconis1984-mv and @Huber1985-zq [expand more here].
@@
 
-> something from nick's paper
+> something from nick's paper.
+> NS: something like figure 3, exlpaining radial tour on lda.
 > figure 9 illustration
+
+```{r}
+##TODO, need to make caption, tag, and optimize figure width/height
+source("./R/make_radial_tour_fig.r")
+make_radial_tour_fig()
+```
 
@@
 > something from ggobi book
+* Brushing as database query
+* Brushing as conditioning
+* Brushing as geometric sectioning
+* Persistent vs. transient 
 > > figure 10: plots showing example

### 2021-03-09 R/make_radial_tour_fig.r

@@ -1,0 +1,73 @@
+### Nicholas Spyrison 08 March 2021
+### Recreate figure 3 from spinifex paper. Silo'd, with minimal dependencies.
+
+#```{r step3, echo=F, warning=F, out.width='100%', fig.width=8, fig.height=2, fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original. "}
+# install.packages("spinifex") ## Made with Cran spinifex_0.2.7
+require("spinifex")
+require("tourr")
+require("ggplot2")
+require("magrittr")
+my_theme <- list(scale_color_brewer(palette = "Dark2"),
+                 theme_void(), 
+                 theme(legend.position ="none"))
+# ## Testing a slight change to scale to see if that works better as framse change
+# scale_axes <-
+#   function (x, position = "center", to = data.frame(x = c(-1L, 1L), y = c(-1L, 1L))){
+#     if(is.null(to)) to <- data.frame(x = c(-1L, 1L), y = c(-1L, 1L))
+#     x_to <- c(min(to[, 1L]), max(to[, 1L]))
+#     y_to <- c(min(to[, 2L]), max(to[, 2L]))
+#     xdiff <- diff(x_to)
+#     ydiff <- diff(y_to)
+#     xcenter <- mean(x_to)
+#     ycenter <- mean(y_to)
+#     if(position == "center") {
+#       scale <- 0.3 * max(ydiff, xdiff)
+#       xoff <- xcenter
+#       yoff <- ycenter
+#     }
+#     x[, 1L] <- scale * x[, 1L] + xoff
+#     x[, 2L] <- scale * x[, 2L] + yoff
+#     return(x)
+#   }
+## Doesn't seem to make the change I expected. may want tyo test in dev version.
+
+make_radial_tour_fig <- function(){
+  dat  <- scale_sd(flea[, 1:6])
+  clas <- factor(flea$species)
+  bas  <- MASS::lda(dat, grouping = clas)$scaling %>% tourr::orthonormalise()
+  mvar <- which(abs(bas[, 1]) == max(abs(bas[, 1])))
+  
+  if(F) ## Manual check the starting LDA basis
+    view_frame(bas, dat, mvar,
+               aes_args = list(color = clas, shape = clas),
+               ggproto = my_theme)
+  
+  ang <- .29
[…]

### 2021-03-09 data/clusters.R
code to create the multi cluster example

@@ -28,10 +28,14 @@ tour_tbl <- dplyr::bind_cols(as.data.frame(Y),
 
 
 tsne_y <- Rtsne::Rtsne(Y)
-embed_tbl <- tidy_tsne(tsne_y, 
-                       list(cluster = factor(X_tbl[["cluster"]])))
-colnames(embed_tbl) <- c("tsneX", "tsneY", "cluster")
-
+#embed_tbl <- tidy_tsne(tsne_y, 
+#                       list(cluster = factor(X_tbl[["cluster"]])))
+#colnames(embed_tbl) <- c("tsneX", "tsneY", "cluster")
+tsne_df <- tibble(tsneX = tsne_y$Y[,1], 
+                    tsneY = tsne_y$Y[,2], 
+                    cluster = X_tbl$cluster)
+ggplot(tsne_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
+  geom_point() + coord_equal()
 
 # multichallenge data set
 url <- "http://ifs.tuwien.ac.at/dm/download/multiChallenge-matrix.txt"
@@ -50,3 +54,8 @@ tsne_multi <- Rtsne::Rtsne(multi[, -1])
 embed_multi <- tidy_tsne(tsne_multi, list(group = multi[["group"]]))
 colnames(embed_multi) <- c("tsneX", "tsneY", "group")
 
+tsne_multi_df <- tibble(tsneX = tsne_multi$Y[,1], 
+                  tsneY = tsne_multi$Y[,2], 
+                  cluster = multi$group)
+ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
+  geom_point() + coord_equal()

### 2021-03-09 data/clusters.R
code to create the multi cluster example

@@ -51,8 +51,8 @@ multi <- read_tsv(url, col_names = FALSE) %>%
 
 set.seed(143280)
 tsne_multi <- Rtsne::Rtsne(multi[, -1])
-embed_multi <- tidy_tsne(tsne_multi, list(group = multi[["group"]]))
-colnames(embed_multi) <- c("tsneX", "tsneY", "group")
+#embed_multi <- tidy_tsne(tsne_multi, list(group = multi[["group"]]))
+#colnames(embed_multi) <- c("tsneX", "tsneY", "group")
 
 tsne_multi_df <- tibble(tsneX = tsne_multi$Y[,1], 
                   tsneY = tsne_multi$Y[,2], 

### 2021-03-09 data/clusters.R
move code over

@@ -6,6 +6,13 @@ library(Rtsne)
 library(dplyr)
 library(readr)
 
+tidy_tsne <- function(model, data) {
+  enframe <- as.data.frame(model[["Y"]])
+  colnames(enframe) <- c("tsneX", "tsneY", "tsneZ")[seq_len(ncol(enframe))]
+  dplyr::bind_cols(enframe, data)
+}
+
+
 # 5 gaussian clusters with equal covariance
 set.seed(63968)
 m <- 500

### 2021-03-09 paper.Rmd
working on initial example

@@
 keywords:
-  formatted: ["tours", "data visualisation"]
+  formatted: ["tours", "data visualisation", "high-dimensional data", "data science", "exploratory data analysis"]
 output: 
@@
 
-> figure 1 illustrating tour views vs nonlinear dimension reduction for Stuart's cluster example
+\begin{figure*}[htp]
+         \includegraphics[width=0.5\textwidth]{figures/cluster_example_r.pdf}
+     \vfill
+   \begin{minipage}{\linewidth}     
+   \centering
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour1_r.pdf}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour2_r.pdf}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour3_r.pdf}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour4_r.pdf}
+   \end{minipage}%
+\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and projections seen during a tour (B.1-4).
+There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied.
+This can be seen in the tour projections.
+The two green clusters are spherical in shape, and very distance from the brown clusters.
+Three of the brown clusters are very close to each other (just visible in B.1), and all brown clusters are elliptical.
+The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the dimension reduction.}
+\label{fig:tsne-tour}
+\end{figure*}
 

### 2021-03-09 data/clusters.R

@@ -52,7 +52,7 @@ multi <- read_tsv(url, col_names = FALSE) %>%
     key = rep(1:5, each = 200L)
   ) %>%
   filter(key %in% c(1, 4)) %>% 
-  mutate(group = rep(c("hierachical", "flat"), each = 200L)) %>% 
+  mutate(group = rep(c("hierarchical", "flat"), each = 200L)) %>% 
   select(group, X1:X10)
 
 
@@ -65,4 +65,25 @@ tsne_multi_df <- tibble(tsneX = tsne_multi$Y[,1],
                   tsneY = tsne_multi$Y[,2], 
                   cluster = multi$group)
 ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
-  geom_point() + coord_equal()
+  geom_point() + coord_equal() +
+  scale_colour_brewer("", palette="Dark2")
+
+# tourr
+library(tourr)
+library(RColorBrewer)
+animate_xy(multi[,2:11], col=multi$group)
+clrs = brewer.pal(3, "Dark2")[as.numeric(factor(multi$group))]
+render_gif(multi[,2:11], 
+           grand_tour(),
+           display_xy(col=clrs, axes="bottomleft"), 
+           frames = 200,
+           "figures/cluster_example.gif")
+
+render_gif(multi[,2:11], 
+           grand_tour(),
+           display_xy(col=clrs, axes="bottomleft", half_range = 0.5, center = TRUE), 
+           frames = 200,
+           "figures/cluster_example_zoom.gif")
+
+
+

### 2021-03-09 paper.Rmd
manual tour figure: R, .png, and paper changes.

@@
 
-```{r}
-##TODO, need to make caption, tag, and optimize figure width/height
-source("./R/make_radial_tour_fig.r")
-make_radial_tour_fig()
+Example fig ref: see figure \ref{fig:manual-tour}.
+
+```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
+We initalise the start basis to the Linear Discriminant Analysis and select the variable with the largest contribution separating the cluster shown as green circles.
+By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.'}
+
+if(F) ## Creation and saving figure at:
+  file.edit("./R/make_radial_tour_fig.r")
+
+knitr::include_graphics("./figures/fig_radial_manual_tour.png")
 ```

### 2021-03-09 R/make_radial_tour_fig.r

@@ -1,8 +1,6 @@
 ### Nicholas Spyrison 08 March 2021
 ### Recreate figure 3 from spinifex paper. Silo'd, with minimal dependencies.
 
-#```{r step3, echo=F, warning=F, out.width='100%', fig.width=8, fig.height=2, fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original. "}
-# install.packages("spinifex") ## Made with Cran spinifex_0.2.7
 require("spinifex")
 require("tourr")
 require("ggplot2")
@@ -10,32 +8,13 @@ require("magrittr")
 my_theme <- list(scale_color_brewer(palette = "Dark2"),
                  theme_void(), 
                  theme(legend.position ="none"))
-# ## Testing a slight change to scale to see if that works better as framse change
-# scale_axes <-
-#   function (x, position = "center", to = data.frame(x = c(-1L, 1L), y = c(-1L, 1L))){
-#     if(is.null(to)) to <- data.frame(x = c(-1L, 1L), y = c(-1L, 1L))
-#     x_to <- c(min(to[, 1L]), max(to[, 1L]))
-#     y_to <- c(min(to[, 2L]), max(to[, 2L]))
-#     xdiff <- diff(x_to)
-#     ydiff <- diff(y_to)
-#     xcenter <- mean(x_to)
-#     ycenter <- mean(y_to)
-#     if(position == "center") {
-#       scale <- 0.3 * max(ydiff, xdiff)
-#       xoff <- xcenter
-#       yoff <- ycenter
-#     }
-#     x[, 1L] <- scale * x[, 1L] + xoff
-#     x[, 2L] <- scale * x[, 2L] + yoff
-#     return(x)
-#   }
-## Doesn't seem to make the change I expected. may want tyo test in dev version.
 
+### Create the 
 make_radial_tour_fig <- function(){
   dat  <- scale_sd(flea[, 1:6])
   clas <- factor(flea$species)
   bas  <- MASS::lda(dat, grouping = clas)$scaling %>% tourr::orthonormalise()
-  mvar <- which(abs(bas[, 1]) == max(abs(bas[, 1])))
+  mvar <- 4 ## aede1, primary disting of 1 cluster.
   
   if(F) ## Manual check the starting LDA basis
     view_frame(bas, dat, mvar,
@@ -52,22 +31,30 @@ make_radial_tour_fig <- function(){
       print(msg)
     }
   }
[…]

### 2021-03-09 paper.Rmd
spinifex citiations, bib file, and manual tour .png

@@
 
-The R package spinifex [@spyrison_spinifex_2020] has 3 primary features: it produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package, renders (manual or other) tours as animations exportable to static .gif or interactive .html widgets, lastly it offers interactive shiny application that offers an graphical user interface to quickly sample tour features.
+The R package spinifex [@spyrison_spinifex_2020] has 3 primary features: it produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package, renders (manual or other) tours as animations exportable to static .gif (gganimate [@pedersen_gganimate_2020] package) or interactive .html widgets (plotly [@sievert_interactive_2020] package), lastly it offers interactive shiny application that offers an graphical user interface to quickly sample tour features.
 

### 2021-03-09 paper.Rmd
clustering example complete

@@
 
-\begin{figure*}[htp]
+Nonlinear dimension reduction techniques such as t-SNE have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary.
+However, it is only a summary, and it likely involves substantial warping of the original data space.
+Using the tour along with these techniques can illuminate the nature of the warped space, and reveal other structure lost in the dimension reduction.
+Figure \ref{fig:tsne-tour} illustrates the difference in what can be learned from a tour in comparison with nonlinear dimension reduction using t-SNE.
+The 10-dimensional data comes from @Rauber2009-kk.
+The t-SNE view (plot A) shows six clearly separated clusters, all spherical with different sizes.
+The tour shows that the clusters look very little like this in the full data space.
+The clusters are at various distances apart and very different sizes.
+In the four projections from the tour this can be seen.
+The two green clusters are large and spherical, and far from the brown clusters.
+The brown clusters have one larger one, and three smaller, very close to each other.
+All of these are elliptical, which means that they have very little variability, actually no variability if you watch the full tour, in some of the 10 dimensions.
+This gives some deeper perspective to what is learned from t-SNE, and illustrates what the t-SNE dimension reduction has done: it has found small gaps between points and expanded these gaps to yield the representation.
+It should be noted, though, that the t-SNE view clearly reveals the presence of six distinct clusters.
+With this information is the invitation to look closer at the data in the tour, to see, that yes, indeed, there are three, tiny, tiny clusters very close to each other.
+
+\begin{figure*}[!t]
          \includegraphics[width=0.5\textwidth]{figures/cluster_example_r.pdf}

### 2021-03-09 paper.Rmd
replacement for manual tour figures

@@
 
+\begin{figure*}[!h]
+   \centering
+         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl1_r.pdf}
+         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl3_r.pdf}
+         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl4_r.pdf}
+\caption{Four projections from a manual tour where variable "bl" is being rotated out of the projection (orange line).
+When this variable is removed the two light green clusters merge, which informs us that "bl" is an important variable for distinguishing between these two groups.}
+\label{fig:manual-tour}
+\end{figure*}
+
 ```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
 We initalise the start basis to the Linear Discriminant Analysis and select the variable with the largest contribution separating the cluster shown as green circles.
-By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.'}
+By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.', eval=FALSE}
 

### 2021-03-09 paper.Rmd
Manual tour and ref.

@@
 
-- Manual (Nick)
+## Manual tour (Nick)
 
-> something from nick's paper.
-> NS: something like figure 3, exlpaining radial tour on lda.
-> figure 9 illustration
+Manual tours [@cook_manual_1997; @spyrison_spinifex_2020] offer a means to interactively control the contribution of a single variable on the projection plane.
+This is particularly useful for exploring a projection once a feature of interest has been identified.
+Manual tours can then be employed to explore local area of, or test the structure of the feature, with respect to a selected variable.
 
-Example fig ref: see figure \ref{fig:manual-tour}.
+The details of performing a manual tour are described by @spyrison_spinifex_2020, but high-level unit-interaction between each frame is as follows:
+
+__Input:__ Starting basis, manipulation variable, an in-plane rotation angle -- $\theta$, an out-of-plane rotation angle -- $\phi$  $\\$
+__Output:__ Rotated basis $\\$
+__Steps:__
+
+1. Initialise a manipulation dimension onto the start basis. Given the manipulation variable a full contribution and orthonormalise this space. 
+2. Post-multiply the space, by a rotation matrix, as solved for in application of Rodrigues’ formula [@rodrigues_lois_1840].
+3. Truncate the manipulation dimension.
+
+Figure \ref{fig:manual-tour} illustrates the manual tour, by radially manipulating the contribution of variable that is crucial, to explain the separation of a particular cluster.
+To start we first identify a basis by orthonormalising the linear discriminant.
+We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, compeletely out of the projection, and then returned to it's original contribution.
+
+```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
+We initalise the start basis to the linear discriminant analysis and select the variable with the largest contribution separating the cluster shown as green circles.
+By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.'}
+if(F) ## Creation and saving figure at:
+  file.edit("./R/make_radial_tour_fig.r")
+knitr::include_graphics("./figures/fig_radial_manual_tour.png")
+```
 
@@
 \end{figure*}
-
-```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
-We initalise the start basis to the Linear Discriminant Analysis and select the variable with the largest contribution separating the cluster shown as green circles.
-By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.', eval=FALSE}
-
-if(F) ## Creation and saving figure at:
-  file.edit("./R/make_radial_tour_fig.r")
-
-knitr::include_graphics("./figures/fig_radial_manual_tour.png")
-```
 

### 2021-03-09 paper.Rmd
Putting down thoughts on my parts

--- 2021-03-09 paper.Rmd
@@
 
+Interpolation is either between frames, or between planes
+Interpolation between planes is preferred to avoid distracting effects (rotation in the plane), exception would be projection pursuit indexes that are sensitive to rotation (but that should be avoided when possible)
+Geodesic interpolation uses SVD to interpolate along the shortest path between planes
+
+XXX need to decide how much math to put in, if/which picture to use (probably something from @Buja2005-cx)
+
 > figure 5 diagram illustrating geodesic interpolation
@@
 
+Projections can hide certain features, sections can provide additional information @Furnas1994-dz
+Sectioning a high-dimensional space in systematic manner is challenging, there are a lot of choices we need to make
+Typical approach is interactive: brushing to select sections in one view, highlight points in a second view, e.g. a tour display
+Slice tour approach: use distance from the projection plane (through a selected center point), update the section systematically with the projection plane @Laa2020-js
+Summary of applications: viewing geometric shapes, model visualization
+Connection with projection pursuit: section pursuit @Laa2020-ni
+
+XXX figure should probably be combination of Fig 1 and 2 in @Laa2020-js
+
 > figure 6 diagram illustrating projection vs slice
@@
 - Transformations on projections (Ursula)
+
+Transformations of the data are common in the preprocessing, but sometimes we want to transform the data after projecting onto lower dimensions
+Example would be to address piling effects that are a result of projecting form high-dimensional space down to low-dimensions
+A solution to this problem was presented in the sage display @Laa2020-uv and is using a radial transformation of the projected data to avoid piling near the center
+Summary of applications: better understanding of clustering in high-dimensional space (discussed below), resolving small features near the center
+
+XXX I would like Fig 5 of @Laa2020-uv as an illustration here
 
@@
 
+Tour for visual cluster analysis, identification of outlying points
+Maybe the physics example from the ggobi book: find group in one view, highlight, restart tour, and so on, in the end we can review the tour with groups highlighted in colour to get an overview of how the data clusters
+Requires persistent brushing with multiple groups
+
 > something from ggobi book
@@
 
+Physics models often consider $\mathcal{O}(10)$ free parameters and are compared to a much larger number of experimental observables ($\mathcal{O}(100)$).
+The comparison typically relies on numeric computation of the model predictions for all observables, potentially obscuring the nature of their dependence on the model parameters.
+Here, multivariate visualisation methods can provide new insights.
+This was demonstrated in @Cook2018-jm.
+For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point.
+For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterix symbol.
+
+XXX need to select plots, also based on how much space we want to use
+
 > figure 13

### 2021-03-10 paper.Rmd
commented out nicks flea data plot

@@
 We initalise the start basis to the linear discriminant analysis and select the variable with the largest contribution separating the cluster shown as green circles.
-By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.'}
+By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.', eval=FALSE}
 if(F) ## Creation and saving figure at:

### 2021-03-10 paper.Rmd
add neural network section

--- 2021-03-09 paper.Rmd
@@
 
+The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg].
+They proposed a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
+They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
+For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance.
+Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
+
 > figure 16

### 2021-03-10 paper.Rmd
plot markers

@@
 
+> Maybe we should clarify What is high-dimensional data? here
+
 The focus of this review will be on linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
@@
 
-> figure 2 diagram illustrating a projection
+> figure 2 diagram illustrating a projection: Di thinks making a Huber plot (2D data - 1D projection) here with projected data on a couple of directions is ideal
 

### 2021-03-11 paper.Rmd
add umap section
fixes references formatting

@@
 > something from ggobi book
+
 * Brushing as database query
@@
 * Persistent vs. transient
+
 > > figure 10: plots showing example
@@
 Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
+@li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
 
-> figure 16
-
-- umap (Mingwei Li paper - Stuart)
-
-> figure 17
+> figure 16 panel a and b
 

### 2021-03-11 paper.Rmd
add sage frames and caption

@@
 
-> figure 14
+\begin{figure}
+         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-001.png}
+         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-016.png}
+         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-038.png}
+         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-074.png}
+
+}
+\caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation described in @Laa2020-uv.
+Here the sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}\label{fig:mouse-cluster}
+\end{figure}
 

### 2021-03-11 paper.Rmd
add interaction figures

@@
 
+\begin{figure*}[!h]
+   \centering
+         \includegraphics[width=0.5\textwidth]{figures/liminal-stop-pause-refresh.png}
+\caption{The clustering example in figure \ref{fig:tsne-tour} using the liminal interface.
+Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.}
+\label{fig:manual-tour}
+\end{figure*}
+
 > pull something from ggobi book?
@@
 
-\begin{figure}
+\begin{figure*}[!h]
+  \centering
            \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-001.png}
@@
 
-}
 \caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation described in @Laa2020-uv.
-Here the sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}\label{fig:mouse-cluster}
-\end{figure}
+Here the sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}
+\label{fig:mouse-cluster}
+\end{figure*}
 

### 2021-03-11 paper.Rmd
fix captioning
update reference
update section
getting artifacts
fix catpion

@@
 
-- Stop/start/reverse
+## Basic Interactions
 
-\begin{figure*}[!h]
+- Stop/start/reverse (Stuart)
+
+\begin{figure*}
    \centering
@@
 Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.}
-\label{fig:manual-tour}
+\label{fig:tour-controls}
 \end{figure*}
-
-> pull something from ggobi book?
-> figure 8 illustration
 
@@
 
-\begin{figure*}[!h]
+\begin{figure*}
   \centering
@@
 
-\caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation described in @Laa2020-uv.
-Here the sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}
+\caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation.
+The sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}
 \label{fig:mouse-cluster}

### 2021-03-14 paper.Rmd
start mapping out what's left
conflict resolved and control image made larger

@@
 
+The rest of the paper is structured as follows...
+
 # Tours for high dimensional visualization
@@
    \centering
-         \includegraphics[width=0.5\textwidth]{figures/liminal-stop-pause-refresh.png}
-\caption{The clustering example in figure \ref{fig:tsne-tour} using the liminal interface.
+         \includegraphics[width=0.9\textwidth]{figures/liminal-stop-pause-refresh.png}
+\caption{The clustering example from Figure \ref{fig:tsne-tour} using the liminal interface.
 Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.}

### 2021-03-14 paper.Rmd
added cubes example

@@
 
-The focus of this review will be on linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+The term "high-dimensional" here means Euclidean space.
+Figure \ref{fig:cubes} shows a way to imagine this.
+It shows a sequence of cube wireframes, ranging from 1D through to 6D, where beyond 2D is a linear projection of the cube.
+As dimension increase, a new orthogonal axis is added.
+For cubes, this is achieved by doubling the cube: a 2D is two 1D cubes, a 3D is two 2D cubes, and so forth.
+
+\begin{figure*}[!h]
+\centerline{\includegraphics[width=0.95\textwidth]{figures/cubes.png}}
+\caption{Illustrating what is meant by "high-dimensional" in this paper, and a "linear projection".
+From a sequence of increasing dimension cubes, from 1D to 6D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles}
+\label{fig:cubes}
+\end{figure*}
+
+The focus of this review will be on visualising high-dimensional data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
 The reason being that it is not feasible to adequately the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.

### 2021-03-15 paper.Rmd
add panel for neural network examples

@@
 
-> figure 16 panel a and b
+\begin{figure*}
+   \centering
+         \includegraphics[width=0.5\textwidth]{figures/scheidigger_nn_tour.png}
+          \includegraphics[width=0.5\textwidth]{figures/scheidigger_umap_tour.png}
+\caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}
+\label{fig:tour-nn}
+\end{figure*}
 

### 2021-03-15 paper.Rmd
add Laa transformation figure

@@
 
-XXX I would like Fig 5 of @Laa2020-uv as an illustration here
-
-> figure 7 diagram of burning sage transformation
+\begin{figure*}
+   \centering
+         \includegraphics[width=0.9\textwidth]{figures/laa_radial_transformation.pdf}
+\caption{An illustration of the radial transformation. Each panel represents linear projections of $p$ dimensional equidistant circles. As $p$ increases the circles get pushed further to the edge.}
+\label{fig:tour-transformation}
+\end{figure*}
 

### 2021-03-15 paper.Rmd
add slice tour figure
adjust layout

@@
 
-XXX figure should probably be combination of Fig 1 and 2 in @Laa2020-js
-
-> figure 6 diagram illustrating projection vs slice
+\begin{figure*}
+   \centering
+         \includegraphics[width=0.9\textwidth]{figures/slice_tour_left.jpg}
+         \includegraphics[width=0.9\textwidth]{figures/slice_tour_right.jpg}
+\caption{Ursula to write}
+\label{fig:tour-slice}
+\end{figure*}
 
@@
    \centering
-         \includegraphics[width=0.5\textwidth]{figures/scheidigger_nn_tour.png}
-          \includegraphics[width=0.5\textwidth]{figures/scheidigger_umap_tour.png}
+         \includegraphics[width=0.45\textwidth]{figures/scheidigger_nn_tour.png}
+          \includegraphics[width=0.45\textwidth]{figures/scheidigger_umap_tour.png}
 \caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}

### 2021-03-15 paper.Rmd
add section header

@@
 By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings is revealed, however the ordering of variables is important to reveal these features.
-<!-- anything else?
-Andrew's curves etc., heat?
+
+- cluster heatmaps
+
+- subplots trellis literature
+
+- data pipelines (?)
+
+<!-- Original tours (Di)
+
+- What do we mean by multivariate data visualisation?
+- multivariate categorical vis, tree maps etc.
 -->
 
-<!-- Original tours (Di)
-- Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
-(picture of the cluster data from Stuart's thesis)
-- Alternatives (Stuart) 
-  - cluster heatmaps 
-  - data pipelines (?)
-  - point to other reviews
-- What do we mean by multivariate data visualisation?
-- multivariate categorical vis, tree maps etc.
-- subplots trellis literature
--->
+The rest of the review is structured as follows: in section \@ref(notation) we define the notation and components of a tour display, in section \@ref(we review the importance of interactive user ..
 
-The rest of the paper is structured as follows...
-
-# Tours for high dimensional visualization
+# Tours for high dimensional visualization {#notation}
 
@@
 
-# Ways of interacting
+# Ways of interacting {#interaction}
 
@@
 
-# Software
+# Software {#software}
+
+- Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
 
@@
 
-# Application
+# Applications {#application}
 
@@
 
-# Discussion
+# Discussion {#discussion}
 

### 2021-03-15 paper.Rmd
add heatmaps

@@
 
-- cluster heatmaps
+The heatmap display is widely used to visualise cluster structure in bioinformatics.
+In a heatmap each element of a numeric matrix is displayed as a single colored square pixel on the canvas.
+The hue of the color is mapped to the magnitude of the element's values.
+A subtlety of using the heatmap is determining how to order the rows and columns of the matrix called seriation.
+In practice, two way hierarchical clustering  or singular value decomposition is used to order rows and columns.
+See @Wilkinson2009-pj for a comprehensive history.
 
@@
 
-The rest of the review is structured as follows: in section \@ref(notation) we define the notation and components of a tour display, in section \@ref(we review the importance of interactive user ..
+The rest of the review is structured as follows: in section \@ref(notation) we define the notation and components of a tour display.
+By its nature a tour is most effective to analyst when combined with interactivity; in section \@ref(interaction) we review the components of user interfaces for manipulating tour views.
+The implementation of tour paths in statistical software is reviewed in section \@ref(software), where we discuss the history and state of the art packages for computing them.
+Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning and applied statistics.
+Finally, in section \@ref(discussion) we point to future research directions for tours.
 

### 2021-03-15 paper.Rmd
fix author formatting

@@
     address: |
-      | Department of Econometrics and Business Statistics,
-      | Monash University
+      | Department of Econometrics and Business Statistics, Monash University
     email: stuart.a.lee@monash.edu
@@
     address:  |
-      | Institute of Statistics,
-      | University of Natural Resources and Life Sciences
+      | Institute of Statistics, University of Natural Resources and Life Sciences
     email:  ursula.laa@boku.ac.at
@@
     address: |
-      | Instituto de Estadística (IESTA), 
-      | Universidad de la República
+      | Instituto de Estadística (IESTA), Universidad de la República
     affiliation: Universidad de la República
@@
     address: |
-      | Faculty of Information and Technology 
-      | Monash University
+      | Faculty of Information and Technology, Monash University
     affiliation: Monash University
@@
     address: |
-      | Department of Econometrics and Business Statistics,
-      | Monash University
+      | Department of Econometrics and Business Statistics, Monash University
     affiliation: Monash University
@@
     address: |
-      | Department of Statistics,
-      | The University of Auckland
+      | Department of Statistics, The University of Auckland
     affiliation: The University of Auckland
@@
 
+<!-- Original tours (Di)
+Not sure how much of this wee need to go into...
 - subplots trellis literature
-
 - data pipelines (?)
-
-<!-- Original tours (Di)
-
 - What do we mean by multivariate data visualisation?
@@
 
-# Applications {#application}
+# Applications {#applications}
 
@@
 
+* why tours are useful and what we learn from them?
+
 # References {-}

### 2021-03-15 paper.Rmd
some discussion points
a note
Figure physics example

@@
 For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point.
-For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterix symbol.
+For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterisk symbol in Fig.
+\ref{fig:physics}.
 
-XXX need to select plots, also based on how much space we want to use
+\begin{figure*}
+  \centering
+           \includegraphics[width=0.45\textwidth]{figures/allcenter-301.png}
+         \includegraphics[width=0.45\textwidth]{figures/jetCluster-027.png}
 
-> figure 13
+\caption{Selected frames from using a grand tour of the physics data.
+Left: the three different types of measurements (shown in different colors) are aligned along different directions in parameter space.
+Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.}
+\label{fig:physics}
+\end{figure*}
 
@@
 
-- neural networks (from Scheidegger - Stuart)
+- neural networks (this could go into model vis?)
 
@@
 
-* why tours are useful and what we learn from them?
+* why tours are useful and what do we learn from them?
+* what research is needed?
 

### 2021-03-16 paper.Rmd
a proposed figure 2 and writing on display section

@@
 A projection basis, $A_{p \times d}$, characterises the direction from which the high-dimensional data are projected.
-A constraint that projection bases needs to satisfy is the orthonormality condition (which requires each axis to be unit vector and perpendicular to each other).
-This gives the basis space a geometrical structure of a Stiefel manifold.
-The projection of the original data on a basis hence can be defined as $Y = X \cdot A$.
-The projection is usually in 1D or 2D for the purpose of visual display.
+Projection bases needs to satisfy an orthonormality condition, which requires each axis to be unit vector and perpendicular to each other, and this gives the basis space a geometrical structure of a Stiefel manifold.
+With a data matrix and a projection basis, a projection can be defined as $Y = X \cdot A$ for visual inspection.
 
@@
 
+# ```{r}
+# PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
+# ```
+
 - Display (Sherry)
 
-[@Wickham2011-uz]
-
-1D: `display_hist`: histogram
-2D: `display_xy`, recent development `display_sage()` - will be detailed in the last section Transformations on projections (Ursula)
-
-[grand/ guided / manual tour or 1D/ 2D display in tourr???]
+Various displays are available to show projections in 1D, 2D, or higher dimensions.
+A 1D projection shows the data in a histogram as if a hand shadow being projected onto a wall.
+Two-d projections are usually displayed with scatterplot since it is the most common plot for showing bivariate relationship.
+Higher dimensional projections can be shown with multivariate display like scatterplot matrix or parallel coordinates plot.
+Figure <!--figure 3 --> shows a 1D display with histogram, a 2D display with scatterplot, and a 5D display with a parallel coordinates plot.
+More display methods can be found in [@Wickham2011-uz].
 

### 2021-03-16 paper.Rmd
first pass at abstract, fix up latex errors

@@
 abstract: > 
-  Abstract goes here.
+  This article discusses a high-dimensional visualization technique called the tour, which can be used to view data in more than three dimensions.
+We review the theory and history behind the technique, as well as modern developments and applications of the tour that are being found across the natural sciences and machine learning.
 keywords:
-  formatted: ["tours", "data visualisation", "high-dimensional data", "data science", "exploratory data analysis"]
+  formatted: ["tours", "data visualization", "high-dimensional data", "data science", "exploratory data analysis"]
 output: 
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!ht]
    \centering
@@
          \includegraphics[width=0.45\textwidth]{figures/scheidigger_nn_tour.png}
-          \includegraphics[width=0.45\textwidth]{figures/scheidigger_umap_tour.png}
+          \includegraphics[width=0.45\textwidth, keepaspectratio=true]{figures/scheidigger_umap_tour.png}
 \caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}

### 2021-03-16 paper.Rmd
tweaks

@@
   This article discusses a high-dimensional visualization technique called the tour, which can be used to view data in more than three dimensions.
-We review the theory and history behind the technique, as well as modern developments and applications of the tour that are being found across the natural sciences and machine learning.
+We review the theory and history behind the technique, as well as modern software developments and applications of the tour that are being found across the sciences and machine learning.
 keywords:

### 2021-03-16 paper.Rmd
add spin and brush figure

@@
 
-> something from ggobi book
-> figure 11: plots showing example
+\begin{figure}[htp]
+\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp1.pdf}}
+ {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp2.pdf}}
+ {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp5.pdf}}}
+\smallskip
+\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp7.pdf}}
+ {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp8.pdf}}
+ {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp9.pdf}}}
+\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp10.pdf}}
+  {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp11.pdf}}
+  {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp13.pdf}}}
+\caption{Spin and brush reprinted from ggobi book}
+\label{prim7-tour}
+\end{figure}
 

### 2021-03-16 paper.Rmd
Paragraphs on interpolation.

@@
 
-Interpolation is either between frames, or between planes
-Interpolation between planes is preferred to avoid distracting effects (rotation in the plane), exception would be projection pursuit indexes that are sensitive to rotation (but that should be avoided when possible)
-Geodesic interpolation uses SVD to interpolate along the shortest path between planes
+The smooth animation between the selected target frame is generated via a step-wise rotation in the high-dimensional space.
+Different methods to derive this rotation were described in @Buja2005-cx, and we can distinguish two types of interpolation.
+When interpolating between frames we rotate the starting basis onto the exact basis representation selected as a basis, and @Buja2005-cx describes three different methods to achieve this (based on decompositions of orthogonal matrices, Givens decompositions, or Householder decompositions).
+Alternatively, when interpolating between planes the rotation is constructed towards any basis that represents the plane corresponding to the target frame, such that it follows the shortest path, i.e.
+the geodesic, avoiding any in-plane rotation.
+The geodesic interpolation is derived using singular value decomposition and is implemented in the `tourr` package.
 
-XXX need to decide how much math to put in, if/which picture to use (probably something from @Buja2005-cx)
+Typically the geodesic interpolation is preferred, since within-plane rotation is distracting and does not provide new information.
+An exception might be the case of a projection pursuit guided tour with an index function that is sensitive to these rotations [@Laa2020-zj].
+
+For any method we need to ensure that each projection along the interpolated path is still orthonormal.
+In the case of one-dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure (XXX adding an illustration using ferrn).
+For higher-dimensional projections this is generalised and we find the path along the surface of a torus instead (need to check about this with Sherry).
 

### 2021-03-16 paper.Rmd
Slice tour paragraphs and figure.

@@
 In the case of one-dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure (XXX adding an illustration using ferrn).
-For higher-dimensional projections this is generalised and we find the path along the surface of a torus instead (need to check about this with Sherry).
+For higher-dimensional projections this is generalized and we find the path along the surface of a torus instead (need to check about this with Sherry).
 
@@
 
-Projections can hide certain features, sections can provide additional information @Furnas1994-dz
-Sectioning a high-dimensional space in systematic manner is challenging, there are a lot of choices we need to make
-Typical approach is interactive: brushing to select sections in one view, highlight points in a second view, e.g. a tour display
-Slice tour approach: use distance from the projection plane (through a selected center point), update the section systematically with the projection plane @Laa2020-js
-Summary of applications: viewing geometric shapes, model visualization
-Connection with projection pursuit: section pursuit @Laa2020-ni
+The classic tour displays work with projections, and interactive systems like GGobi also provide the option of adding sectioning information via linked brushing.
+Sectioning can be considered as providing complementary information, and in combination with projections we can for example gain insights into the dimensionality of a data structure [@Furnas1994-dz].
+Sectioning a high-dimensional space in systematic manner (without interactive selection of data sections) is challenging, since there is a lot of freedom in choosing a section.
+A new approach is implemented in the slice tour display [@Laa2020-js], which is using the distance of each from the current projection plane (typically placed such that it passes through the data mean) to highlight points that are nearby and fade out points further from the plane.
 
+The slice tour can reveal concave or non-linear structures obscured in projections, as well as small structures hidden near the center of a distribution.
+An example showing snapshots of the slice tour of points distributed on the surface of geometric shapes is shown in Fig.
+\ref{fig:tour-slice}.
+As with projections, an index function may be used to select target planes showing interesting slices of the data, and we can define a section pursuit guided tour [@Laa2020-ni].
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.2\textwidth]{figures/roman-surface-centered-010.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/torus-4-centered-015.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/cube-6-centered-001.png}}
+\centerline{
+\includegraphics[width=0.2\textwidth]{figures/roman-surface-centered-032.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/torus-4-centered-033.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/cube-6-centered-063.png}
+}
+\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right).}
+\label{fig:tour-slice}
+\end{figure*}
+
+<!--
 \begin{figure*}
@@
 \end{figure*}
+-->
 

### 2021-03-16 paper.Rmd
Paragraph, figure on sage display.

@@
 
-Transformations of the data are common in the preprocessing, but sometimes we want to transform the data after projecting onto lower dimensions
-Example would be to address piling effects that are a result of projecting form high-dimensional space down to low-dimensions
-A solution to this problem was presented in the sage display @Laa2020-uv and is using a radial transformation of the projected data to avoid piling near the center
-Summary of applications: better understanding of clustering in high-dimensional space (discussed below), resolving small features near the center
+Transformations of the data are common in the preprocessing, examples would be scaling, sphering, or a logarithmic transformation for skewed distributions.
+However, we may also want to transform the data after projecting onto lower dimensions, to correct for unwanted effects of the projections.
+One example is the piling effect that is observed when projecting a high-dimensional distribution [@Diaconis1984-mv]: projected points are approximately Gaussian in most views, and increasingly concentrating near the center as dimensionality increases.
+The sage display [@Laa2020-uv] proposes a solution to this issue, via a radial transformation of the projected data.
+This non-linear transformation is defined in each point of the projection plane and is sensitive to the overall scale of the data as well as the original dimensionality $p$.
+In addition, tuning parameters can be used to obtain a more (or less) aggressive redistribution of the points.
+This is illustrated in Fig.
+\ref{fig:pollen} showing the application of the sage transformation to the classical pollen dataset [@Coleman1986-so].
+The projection without any rescaling looks similar to what is found when rescaling with default options (left), and we can use either of the two tuning parameters to better resolve the distribution near the center and reveal the hidden structure (middle and right).
 
+\begin{figure}
+\centering
+\includegraphics[width=0.3\textwidth]{figures/pollen_sage-088.png}
+\includegraphics[width=0.3\textwidth]{figures/pollen_sage_R1-088.png}
+\includegraphics[width=0.3\textwidth]{figures/pollen_sage_gam20-088.png}
+\caption{Snapshots of the same projection from the sage tour on the pollen data, illustrating the different tuning options.
+Left: using the default settins for the transformation results in a similar view as found in projections without the sage transformation.
+Middle: using the tuning parameter $R$ for better resolution near the center reveals the word "EUREKA".
+Right: using the $\gamma$ tuning parameter can also reveal the structure.}
+\label{fig:pollen}
+\end{figure}
+
+<!--
 \begin{figure*}

### 2021-03-16 paper.Rmd
Spin-and-brush paragraph added.

@@
 
-Tour for visual cluster analysis, identification of outlying points
-Maybe the physics example from the ggobi book: find group in one view, highlight, restart tour, and so on, in the end we can review the tour with groups highlighted in colour to get an overview of how the data clusters
-Requires persistent brushing with multiple groups
+Tasks such as the identification of outlying points or cluster analysis can be performed visually using the tour.
+Here, a "spin-and-brush" approach works well, since different views will reveal different features and a persistent brush helps us to connect this to the previously observed information.
+As an example we briefly summarize an analysis from @Cook2007-be that explores clustering of the physics data previously used to in the context of projection pursuit [@Friedman1974-ck; @Cook1995-ae].
+This data is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
+These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Fig.
+\ref{prim7-tour}.
+We stop the tour each time a cluster is clearly separate from the main distribution in the current projection and brush the points in a new color.
+In the end the full structure becomes apparent and can be visualized by replaying the tour with all clusters highlighted in different colors.
 
@@
   {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp13.pdf}}}
-\caption{Spin and brush reprinted from ggobi book}
+\caption{Stepwise identification of clusters in the physics data.
+Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.}
 \label{prim7-tour}

### 2021-03-16 paper.Rmd
cleaning up a little

@@
 
-The focus of this review will be on visualising high-dimensional data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
-The reason being that it is not feasible to adequately the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.
+The focus of this review will be on visualising high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+The reason being that it is not feasible to adequately review the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.
 An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
@@
 
-<!-- Original tours (Di)
+<!-- above paragraphs could be shortened into a single paragraph? -->
+
+<!-- 
 Not sure how much of this wee need to go into...
@@
 > figure 2 diagram illustrating a projection: Di thinks making a Huber plot (2D data - 1D projection) here with projected data on a couple of directions is ideal
-
-# ```{r}
-# PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
-# ```
+```{r, eval = FALSE}
+PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
+```
 
@@
 
-- Stop/start/reverse (Stuart)
+Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualisation.
+Often one will want to pause on a particularly interesting projection and return it for use in a downstream analysis or static plot.
+Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown in figure \ref{fig:tour-controls}.
 
@@
 
-## Manual tour (Nick)
+## Manual tour
+
+<!-- needs to be shortend -->
 
@@
 To start we first identify a basis by orthonormalising the linear discriminant.
-We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, compeletely out of the projection, and then returned to it's original contribution.
+We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, completely out of the projection, and then returned to it's original contribution.
 
@@
 - tourr (Di)
+
 - spinifex (Nick)

### 2021-03-17 paper.Rmd
brushing section draft

@@
 
-- Linked brushing (Earo)
+## Linked Brushing
 
-> something from ggobi book
+Brushing is an interaction that results in a region being drawn onto the view of visualisation.
+When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a lookup to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
+The effect of brushing can be transient, that is, the changes triggered by a brush are removed after use, or they can be persistent i.e.
+the changes to data within the same view or adjacent views remain (Figure \ref{fig:prim7-tour}).
+Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or used to section lower dimensional views.
 
-* Brushing as database query
-* Brushing as conditioning
-* Brushing as geometric sectioning
-* Persistent vs. transient
+\begin{figure*}[!ht]
+   \centering
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush_query.pdf}
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush_response.pdf}
+\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by higlighting the corresponding points.}
+\label{fig:brushing}
+\end{figure*}
 
-> > figure 10: plots showing example
-
-- Spin-and-brush  (Ursula)
+## Spin-and-brush
 

### 2021-03-17 paper.Rmd
fix paths

@@
    \centering
-         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush_query.pdf}
-         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush_response.pdf}
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-query.pdf}
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-response.pdf}
 \caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by higlighting the corresponding points.}

### 2021-03-17 paper.Rmd
shuffling and add references
section headings

@@
 
-- Bases notation (Sherry)
+## Notation
 
@@
 
-- Display (Sherry)
+## Display
 
@@
 
-- Choosing targets, inc pp indexes (Sherry)
+## Finding targets
 
@@
 
-- Geodesic interpolation (Ursula)
+## Geodesic interpolation
 
@@
 
-- Slices vs projections (Ursula)
+## Slices, sections and projections
 
@@
 
-- Transformations on projections (Ursula)
+## Transformations on projections
 
@@
 
-<!-- needs to be shortend -->
+<!-- needs to be shortened -->
 
@@
 This data is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
-These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Fig.
-\ref{prim7-tour}.
+These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Figure \ref{prim7-tour}.
 We stop the tour each time a cluster is clearly separate from the main distribution in the current projection and brush the points in a new color.
@@
 
-- spinifex (Nick)
-
 The R package spinifex [@spyrison_spinifex_2020] has 3 primary features: it produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package, renders (manual or other) tours as animations exportable to static .gif (gganimate [@pedersen_gganimate_2020] package) or interactive .html widgets (plotly [@sievert_interactive_2020] package), lastly it offers interactive shiny application that offers an graphical user interface to quickly sample tour features.
 
-- liminal (Stuart)
-
-The liminal R package uses the tour to explore the quality of non-linear dimensionality reduction algorithms.
+The liminal R package uses the tour to explore the quality of non-linear dimensionality reduction algorithms [@Lee2020-kg].
 The interface consists of two side by side views consisting of a scatter plot displaying a reduced form of the data, and an interactive tour.
@@
 
-- python (Di to follow up with Csaba)
+- other work (Di)
 
@@
 
-- Model vis examples eg random forest (from Hadley, removing the blindfold) (Natalia)
+## Model Visualization
+
+- Model vis examples eg random forest (from Hadley, removing the blindfold) (Stuart)
 
@@
 
-- Physics (Ursula)
+The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg].
+They proposed a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
+They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
+For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance.
+Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
+@li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
+
+\begin{figure*}
+   \centering
+         \includegraphics[width=0.45\textwidth]{figures/scheidigger_nn_tour.png}
+          \includegraphics[width=0.45\textwidth, keepaspectratio=true]{figures/scheidigger_umap_tour.png}
+\caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}
+\label{fig:tour-nn}
+\end{figure*}
+
+## Physics
 
@@
          \includegraphics[width=0.45\textwidth]{figures/jetCluster-027.png}
-
 \caption{Selected frames from using a grand tour of the physics data.
@@
 
-- Bioinformatics (Stuart)
+## Bioinformatics
 
@@
 
+## Geometry of data
+
 - High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)
-
-> figure 15
-
-- neural networks (this could go into model vis?)
-
-The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg].
-They proposed a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
-They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
-For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance.
-Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
-@li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
-
-\begin{figure*}
-   \centering
-         \includegraphics[width=0.45\textwidth]{figures/scheidigger_nn_tour.png}
-          \includegraphics[width=0.45\textwidth, keepaspectratio=true]{figures/scheidigger_umap_tour.png}
-\caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}
-\label{fig:tour-nn}
-\end{figure*}
 

### 2021-03-17 paper.Rmd
add applications: model viz

@@
 
-Tasks such as the identification of outlying points or cluster analysis can be performed visually using the tour.
-Here, a "spin-and-brush" approach works well, since different views will reveal different features and a persistent brush helps us to connect this to the previously observed information.
-As an example we briefly summarize an analysis from @Cook2007-be that explores clustering of the physics data previously used to in the context of projection pursuit [@Friedman1974-ck; @Cook1995-ae].
-This data is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
-These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Figure \ref{prim7-tour}.
-We stop the tour each time a cluster is clearly separate from the main distribution in the current projection and brush the points in a new color.
-In the end the full structure becomes apparent and can be visualized by replaying the tour with all clusters highlighted in different colors.
-
 \begin{figure}[htp]
@@
 Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.}
-\label{prim7-tour}
+\label{fig:prim7-tour}
 \end{figure}
+
+Tasks such as the identification of outlying points or cluster analysis can be performed visually using the tour.
+Here, a "spin-and-brush" approach works well, since different views will reveal different features and a persistent brush helps us to connect this to the previously observed information.
+As an example we briefly summarize an analysis from @Cook2007-be that explores clustering of the physics data previously used to in the context of projection pursuit [@Friedman1974-ck; @Cook1995-ae].
+This data is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
+These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Figure \ref{fig:prim7-tour}.
+We stop the tour each time a cluster is clearly separate from the main distribution in the current projection and brush the points in a new color.
+In the end the full structure becomes apparent and can be visualized by replaying the tour with all clusters highlighted in different colors.
 
@@
 
-- Model vis examples eg random forest (from Hadley, removing the blindfold) (Stuart)
+\begin{figure*}
+  \centering
+    \includegraphics[width=0.32\textwidth]{figures/svm-radial-orange.png}
+    \includegraphics[width=0.32\textwidth]{figures/svm-radial-green.png}
+    \includegraphics[width=0.32\textwidth]{figures/svm-radial-purple.png}
+  \caption{A grand tour run on the prediction boundaries of a radial support vector machine.
+(Left) The prediction region for points classified as orange produces a slice of a cube.
+(Middle) The predicition region for points classified as green produces another slice of the cube.
+(Right) The complete prediction region shows the points classified as purple are embedded within the green slice.}
+  \label{fig:svm-tour}
+\end{figure*}
 
-> figure 12 tour plot(s) of vote matrix and tour plot(s) of variables, with some annotation of cases?
+@Wickham2015-do espouses the tour as a key component of their approach for visualizing models in data space.
+They propose using the tour to explore the high dimensional surfaces produced by model fits instead of simple summary statistics.
+In Figure \ref{fig:svm-tour} samples have been taken along the prediction region (i.e.
+the value the model predicts over a grid of observations) of a 3-$d$ support vector machine used to classify three different classes (green, orange purple), by touring over the prediction region we can see how the classifier splits the decision boundaries.
+
+\begin{figure*}
+   \centering
+         \includegraphics[width=0.9\textwidth]{figures/scheidigger_nn_tour.png}
+\caption{A grand tour run on the output of a deep neural network model for classification. The axes correspond to the probabilities that the model of an image belonging to a certain class.}
+\label{fig:tour-nn}
+\end{figure*}
 
@@
 They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
-For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance.
+For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance (Figure \ref{fig:tour-nn}).
 Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
 @li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
-
-\begin{figure*}
-   \centering
-         \includegraphics[width=0.45\textwidth]{figures/scheidigger_nn_tour.png}
-          \includegraphics[width=0.45\textwidth, keepaspectratio=true]{figures/scheidigger_umap_tour.png}
-\caption{Left: A grand tour run on the output of a deep neural network model for classification. Right: A grand tour on UMAP components of a layer of deep neural network.}
-\label{fig:tour-nn}
-\end{figure*}
 

### 2021-03-17 paper.Rmd
Clarification on sectioning.

@@
 The classic tour displays work with projections, and interactive systems like GGobi also provide the option of adding sectioning information via linked brushing.
-Sectioning can be considered as providing complementary information, and in combination with projections we can for example gain insights into the dimensionality of a data structure [@Furnas1994-dz].
+Sectioning means we select points that fall in a section of the full parameter space, for example for the purpose of highlighting them in a tour display.
+This often reveals complementary information, and combining sections and projections can for example provide insights into the dimensionality of a data structure [@Furnas1994-dz].
+
 Sectioning a high-dimensional space in systematic manner (without interactive selection of data sections) is challenging, since there is a lot of freedom in choosing a section.
-A new approach is implemented in the slice tour display [@Laa2020-js], which is using the distance of each from the current projection plane (typically placed such that it passes through the data mean) to highlight points that are nearby and fade out points further from the plane.
+One approach is to define sections based on projection planes, we refer to such sections as "slices" of the data.
+This is implemented in the slice tour display [@Laa2020-js], and uses the orthogonal distance of each data point from the current projection plane (typically placed such that it passes through the data mean) to highlight points that are nearby and fade out points further from the plane.
 

### 2021-03-18 paper.Rmd
title change and authors alphabetically except for Stuart

@@
     email: stuart.a.lee@monash.edu
+  - name: Dianne Cook
+    address: |
+      | Department of Econometrics and Business Statistics, Monash University
+    affiliation: Monash University
+    email: dicook@monash.edu
+  - name: Natalia Da Silva
+    address: |
+      | Instituto de Estadística (IESTA), Universidad de la República
+    affiliation: Universidad de la República
+    email: natalia@iesta.edu.uy
   - name: Ursula Laa 
@@
     email:  ursula.laa@boku.ac.at
-  - name: Natalia da Silva
-    address: |
-      | Instituto de Estadística (IESTA), Universidad de la República
-    affiliation: Universidad de la República
-    email: natalia@iesta.edu.uy
   - name: Nicholas Spyrison
@@
     email: nicholas.spyrison@monash.edu
-  - name: Dianne Cook
-    address: |
-      | Department of Econometrics and Business Statistics, Monash University
-    affiliation: Monash University
-    email: dicook@monash.edu
   - name: Earo Wang
@@
 title:
-  formatted: "An Overview of Tours for Dynamic Visualisation of High-dimensional Data"
+  formatted: "A Review of the State-of-the-Art on Tours for Dynamic Visualisation of High-dimensional Data"
 abstract: > 

### 2021-03-18 paper.Rmd
starting to flesh out history

@@
 
-> Maybe we should clarify What is high-dimensional data? here
-
 The term "high-dimensional" here means Euclidean space.
@@
 \caption{Illustrating what is meant by "high-dimensional" in this paper, and a "linear projection".
-From a sequence of increasing dimension cubes, from 1D to 6D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles}
+From a sequence of increasing dimension cubes, from 1D to 6D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles.
+(Note to Di: to change this diagram to highlight the doubling.)}
 \label{fig:cubes}
@@
 
+There has been a long history of software including tour methods, or something similar.
+These have laid the foundation for the current tools.
+The American Statistical Association Statistical Graphics Video Library web site [@ASASGVL] hosts videos documenting the history.
+It is fabulous watching the videos titled "Multidimensional Scaling" (Kruskal, 1962), "Real-Time Rotation" (Chang, 1970) and "Prim-9" (Tukey, 1973) show some preliminary methods leading to the development of the grand tour.
+The video titled "Use of the Grand Tour in Remote Sensing" (McDonald and Willis, 1987) is the first to show a tour, and the video titled "Dataviewer: A Program for Looking at Data in Several Dimensions" (Buja and Tukey, 1987) demonstrates the tour as part of a larger data analysis system.
+
 - Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)

### 2021-03-18 paper.Rmd
more history and references

@@
 The video titled "Use of the Grand Tour in Remote Sensing" (McDonald and Willis, 1987) is the first to show a tour, and the video titled "Dataviewer: A Program for Looking at Data in Several Dimensions" (Buja and Tukey, 1987) demonstrates the tour as part of a larger data analysis system.
+The videos titled "XGobi: Dynamic Graphics for Data Analysis" (Swayne et al, 1991) and "Grand Tour and Projection Pursuit" (Cook et al, 1993) show the tour tools in the XGobi software system.
+Some work on tours, for example @Wegman92 and @Tierney1991 is not documented in the video collection.
 
-- Dataviewer, XGobi, GGobi, cranvas (XLispStat, Orca, mmvz) (Di)
+Recently, some other review papers have summarized different aspects of tour methods.
+@Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve.
+@Moustafa2010 describes a computational shortcut for a tour.
+@Martinez2013 explains the image grand tour.
 
-- tourr (Di)
+The most accessible current software is the R package called tourr [@Wickham2011-uz].
+This software evolved from the Dataviewer [@Buja1986-la], XGobi [@Swayne1998], GGobi [@Swayne2003], Orca [@Sutherland2000] and cranvas [@Xie2014] ancestry.
 

### 2021-03-18 paper.Rmd
history done enough
torii example added

@@
 This software evolved from the Dataviewer [@Buja1986-la], XGobi [@Swayne1998], GGobi [@Swayne2003], Orca [@Sutherland2000] and cranvas [@Xie2014] ancestry.
+The tourr package has a wide range of display types, for different projection dimensions, and a selection of target generating methods including grand, guided, little, local, section, sage and frozen.
 
@@
 
-- High-d shapes/distributions: normal, skewed, outliers, (Di from tutorial notes)
+High-dimensional data is on many analysts' minds in recent years, and a way to build understanding of high-dimensional spaces is to examine common high-dimensional shapes using a tour.
+Many different shapes can be simulated using the R package geozoo.
+Figure \ref{fig:geozoo} shows points on the surface of two different types of high-dimensional torii.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=0.24\textwidth]{figures/torus-rings-4-with-torus-1.pdf}\includegraphics[width=0.24\textwidth]{figures/torus-rings-4-with-torus-2.pdf}
+  \includegraphics[width=0.24\textwidth]{figures/torus-flat-6-1.pdf}\includegraphics[width=0.24\textwidth]{figures/torus-flat-6-2.pdf}
+\caption{Several views of torii: (two at left) 4D ring torus, and (two at right) 6D flat torus. In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes.}
+\label{fig:geozoo}
+\end{figure*}
 

### 2021-03-18 paper.Rmd
adding dot points for discussion

@@
 
-* why tours are useful and what do we learn from them?
-* what research is needed?
+* summarising the paper again
+* highlighting the recent developments, spinifex, liminal, sage, slicing and sectioning, ferrn etc. 
+* what's next - conditional model visualisation (Catherine Hurley's work), a interactive graphics system that easily integrates tours that provides the direct manipulations we want, pre-processing / transformations for high dimensional data, penalised projection pursuit, different kinds of data and spaces
 

### 2021-03-18 paper.Rmd
huber plot

@@
 
-Some notations are defined for describing $d$-dimensional projections in $p$-dimensional space.
-Let $X_{n \times p}$ be the original data whose projections are of our interest.
-A projection basis, $A_{p \times d}$, characterises the direction from which the high-dimensional data are projected.
-Projection bases needs to satisfy an orthonormality condition, which requires each axis to be unit vector and perpendicular to each other, and this gives the basis space a geometrical structure of a Stiefel manifold.
-With a data matrix and a projection basis, a projection can be defined as $Y = X \cdot A$ for visual inspection.
+Below introduces the notations used for describing $d$-dimensional projections in a $p$-dimensional space.
+Let $X_{n \times p}$ be the data whose projections are of interest.
+A projection basis, $A_{p \times d}$, characterises the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires each column to be a unit vector and to be perpendicular to each other.
+This condition gives the $p$-dimensional basis space a geometrical structure of a Stiefel manifold.
+With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
+Figure \ref{fig:notation-huber} shows an example of the 2D iris data (sepal measures only) in a huber's plot, along with a 1D projection in the histogram.
+This projection is created based on projecting the data onto the solid line in the huber plot and it separates group setosa from versicolor and virginica.
+(This plot is created with package PPtreeViz, whose reference is @Lee2005-gl)
 
-> figure 2 diagram illustrating a projection: Di thinks making a Huber plot (2D data - 1D projection) here with projected data on a couple of directions is ideal
-```{r, eval = FALSE}
-PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
-```
+\begin{figure*}
+   \centering
+         \includegraphics[width=\textwidth]{figures/notation-huber.pdf}
+\caption{Huber's plot (left) of the 2D iris data (sepal length and width only).
+The solid ellipse represents how well the projection, onto different directions, separates the three species and the dashed circle is a reference guide.
+The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.}
+\label{fig:notation-huber}
+\end{figure*}
 

### 2021-03-18 R/tour-notation.R

@@ -1,0 +1,7 @@
+# Figure 2 diagram illustrating a projection: Huber plot of 2D data with 1D projection
+library(ggplot2)
+p <- PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA") 
+
+ggsave(p, filename = "notation-huber.pdf", 
+       path = here::here("figures"),
+       width = 12, height = 5, units = "in")

### 2021-03-18 paper.Rmd
a plot for various tour display

@@
 Various displays are available to show projections in 1D, 2D, or higher dimensions.
-A 1D projection shows the data in a histogram as if a hand shadow being projected onto a wall.
-Two-d projections are usually displayed with scatterplot since it is the most common plot for showing bivariate relationship.
+A 1D projection displays the data in a histogram as if a hand shadow being projected onto a wall.
+Two-d projections are usually shown as scatterplot since it is the most common display for bivariate relationship.
 Higher dimensional projections can be shown with multivariate display like scatterplot matrix or parallel coordinates plot.
-Figure <!--figure 3 --> shows a 1D display with histogram, a 2D display with scatterplot, and a 5D display with a parallel coordinates plot.
 More display methods can be found in [@Wickham2011-uz].
+Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables.
 
-grand tour, random, may miss interesting, take long to rotate to that specific angle, guided tour, along with projection pursuit, by generating target bases, interestingness
-
-> figure 3 diagram illustrating different displays for different d
+\begin{figure}
+\centering
+\includegraphics[width=0.3\textwidth]{figures/notation-display-1d.pdf}
+\includegraphics[width=0.3\textwidth]{figures/notation-display-2d.pdf}
+\includegraphics[width=0.3\textwidth]{figures/notation-display-5d.pdf}
+\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5d display in parallel coordinate plot (right).}
+\label{fig:display}
+\end{figure}
 

### 2021-03-18 R/tour-notation.R

@@ -5,3 +5,26 @@ p <- PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
 ggsave(p, filename = "notation-huber.pdf", 
        path = here::here("figures"),
        width = 12, height = 5, units = "in")
+
+# figure 3 diagram illustrating different displays for different d
+library(tourr)
+render(flea[,1:6], 
+       guided_tour(holes(), d= 1),
+       display_dist(),
+       "pdf", 
+       frames = 1,
+       file.path(here::here("figures", "notation-display-1d.pdf")))
+
+render(flea[,1:6], 
+       guided_tour(holes(), d= 2),
+       display_xy(),
+       "pdf", 
+       frames = 1,
+       file.path(here::here("figures", "notation-display-2d.pdf")))
+
+render(flea[,1:6], 
+       guided_tour(holes(), d= 5),
+       display_pcp(),
+       "pdf", 
+       frames = 1,
+       file.path(here::here("figures", "notation-display-5d.pdf")))

### 2021-03-18 paper.Rmd
all figures done

@@
 
-> figure 4 diagram illustrating different types of target generation
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{figures/notation-target-gen.pdf}
+\caption{A comparison of guided and grand tour path in the PCA reduced space.
+The start point of tour paths is smaller and ligher than the end point.
+Guided tour optimises the index value iteratively and finishes the search quickly while grand tour wanders in the parameter space to look at possible interesting projections.}
+\label{fig:notation-target-gen}
+\end{figure}
 

### 2021-03-18 R/tour-notation.R

@@ -28,3 +28,45 @@ render(flea[,1:6],
        "pdf", 
        frames = 1,
        file.path(here::here("figures", "notation-display-5d.pdf")))
+
+# figure 4 diagram illustrating different types of target generation
+library(tourr)
+library(ferrn)
+library(dplyr)
+n_basis <- 20
+set.seed(1234)
+grand <- save_history(
+  boa5, 
+  grand_tour(d = 1),
+  max_bases = n_basis
+)
+
+interp <- interpolate(grand)
+
+attr(interp, "class") <- NULL
+attr(interp, "data") <- NULL
+grand_list <- list()
+for (i in 1: dim(interp)[3]) grand_list[[i]] <- interp[,,i]
+index <- holes()
+grand_tibble <- purrr::map_dfr(grand_list, 
+                               ~tibble::tibble(basis = list(matrix(.x, ncol = 1)))) %>% 
+  mutate(method = "grand_tour", 
+         id = row_number(),
+         info = "interpolation",
+         index_val = vapply(basis, function(x) index(x), double(1)))
+
+dt <- bind_rows(grand_tibble, holes_1d_better %>% mutate(method = "guided_tour")) %>% 
+  compute_pca(group = method, flip = FALSE) %>% purrr::pluck("aug")
+
+p <- ggplot() + 
+  add_space(get_space_param(dt)) + 
+  add_start(get_start(dt), start_color = method, start_alpha = 0.8) + 
+  add_end(get_best(dt, group = method), end_color = method, end_size = 9) + 
+  add_interp(get_interp(dt), interp_color = method) + 
+  ggplot2::theme_void() +
+  ggplot2::theme(aspect.ratio = 1, legend.position = "bottom", legend.title = ggplot2::element_blank()) + 
+  scale_color_discrete_botanical()
+
+ggsave(p, filename = "notation-target-gen.pdf", 
+       path = here::here("figures"),
+       width = 5, height = 5, units = "in")

### 2021-03-18 paper.Rmd
notes to urusla

@@
 For higher-dimensional projections this is generalized and we find the path along the surface of a torus instead (need to check about this with Sherry).
+
+Notes to Ursula:
+
+- The basis space constrained by the orthonormality condition forms a Stiefel manifold, which is a sphere in 1D case and a torus in 2D case (I have one sentence for that in the notation section).
+- The space plot with tour path embedded are `figure/notation-1d-space.pdf` and `figure/notation-2d-space.pdf` - let me know if this is what you want
 

### 2021-03-18 paper.Rmd
Updated interpolation part based on notes from Sherry

--- 2021-03-18 paper.Rmd
@@
 For any method we need to ensure that each projection along the interpolated path is still orthonormal.
-In the case of one-dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure (XXX adding an illustration using ferrn).
-For higher-dimensional projections this is generalized and we find the path along the surface of a torus instead (need to check about this with Sherry).
+As explained above this means a tour path is a trace along a Stiefel manifold.
+In the case of one-dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure \ref{fig:notation-target-gen}, and for two-dimensional projections projections fall onto a torus shape.
+We can use a tour to visualize such a path, as shown in Figure \ref{fig:interpolation-2d}.
 
-> figure 5 diagram illustrating geodesic interpolation
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/notation-2d-space.pdf}
+\caption{Still views of a tour showing the space of possible two-dimensional projections as a torus shape in grey, and the trace of two short paths obtained via geodesic interpolation traced in color.}
+\label{fig:interpolation-2d}
+\end{figure}
 

### 2021-03-20 paper.Rmd
compress intro paragraph and reorder

@@
 
+While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualising structure in high dimensional data.
+An early technique proposed for assessing pairwise relationships between variables is the scatter plot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
+The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables.
+Similarly, parallel coordinates plots (PCP) can be used to explore correlation and multicollinearity [@Inselberg1985-kf; @Wegman1990-dy].
+By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings is revealed, however the ordering of variables along the axis changes what can be learned.
+Another display that relies on variable ordering is the heatmap, which is widely used to visualise cluster structure in bioinformatics.
+@Wilkinson2009-pj provides a comprehensive history of this display in the social and natural sciences.
+
 A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
@@
 \end{figure*}
-
-While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualising structure in high dimensional data.
-An early technique proposed for assessing pairwise relationships between variables is the scatter plot matrix (SPLOM) [@Tukey1983-fj] <!-- there might be others here(?) -->.
-In this graphic, the canvas is split into a matrix where each off-diagonal element is a scatter plot of variable pairs, while the diagonal elements are either left blank or represent univariate structure.
-The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables.
-
-Similarly, parallel coordinates plots (PCP) can be used to explore correlation and co-linearities in multivariate data [@Inselberg1985-kf].
-In PCPs, variables are arranged along the (vertical) horizontal axis with vertical (horizontal) axes drawn representing the range of values for each variable.
-Points (observations) of the data are represented by drawing lines between each variables axis.
-If there is positive correlation between two variables the lines will be parallel, while for negative correlation the lines will cross.
-By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings is revealed, however the ordering of variables is important to reveal these features.
-
-The heatmap display is widely used to visualise cluster structure in bioinformatics.
-In a heatmap each element of a numeric matrix is displayed as a single colored square pixel on the canvas.
-The hue of the color is mapped to the magnitude of the element's values.
-A subtlety of using the heatmap is determining how to order the rows and columns of the matrix called seriation.
-In practice, two way hierarchical clustering  or singular value decomposition is used to order rows and columns.
-See @Wilkinson2009-pj for a comprehensive history.
-
-<!-- above paragraphs could be shortened into a single paragraph? -->
-
-<!-- 
-Not sure how much of this wee need to go into...
-- subplots trellis literature
-- data pipelines (?)
-- What do we mean by multivariate data visualisation?
-- multivariate categorical vis, tree maps etc.
--->
 

### 2021-04-04 paper.Rmd
proofreading and commenting the first draft #3

@@
 The t-SNE view (plot A) shows six clearly separated clusters, all spherical with different sizes.
-The tour shows that the clusters look very little like this in the full data space.
+The tour shows that the clusters look very little like this [EW: not sure what "like this" referes to] in the full data space.
 The clusters are at various distances apart and very different sizes.
@@
 
-The rest of the review is structured as follows: in section \@ref(notation) we define the notation and components of a tour display.
+The rest of the review is structured as follows: in section \@ref(notation) we [EW: all of a sudden, from third person to first person?] define the notation and components of a tour display.
 By its nature a tour is most effective to analyst when combined with interactivity; in section \@ref(interaction) we review the components of user interfaces for manipulating tour views.
@@
 Let $X_{n \times p}$ be the data whose projections are of interest.
-A projection basis, $A_{p \times d}$, characterises the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires each column to be a unit vector and to be perpendicular to each other.
+A projection basis, $A_{p \times d}$, characterizes the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires each column to be a unit vector and to be perpendicular to each other.
 This condition gives the $p$-dimensional basis space a geometrical structure of a Stiefel manifold.
@@
 Figure \ref{fig:notation-huber} shows an example of the 2D iris data (sepal measures only) in a huber's plot, along with a 1D projection in the histogram.
-This projection is created based on projecting the data onto the solid line in the huber plot and it separates group setosa from versicolor and virginica.
+This projection is created based on projecting the data onto the solid line in the huber's plot and it separates group setosa from versicolor and virginica.
 (This plot is created with package PPtreeViz, whose reference is @Lee2005-gl)
@@
 \caption{Huber's plot (left) of the 2D iris data (sepal length and width only).
-The solid ellipse represents how well the projection, onto different directions, separates the three species and the dashed circle is a reference guide.
+The solid ellipse represents how well the projection, onto different directions, separating the three species and the dashed circle is a reference guide.
 The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.}
@@
 
-To define whether a projection is interesting or not in the guided tour, an index function is used.
+To define whether a projection is interesting or not in the guided tour [EW: should we add a new section on tour types before this section?], an index function is used.
 An index function defines a surjection function from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
@@
 Numerous index functions have been developed since the initial index proposed by @Friedman1974-ck: <!-- entropy-based index?
---> holes and central mass index [@Cook2008-nl] computes the deviation of the projected data from a normal distribution; @Posse1995-ih proposed a chi-square index for detecting the spiral structure in the data;  kurtosis and skewness indices are raised in @Loperfido2020-xj and @Loperfido2018-ga to detect outliers in financial time series; and most recently, scagnostic indices [@Laa2020-zj] provides eight measures for summarising information between pairs of variables in scatter plot matrices.
+--> holes and central mass index [@Cook2008-nl] computes the deviation of the projected data from a normal distribution; @Posse1995-ih proposed a chi-square index for detecting the spiral structure in the data;  kurtosis and skewness indices are raised in @Loperfido2020-xj and @Loperfido2018-ga to detect outliers in financial time series; and most recently, scagnostic indices [@Laa2020-zj] provide eight measures for summarizing information between pairs of variables in SPLOM.
 
@@
 \caption{A comparison of guided and grand tour path in the PCA reduced space.
-The start point of tour paths is smaller and ligher than the end point.
+The start point of tour paths is smaller and lighter than the end point.
 Guided tour optimises the index value iteratively and finishes the search quickly while grand tour wanders in the parameter space to look at possible interesting projections.}
@@
 
-The smooth animation between the selected target frame is generated via a step-wise rotation in the high-dimensional space.
-Different methods to derive this rotation were described in @Buja2005-cx, and we can distinguish two types of interpolation.
+The smooth animation between the selected target frames is generated via a step-wise rotation in the high-dimensional space.
+Different methods to derive this rotation were described in @Buja2005-cx, and we can distinguish two types of interpolation [EW: ??].
 When interpolating between frames we rotate the starting basis onto the exact basis representation selected as a basis, and @Buja2005-cx describes three different methods to achieve this (based on decompositions of orthogonal matrices, Givens decompositions, or Householder decompositions).
@@
 The geodesic interpolation is derived using singular value decomposition and is implemented in the `tourr` package.
+[EW: this paragraph perhaps needs some rework.]
 
@@
 
-The classic tour displays work with projections, and interactive systems like GGobi also provide the option of adding sectioning information via linked brushing.
+Interactive systems like GGobi provide the option of adding sectioning information via linked brushing.
 Sectioning means we select points that fall in a section of the full parameter space, for example for the purpose of highlighting them in a tour display.
@@
 The slice tour can reveal concave or non-linear structures obscured in projections, as well as small structures hidden near the center of a distribution.
-An example showing snapshots of the slice tour of points distributed on the surface of geometric shapes is shown in Fig.
-\ref{fig:tour-slice}.
-As with projections, an index function may be used to select target planes showing interesting slices of the data, and we can define a section pursuit guided tour [@Laa2020-ni].
+Fig.
+\ref{fig:tour-slice} shows snapshots of the slice tour of points distributed on the surface of geometric shapes.
+Along with projections, an index function can be used to select target planes to show interesting slices of the data, defined as a section pursuit guided tour [@Laa2020-ni].
 
@@
 
-Transformations of the data are common in the preprocessing, examples would be scaling, sphering, or a logarithmic transformation for skewed distributions.
+Transformations of the data are common in the preprocessing [EW: "preprocessing" means b/f projecting?], such as scaling, sphering, or a logarithmic transformation for skewed distributions.
 However, we may also want to transform the data after projecting onto lower dimensions, to correct for unwanted effects of the projections.
@@
 \caption{Snapshots of the same projection from the sage tour on the pollen data, illustrating the different tuning options.
-Left: using the default settins for the transformation results in a similar view as found in projections without the sage transformation.
+Left: using the default settings for the transformation results in a similar view as found in projections without the sage transformation.
 Middle: using the tuning parameter $R$ for better resolution near the center reveals the word "EUREKA".
@@
 This is particularly useful for exploring a projection once a feature of interest has been identified.
-Manual tours can then be employed to explore local area of, or test the structure of the feature, with respect to a selected variable.
+Manual tours can then be employed to explore local area of interest, or test the structure of the feature, with respect to a selected variable.
 
@@
          \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl4_r.pdf}
-\caption{Four projections from a manual tour where variable "bl" is being rotated out of the projection (orange line).
+\caption{Three projections from a manual tour where variable "bl" is being rotated out of the projection (orange line).
 When this variable is removed the two light green clusters merge, which informs us that "bl" is an important variable for distinguishing between these two groups.}
@@
 ## Linked Brushing
+
+[EW: seems unrelated to the rest of the article; alternatively briefly explain "brushing" in s2.5?]
 
@@
          \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-response.pdf}
-\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by higlighting the corresponding points.}
+\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by highlighting the corresponding points.}
 \label{fig:brushing}

### 2021-04-05 paper.Rmd
Reduce manual tour section, remove spinifex manual tour figure.

--- 2021-03-20 paper.Rmd
@@
 
-<!-- needs to be shortened -->
-
 Manual tours [@cook_manual_1997; @spyrison_spinifex_2020] offer a means to interactively control the contribution of a single variable on the projection plane.
 This is particularly useful for exploring a projection once a feature of interest has been identified.
-Manual tours can then be employed to explore local area of, or test the structure of the feature, with respect to a selected variable.
-
-The details of performing a manual tour are described by @spyrison_spinifex_2020, but high-level unit-interaction between each frame is as follows:
-
-__Input:__ Starting basis, manipulation variable, an in-plane rotation angle -- $\theta$, an out-of-plane rotation angle -- $\phi$  $\\$
-__Output:__ Rotated basis $\\$
-__Steps:__
-
-1. Initialise a manipulation dimension onto the start basis. Given the manipulation variable a full contribution and orthonormalise this space. 
-2. Post-multiply the space, by a rotation matrix, as solved for in application of Rodrigues’ formula [@rodrigues_lois_1840].
-3. Truncate the manipulation dimension.
-
-Figure \ref{fig:manual-tour} illustrates the manual tour, by radially manipulating the contribution of variable that is crucial, to explain the separation of a particular cluster.
-To start we first identify a basis by orthonormalising the linear discriminant.
-We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, completely out of the projection, and then returned to it's original contribution.
-
-```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
-We initalise the start basis to the linear discriminant analysis and select the variable with the largest contribution separating the cluster shown as green circles.
-By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.', eval=FALSE}
-if(F) ## Creation and saving figure at:
-  file.edit("./R/make_radial_tour_fig.r")
-knitr::include_graphics("./figures/fig_radial_manual_tour.png")
-```
+Manual tours can then be employed to test the structure of the feature, with respect to a selected variable.
+For instance, figure \ref{fig:manual-tour}, starts from the orthonormal linear discriminant and explores the sensitivity of class separation as the contribution of a single variable is altered.
 

### 2021-04-06 paper.Rmd

--- 2021-04-04 paper.Rmd
@@
 
-## Manual tour
-
-<!-- needs to be shortened -->
+# Manual Tours
 
@@
 This is particularly useful for exploring a projection once a feature of interest has been identified.
-Manual tours can then be employed to explore local area of interest, or test the structure of the feature, with respect to a selected variable.
-
-The details of performing a manual tour are described by @spyrison_spinifex_2020, but high-level unit-interaction between each frame is as follows:
-
-__Input:__ Starting basis, manipulation variable, an in-plane rotation angle -- $\theta$, an out-of-plane rotation angle -- $\phi$  $\\$
-__Output:__ Rotated basis $\\$
-__Steps:__
-
-1. Initialise a manipulation dimension onto the start basis. Given the manipulation variable a full contribution and orthonormalise this space. 
-2. Post-multiply the space, by a rotation matrix, as solved for in application of Rodrigues’ formula [@rodrigues_lois_1840].
-3. Truncate the manipulation dimension.
-
-Figure \ref{fig:manual-tour} illustrates the manual tour, by radially manipulating the contribution of variable that is crucial, to explain the separation of a particular cluster.
-To start we first identify a basis by orthonormalising the linear discriminant.
-We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, completely out of the projection, and then returned to it's original contribution.
-
-```{r manual-tour, out.width="100%", fig.show='asis', fig.cap='Manual tour of flea data.
-We initalise the start basis to the linear discriminant analysis and select the variable with the largest contribution separating the cluster shown as green circles.
-By controlling the contribution of the variable we see that it is quite sensitive to the sepration of this cluster.', eval=FALSE}
-if(F) ## Creation and saving figure at:
-  file.edit("./R/make_radial_tour_fig.r")
-knitr::include_graphics("./figures/fig_radial_manual_tour.png")
-```
+Manual tours can then be employed to test the structure of the feature, with respect to a selected variable.
+For instance, figure \ref{fig:manual-tour}, starts from the orthonormal linear discriminant and explores the sensitivity of class separation as the contribution of a single variable is altered.
 

### 2021-04-09 paper.Rmd
update todo
Updates interactivity section, corrects figure captions for Stuart's sections, clarifies final paragraph from introduction

@@
 
+All of the aforementioned techniques can be enhanced through the use of interactivity.
+By combining views with interactive elements like tooltips or highlighting the analysts is able to quickly interrogate interesting features of the data.
+One  particularly important interaction technique in the history of statistical graphics is called brushing [@Becker1987-dw].
+When brushing an analyst drags their mouse over the view which results in a region being drawn onto the canvas.
+When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a lookup to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
+This technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).
+
+\begin{figure*}[!ht]
+   \centering
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-query.pdf}
+         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-response.pdf}
+\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange.
+The view on the right responds by highlighting the corresponding points.
+(Adapted from Cook and Swayne, 2007 Figure 2.12)}
+\label{fig:brushing}
+\end{figure}
+
 A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
@@
 
-The rest of the review is structured as follows: in section \@ref(notation) we [EW: all of a sudden, from third person to first person?] define the notation and components of a tour display.
-By its nature a tour is most effective to analyst when combined with interactivity; in section \@ref(interaction) we review the components of user interfaces for manipulating tour views.
-The implementation of tour paths in statistical software is reviewed in section \@ref(software), where we discuss the history and state of the art packages for computing them.
+The rest of the review is structured as follows: section \@ref(notation) defines the notation and components of a tour displays.
+By its nature a tour is most effective to analyst when combined with interactivity; section \@ref(interaction) reviews the components of user interfaces for manipulating tour views.
+The implementation of tour paths in statistical software is reviewed in section \@ref(software).
 Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning and applied statistics.
-Finally, in section \@ref(discussion) we point to future research directions for tours.
+Finally, section \@ref(discussion) discusses future research directions for tours.
 
@@
 \end{figure*}
+
+<!-- Sherry to change to penguins data -->
 
@@
 \end{figure}
+
+<!-- Sherry to include two spheres, so the figure fills the whole page -->
 
@@
 The slice tour can reveal concave or non-linear structures obscured in projections, as well as small structures hidden near the center of a distribution.
-Fig.
-\ref{fig:tour-slice} shows snapshots of the slice tour of points distributed on the surface of geometric shapes.
+Figure \ref{fig:tour-slice} shows snapshots of the slice tour of points distributed on the surface of geometric shapes.
 Along with projections, an index function can be used to select target planes to show interesting slices of the data, defined as a section pursuit guided tour [@Laa2020-ni].
@@
 In addition, tuning parameters can be used to obtain a more (or less) aggressive redistribution of the points.
-This is illustrated in Fig.
-\ref{fig:pollen} showing the application of the sage transformation to the classical pollen dataset [@Coleman1986-so].
+This is illustrated in Figure \ref{fig:pollen} showing the application of the sage transformation to the classical pollen dataset [@Coleman1986-so].
 The projection without any rescaling looks similar to what is found when rescaling with default options (left), and we can use either of the two tuning parameters to better resolve the distribution near the center and reveal the hidden structure (middle and right).
@@
 
-<!--
-\begin{figure*}
-   \centering
-         \includegraphics[width=0.9\textwidth]{figures/laa_radial_transformation.pdf}
-\caption{An illustration of the radial transformation. Each panel represents linear projections of $p$ dimensional equidistant circles. As $p$ increases the circles get pushed further to the edge.}
-\label{fig:tour-transformation}
-\end{figure*}
--->
-
 # Ways of interacting {#interaction}
@@
 \caption{The clustering example from Figure \ref{fig:tsne-tour} using the liminal interface.
-Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.}
+Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.
+Adapted from Lee, Laa and Cook, 2020 Figure 2.}
 \label{fig:tour-controls}
@@
 
-## Linked Brushing
-
-[EW: seems unrelated to the rest of the article; alternatively briefly explain "brushing" in s2.5?]
-
-Brushing is an interaction that results in a region being drawn onto the view of visualisation.
-When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a lookup to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
-The effect of brushing can be transient, that is, the changes triggered by a brush are removed after use, or they can be persistent i.e.
-the changes to data within the same view or adjacent views remain (Figure \ref{fig:prim7-tour}).
-Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or used to section lower dimensional views.
-
-\begin{figure*}[!ht]
-   \centering
-         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-query.pdf}
-         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-response.pdf}
-\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by highlighting the corresponding points.}
-\label{fig:brushing}
-\end{figure*}
-
-## Spin-and-brush
+## Spin-and-brush {#sec:spin}
 
@@
 \caption{Stepwise identification of clusters in the physics data.
-Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.}
+Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.
+Adapted from Cook and Swayne, 2007 Figure 5.3.}
 \label{fig:prim7-tour}
@@
 
-Tasks such as the identification of outlying points or cluster analysis can be performed visually using the tour.
+Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or used to section lower dimensional views.
+This aids tasks such as the identification of outlying points or visual cluster analysis when combined with the tour.
 Here, a "spin-and-brush" approach works well, since different views will reveal different features and a persistent brush helps us to connect this to the previously observed information.
@@
 (Middle) The predicition region for points classified as green produces another slice of the cube.
-(Right) The complete prediction region shows the points classified as purple are embedded within the green slice.}
+(Right) The complete prediction region shows the points classified as purple are embedded within the green slice.
+Adapted from Wickham, Hoffman and Cook, 2015.}
   \label{fig:svm-tour}
@@
          \includegraphics[width=0.9\textwidth]{figures/scheidigger_nn_tour.png}
-\caption{A grand tour run on the output of a deep neural network model for classification. The axes correspond to the probabilities that the model of an image belonging to a certain class.}
+\caption{A grand tour run on the output of a deep neural network model for classification.
+The axes correspond to the probabilities that the model of an image belonging to a certain class.
+Adapted from Li, Zhao and Scheidigger, 2020.}
 \label{fig:tour-nn}
@@
 For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point.
-For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterisk symbol in Fig.
-\ref{fig:physics}.
+For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterisk symbol in Figure \ref{fig:physics}.
 
@@
 \caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation.
-The sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.}
+The sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.
+Adapted from Laa, Lee and Cook, 2020 Figure 6.}
 \label{fig:mouse-cluster}

### 2021-04-09 paper.Rmd
figure caption correction
rejig sentence

@@
 \label{fig:brushing}
-\end{figure}
+\end{figure*}
 
@@
 The t-SNE view (plot A) shows six clearly separated clusters, all spherical with different sizes.
-The tour shows that the clusters look very little like this [EW: not sure what "like this" referes to] in the full data space.
+The tour shows that the clusters do not look like this in the full data space.
 The clusters are at various distances apart and very different sizes.
@@
 \end{figure*}
-
-<!--
-\begin{figure*}
-   \centering
-         \includegraphics[width=0.9\textwidth]{figures/slice_tour_left.jpg}
-         \includegraphics[width=0.9\textwidth]{figures/slice_tour_right.jpg}
-\caption{Ursula to write}
-\label{fig:tour-slice}
-\end{figure*}
--->
 

### 2021-04-09 paper.Rmd
add discussion

@@
 
-* summarising the paper again
+This review highlights the history of the grand tour and its modern extensions for the visualization of high dimensional numerical data.
+A rich ecosystem of tour methods are being actively developed and are finding applications in a broad range of areas.
+Tours are being combined with interactive graphics to assist with cluster analysis (the spinifex and liminal packages), to understand higher dimensional space and find lower dimensional structure (the sage algorithm, slicing and sectioning) <!-- Ursula can you edit this?
+-->, and to diagnose optimization procedures for projection pursuit (ferrn).
+
+There are many possible future directions for tour research, one big limitation
+of the current software ecosystem is that analysts are no longer able to perform interactions that were possible in software packages from the 80s and 90s.
+An interactive graphics system that integrated the tour methodology directly would be a big step forward.
+<!-- Ursula to add sentence on Catherine Hurley's work and other aspects -->
+
 * highlighting the recent developments, spinifex, liminal, sage, slicing and sectioning, ferrn etc. 

### 2021-04-09 paper.Rmd
changes on the plot as per comments

@@
 With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows an example of the 2D iris data (sepal measures only) in a huber's plot, along with a 1D projection in the histogram.
-This projection is created based on projecting the data onto the solid line in the huber's plot and it separates group setosa from versicolor and virginica.
-(This plot is created with package PPtreeViz, whose reference is @Lee2005-gl)
+Figure \ref{fig:notation-huber} shows an example of the 2D palmerpenguins data (bill measures only) in a huber's plot, along with a 1D projection in the histogram.
+This projection is created based on projecting the data onto the solid line in the huber's plot and the it shows the separation among three species.
 
@@
          \includegraphics[width=\textwidth]{figures/notation-huber.pdf}
-\caption{Huber's plot (left) of the 2D iris data (sepal length and width only).
+\caption{Huber's plot (left) of the 2D penguins data (bill length nad depth only).
 The solid ellipse represents how well the projection, onto different directions, separating the three species and the dashed circle is a reference guide.
-The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.}
+The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.
+Adapted from E.-K.
+Lee et al.
+(2005), Figure 1.
+}
 \label{fig:notation-huber}
 \end{figure*}
-
-<!-- Sherry to change to penguins data -->
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/notation-target-gen.pdf}
+\includegraphics[width=\textwidth]{figures/notation-target-gen.pdf}
 \caption{A comparison of guided and grand tour path in the PCA reduced space.
@@
 \end{figure}
-
-<!-- Sherry to include two spheres, so the figure fills the whole page -->
 

### 2021-04-09 R/tour-notation.R

@@ -1,6 +1,9 @@
 # Figure 2 diagram illustrating a projection: Huber plot of 2D data with 1D projection
 library(ggplot2)
-p <- PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA") 
+library(palmerpenguins)
+library(dplyr)
+penguins_small <- penguins %>% filter(!is.na(bill_length_mm))
+p <- PPtreeViz::Huberplot(penguins_small[,3:4],penguins_small$species,PPmethod="LDA") 
 
 ggsave(p, filename = "notation-huber.pdf", 
        path = here::here("figures"),
@@ -65,8 +68,11 @@ p <- ggplot() +
   add_interp(get_interp(dt), interp_color = method) + 
   ggplot2::theme_void() +
   ggplot2::theme(aspect.ratio = 1, legend.position = "bottom", legend.title = ggplot2::element_blank()) + 
-  scale_color_discrete_botanical()
+  scale_color_discrete_botanical() + 
+  facet_wrap(vars(method)) + 
+  theme(strip.text.x = element_blank(),
+        legend.text = element_text(size = 20))
 
 ggsave(p, filename = "notation-target-gen.pdf", 
        path = here::here("figures"),
-       width = 5, height = 5, units = "in")
+       width = 10, height = 5, units = "in")

### 2021-04-09 paper.Rmd
rework finding targets section

@@
 
-To define whether a projection is interesting or not in the guided tour [EW: should we add a new section on tour types before this section?], an index function is used.
-An index function defines a surjection function from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
-For the high-dimensional data, an interesting projection is the one that deviates from normality, as discussed in @Diaconis1984-mv and @Huber1985-zq [expand more here].
-Numerous index functions have been developed since the initial index proposed by @Friedman1974-ck: <!-- entropy-based index?
---> holes and central mass index [@Cook2008-nl] computes the deviation of the projected data from a normal distribution; @Posse1995-ih proposed a chi-square index for detecting the spiral structure in the data;  kurtosis and skewness indices are raised in @Loperfido2020-xj and @Loperfido2018-ga to detect outliers in financial time series; and most recently, scagnostic indices [@Laa2020-zj] provide eight measures for summarizing information between pairs of variables in SPLOM.
+A sequence of projections can be generated via different tour paths, i.e.
+grand tour, guided tour, manual tour, little tour, local tour, and most recently, slice tour.
+The focus of this section will be on the grand and guided tour, with manual and slice tour being introduced in the later sections of the review.
+
+Grand tour, coined by @Asimov1985-xr, randomly samples target bases in the space and constructs an interpolation between them to produce a sequence of projections as in Figure  \ref{fig:notation-target-gen}.
+This tour path is suitable for an initial view of the data, however, as the dimension increases, it could take a long time before the grand tour finds a projection.
+
+Instead of randomly generating targets, guided tour [@Cook1995-ae] finds its targets with a technique called projection pursuit [@Friedman1974-ck].
+In projection pursuit, an index function is a surjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
+Guided tour iteratively optimises the index function in the projection pursuit and produces the most interesting projection according to the index measure.
+The right panel of Figure \ref{fig:notation-target-gen} shows an example of a guided tour path.
 
@@
 \includegraphics[width=\textwidth]{figures/notation-target-gen.pdf}
-\caption{A comparison of guided and grand tour path in the PCA reduced space.
+\caption{A comparison of guided and grand tour path in the PCA reduced 2D space.
 The start point of tour paths is smaller and lighter than the end point.

### 2021-04-09 R/tour-notation.R

@@ -37,10 +37,11 @@ library(tourr)
 library(ferrn)
 library(dplyr)
 n_basis <- 20
+start_basis <- holes_1d_better %>% filter(id == 1) %>% pull(basis) %>% .[[1]]
 set.seed(1234)
 grand <- save_history(
   boa5, 
-  grand_tour(d = 1),
+  tour_path = grand_tour(d = 1),
   max_bases = n_basis
 )
 
@@ -54,9 +55,12 @@ index <- holes()
 grand_tibble <- purrr::map_dfr(grand_list, 
                                ~tibble::tibble(basis = list(matrix(.x, ncol = 1)))) %>% 
   mutate(method = "grand_tour", 
-         id = row_number(),
+         id = row_number() + 1,
          info = "interpolation",
-         index_val = vapply(basis, function(x) index(x), double(1)))
+         index_val = vapply(basis, function(x) index(x), double(1))) %>% 
+  add_row(basis = list(start_basis), method = "grand_tour", id = 1, 
+          info = "interpolation", index_val = index(start_basis)) %>% 
+  arrange(id)
 
 dt <- bind_rows(grand_tibble, holes_1d_better %>% mutate(method = "guided_tour")) %>% 
   compute_pca(group = method, flip = FALSE) %>% purrr::pluck("aug")
@@ -76,3 +80,4 @@ p <- ggplot() +
 ggsave(p, filename = "notation-target-gen.pdf", 
        path = here::here("figures"),
        width = 10, height = 5, units = "in")
+

### 2021-04-09 paper.Rmd
References in captions added.

@@
 }
-\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right).}
+\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from Laa, Cook and Valencia (2020), Fig. 4.}
 \label{fig:tour-slice}
@@
 Middle: using the tuning parameter $R$ for better resolution near the center reveals the word "EUREKA".
-Right: using the $\gamma$ tuning parameter can also reveal the structure.}
+Right: using the $\gamma$ tuning parameter can also reveal the structure.
+Adapted from Laa, Cook and Lee, Fig.
+9.}
 \label{fig:pollen}
@@
 Left: the three different types of measurements (shown in different colors) are aligned along different directions in parameter space.
-Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.}
+Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.
+Adapted from Cook, Laa and Valencia, 2018, Figures 6 and 7.}
 \label{fig:physics}

### 2021-04-09 paper.Rmd
Small additions to the discussion.

@@
 A rich ecosystem of tour methods are being actively developed and are finding applications in a broad range of areas.
-Tours are being combined with interactive graphics to assist with cluster analysis (the spinifex and liminal packages), to understand higher dimensional space and find lower dimensional structure (the sage algorithm, slicing and sectioning) <!-- Ursula can you edit this?
--->, and to diagnose optimization procedures for projection pursuit (ferrn).
+Tours are being combined with interactive graphics to assist with cluster analysis (the spinifex and liminal packages), displays of projected data are augmented to allow the identification of local structures (the sage algorithm, slicing and section pursuit), and  new visualizations based on tour methods are used to better diagnose optimization procedures for projection pursuit (ferrn).
 
@@
 An interactive graphics system that integrated the tour methodology directly would be a big step forward.
-<!-- Ursula to add sentence on Catherine Hurley's work and other aspects -->
+Such an interface could also be combined with methods for conditional model visualization, for example as implemented in the condviz package [@JSSv081i05], allowing the visual exploration of multivariate fitted models.
 

### 2021-04-12 paper.Rmd
use American spelling

@@
 This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936].
-It is also common to reduce dimension nonlinearly with techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbour Embedding (t-SNE) [@Van_der_Maaten2008-qa].
+It is also common to reduce dimension with nonlinear techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbor Embedding (t-SNE) [@Van_der_Maaten2008-qa].
 
@@
 
-The focus of this review will be on visualising high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
-The reason being that it is not feasible to adequately review the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.
+The focus of this review will be on visualizing high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+The reason being that it is not feasible to adequately review the very large area of visualizing high-dimensions, and there have been numerous developments in tours recently.
 An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
 
-While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualising structure in high dimensional data.
+While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualizing structure in high dimensional data.
 An early technique proposed for assessing pairwise relationships between variables is the scatter plot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
 The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables.
-Similarly, parallel coordinates plots (PCP) can be used to explore correlation and multicollinearity [@Inselberg1985-kf; @Wegman1990-dy].
-By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings is revealed, however the ordering of variables along the axis changes what can be learned.
+Similarly, parallel coordinates plots (PCP) can be used to explore correlation and collinearity [@Inselberg1985-kf; @Wegman1990-dy].
+By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings are revealed, however the ordering of variables along the axis changes what can be learned.
 Another display that relies on variable ordering is the heatmap, which is widely used to visualise cluster structure in bioinformatics.

### 2021-04-12 paper.Rmd
use American spelling

@@
 By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings are revealed, however the ordering of variables along the axis changes what can be learned.
-Another display that relies on variable ordering is the heatmap, which is widely used to visualise cluster structure in bioinformatics.
+Another display that relies on variable ordering is the heatmap, which is widely used to visualize cluster structure in bioinformatics.
 @Wilkinson2009-pj provides a comprehensive history of this display in the social and natural sciences.
@@
 All of the aforementioned techniques can be enhanced through the use of interactivity.
-By combining views with interactive elements like tooltips or highlighting the analysts is able to quickly interrogate interesting features of the data.
+By combining views with interactive elements like tool tips or highlighting the analysts is able to quickly interrogate interesting features of the data.
 One  particularly important interaction technique in the history of statistical graphics is called brushing [@Becker1987-dw].
 When brushing an analyst drags their mouse over the view which results in a region being drawn onto the canvas.
-When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a lookup to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
+When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a look-up query to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
 This technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).
@@
 With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows an example of the 2D palmerpenguins data (bill measures only) in a huber's plot, along with a 1D projection in the histogram.
-This projection is created based on projecting the data onto the solid line in the huber's plot and the it shows the separation among three species.
+Figure \ref{fig:notation-huber} shows an example of the 2D palmerpenguins data (bill measures only) in a Huber plot, along with a 1D projection in the histogram.
+This projection is created based on projecting the data onto the solid line in the Huber plot and the it shows the separation among three species.
 
@@
 In projection pursuit, an index function is a surjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
-Guided tour iteratively optimises the index function in the projection pursuit and produces the most interesting projection according to the index measure.
+Guided tour iteratively optimizes the index function in the projection pursuit and produces the most interesting projection according to the index measure.
 The right panel of Figure \ref{fig:notation-target-gen} shows an example of a guided tour path.
@@
 For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point.
-For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the outlyingness of the data point marked with an asterisk symbol in Figure \ref{fig:physics}.
+For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the "outlyingness" of the data point marked with an asterisk symbol in Figure \ref{fig:physics}.
 

### 2021-04-12 paper.Rmd
wordsmithing

@@
 The reason being that it is not feasible to adequately review the very large area of visualizing high-dimensions, and there have been numerous developments in tours recently.
-An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.
+An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications.
+
+A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
+It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (see for e.g.
+@Swayne2003-cc or @Tierney1991).
+To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
+It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
 
 While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualizing structure in high dimensional data.
-An early technique proposed for assessing pairwise relationships between variables is the scatter plot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
+An early technique proposed for assessing pairwise relationships between variables is the scatterplot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
 The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables.
@@
 When brushing an analyst drags their mouse over the view which results in a region being drawn onto the canvas.
-When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used as a look-up query to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
+When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
 This technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).
@@
 \end{figure*}
-
-A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
-It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (see for e.g.
-@Swayne2003-cc or @Tierney1991).
-To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
-It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
 
@@
 
-Below introduces the notations used for describing $d$-dimensional projections in a $p$-dimensional space.
-Let $X_{n \times p}$ be the data whose projections are of interest.
-A projection basis, $A_{p \times d}$, characterizes the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires each column to be a unit vector and to be perpendicular to each other.
-This condition gives the $p$-dimensional basis space a geometrical structure of a Stiefel manifold.
-With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows an example of the 2D palmerpenguins data (bill measures only) in a Huber plot, along with a 1D projection in the histogram.
-This projection is created based on projecting the data onto the solid line in the Huber plot and the it shows the separation among three species.
+When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$.
+Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest.
+A projection basis, $A_{p \times d}$, is matrix that characterizes the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires that each column in $A$ has unit length and are perpendicular.
+<!-- This condition gives the $p$-dimensional basis space a geometrical structure of a Stiefel manifold.
+(Not sure this adds anything) --> With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
+Figure \ref{fig:notation-huber} shows an example of lower dimensional projections of the palmer penguins data in a Huber plot and a histogram [@penguins].
+
+<!-- SL: I think there needs to be more detail here to link to the next paragraph and more explanation about what the huber plot is showing -->
 
@@
          \includegraphics[width=\textwidth]{figures/notation-huber.pdf}
-\caption{Huber's plot (left) of the 2D penguins data (bill length nad depth only).
-The solid ellipse represents how well the projection, onto different directions, separating the three species and the dashed circle is a reference guide.
+\caption{Huber's plot (left) of the penguins data (bill length and depth only).
+The solid ellipse represents how well the projection, onto different directions, separates the three species and the dashed circle is a reference guide.
 The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.
@@
 
-Various displays are available to show projections in 1D, 2D, or higher dimensions.
-A 1D projection displays the data in a histogram as if a hand shadow being projected onto a wall.
-Two-d projections are usually shown as scatterplot since it is the most common display for bivariate relationship.
-Higher dimensional projections can be shown with multivariate display like scatterplot matrix or parallel coordinates plot.
-More display methods can be found in [@Wickham2011-uz].
+<!-- SL: I think this paragraph could moved down to the bottom of this section -->
+Various displays are available to show projections in 1-$d$, 2-$d$ and higher dimensions.
+A 1-$d$ projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2-$d$ projections are displayed as scatterplots.
+Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates.
 Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables.
@@
 
+<!-- SL: I think this needs to be linked up to the notation section -->
+
 A sequence of projections can be generated via different tour paths, i.e.
@@
 Instead of randomly generating targets, guided tour [@Cook1995-ae] finds its targets with a technique called projection pursuit [@Friedman1974-ck].
-In projection pursuit, an index function is a surjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
-Guided tour iteratively optimizes the index function in the projection pursuit and produces the most interesting projection according to the index measure.
+In projection pursuit, an index function is defined that is a surjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
+The guided tour iteratively optimizes the index function in the projection pursuit and produces the most interesting projection according to the index measure.
 The right panel of Figure \ref{fig:notation-target-gen} shows an example of a guided tour path.
@@
 The smooth animation between the selected target frames is generated via a step-wise rotation in the high-dimensional space.
-Different methods to derive this rotation were described in @Buja2005-cx, and we can distinguish two types of interpolation [EW: ??].
-When interpolating between frames we rotate the starting basis onto the exact basis representation selected as a basis, and @Buja2005-cx describes three different methods to achieve this (based on decompositions of orthogonal matrices, Givens decompositions, or Householder decompositions).
+There are two different methods to derive this rotation, first described in @Buja2005-cx., When interpolating between frames we rotate the starting basis onto the exact basis representation selected as a basis, and @Buja2005-cx describes three different methods to achieve this (based on decompositions of orthogonal matrices, Givens decompositions, or Householder decompositions) <!--SL : this sentence is confusing -->.
 Alternatively, when interpolating between planes the rotation is constructed towards any basis that represents the plane corresponding to the target frame, such that it follows the shortest path, i.e.
@@
 The geodesic interpolation is derived using singular value decomposition and is implemented in the `tourr` package.
-[EW: this paragraph perhaps needs some rework.]
+
+<!-- SL: The above paragraph is very confusing to me, what is a plane and frame? what is a geodesic? why should it be a geodesic? -->
 
@@
 As explained above this means a tour path is a trace along a Stiefel manifold.
-In the case of one-dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure \ref{fig:notation-target-gen}, and for two-dimensional projections projections fall onto a torus shape.
+<!-- I still don't know if this adds anything, without it being defined --> In the case of one dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure \ref{fig:notation-target-gen}, and for two dimensional projections projections fall onto a torus shape.
 We can use a tour to visualize such a path, as shown in Figure \ref{fig:interpolation-2d}.
@@
 
-Transformations of the data are common in the preprocessing [EW: "preprocessing" means b/f projecting?], such as scaling, sphering, or a logarithmic transformation for skewed distributions.
+Transformations of data are often useful prior to touring such as scaling, sphering, or a logarithmic transformation for skewed distributions.
 However, we may also want to transform the data after projecting onto lower dimensions, to correct for unwanted effects of the projections.
@@
 The sage display [@Laa2020-uv] proposes a solution to this issue, via a radial transformation of the projected data.
-This non-linear transformation is defined in each point of the projection plane and is sensitive to the overall scale of the data as well as the original dimensionality $p$.
+This nonlinear transformation is defined in each point of the projection plane and is sensitive to the overall scale of the data as well as the original dimensionality $p$.
 In addition, tuning parameters can be used to obtain a more (or less) aggressive redistribution of the points.
@@
 This review highlights the history of the grand tour and its modern extensions for the visualization of high dimensional numerical data.
-A rich ecosystem of tour methods are being actively developed and are finding applications in a broad range of areas.
+A rich ecosystem of tour methods are being actively developed and are finding applications in a broad range of areas such as machine learning and physics.
 Tours are being combined with interactive graphics to assist with cluster analysis (the spinifex and liminal packages), displays of projected data are augmented to allow the identification of local structures (the sage algorithm, slicing and section pursuit), and  new visualizations based on tour methods are used to better diagnose optimization procedures for projection pursuit (ferrn).
@@
 
+<!--
 * highlighting the recent developments, spinifex, liminal, sage, slicing and sectioning, ferrn etc. 
 * what's next - conditional model visualisation (Catherine Hurley's work), a interactive graphics system that easily integrates tours that provides the direct manipulations we want, pre-processing / transformations for high dimensional data, penalised projection pursuit, different kinds of data and spaces
+-->
 

### 2021-04-13 paper.Rmd
figure references made consistent

@@
 The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.
-Adapted from E.-K.
+(Adapted from E.-K.
 Lee et al.
-(2005), Figure 1.
-}
+2005, Figure 1.) }
 \label{fig:notation-huber}
@@
 }
-\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from Laa, Cook and Valencia (2020), Fig. 4.}
+\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). (Adapted from Laa, Cook and Valencia, 2020, Figure 4.)}
 \label{fig:tour-slice}
@@
 Right: using the $\gamma$ tuning parameter can also reveal the structure.
-Adapted from Laa, Cook and Lee, Fig.
-9.}
+(Adapted from Laa, Cook and Lee, 2020, Figure 9.)}
 \label{fig:pollen}
@@
 Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.
-Adapted from Lee, Laa and Cook, 2020 Figure 2.}
+(Adapted from Lee, Laa and Cook, 2020, Figure 2.)}
 \label{fig:tour-controls}
@@
 Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.
-Adapted from Cook and Swayne, 2007 Figure 5.3.}
+(Adapted from Cook and Swayne, 2007, Figure 5.3.)}
 \label{fig:prim7-tour}
@@
 (Right) The complete prediction region shows the points classified as purple are embedded within the green slice.
-Adapted from Wickham, Hoffman and Cook, 2015.}
+(Adapted from Wickham, Hofmann and Cook, 2015, Figure 6.)}
   \label{fig:svm-tour}
@@
 The axes correspond to the probabilities that the model of an image belonging to a certain class.
-Adapted from Li, Zhao and Scheidigger, 2020.}
+(Adapted from Li, Zhao and Scheidigger, 2020.)}
 \label{fig:tour-nn}
@@
 Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.
-Adapted from Cook, Laa and Valencia, 2018, Figures 6 and 7.}
+(Adapted from Cook, Laa and Valencia, 2018, Figures 6 and 7.)}
 \label{fig:physics}
@@
 The sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.
-Adapted from Laa, Lee and Cook, 2020 Figure 6.}
+(Adapted from Laa, Lee and Cook, 2020 Figure 6.)}
 \label{fig:mouse-cluster}
@@
   \includegraphics[width=0.24\textwidth]{figures/torus-flat-6-1.pdf}\includegraphics[width=0.24\textwidth]{figures/torus-flat-6-2.pdf}
-\caption{Several views of torii: (two at left) 4D ring torus, and (two at right) 6D flat torus. In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes.}
+\caption{Several views of torii: (two at left) 4D ring torus, and (two at right) 6D flat torus.
+In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes.
+(Adapted from Schloerke, et al., 2016, Figures 10 and 13.)}
 \label{fig:geozoo}

### 2021-04-13 paper.Rmd
reflect 5D cubes not 6D

@@
 Figure \ref{fig:cubes} shows a way to imagine this.
-It shows a sequence of cube wireframes, ranging from 1D through to 6D, where beyond 2D is a linear projection of the cube.
+It shows a sequence of cube wireframes, ranging from 1D through to 5D, where beyond 2D is a linear projection of the cube.
 As dimension increase, a new orthogonal axis is added.
@@
 \caption{Illustrating what is meant by "high-dimensional" in this paper, and a "linear projection".
-From a sequence of increasing dimension cubes, from 1D to 6D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles.
-(Note to Di: to change this diagram to highlight the doubling.)}
+From a sequence of increasing dimension cubes, from 1D to 5D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles.
+}
 \label{fig:cubes}

### 2021-04-13 paper.Rmd
small changes, typos

@@
     email: dicook@monash.edu
-  - name: Natalia Da Silva
+  - name: Natalia da Silva
     address: |
@@
 Instead of randomly generating targets, guided tour [@Cook1995-ae] finds its targets with a technique called projection pursuit [@Friedman1974-ck].
-In projection pursuit, an index function is defined that is a surjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
+In projection pursuit, an index function is defined that is a subjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
 The guided tour iteratively optimizes the index function in the projection pursuit and produces the most interesting projection according to the index measure.

### 2021-04-13 paper.Rmd
Section 4 information was missing
revised fig 3

@@
    \centering
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour1_r.pdf}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour2_r.pdf}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour3_r.pdf}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour4_r.pdf}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour1.png}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour2.png}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour3.png}
+         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour4.png}
    \end{minipage}%
-\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and projections seen during a tour (B.1-4).
+\caption{Comparison of structure perception between nonlinear dimension reduction from 10D, using t-SNE (A) and a tour (B).
 There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied.
-This can be seen in the tour projections.
-The two green clusters are spherical in shape, and very distance from the brown clusters.
+This can be seen in sample of tour projections shown.
+The two green clusters are (almost) spherical in shape, and very distance from the brown clusters.
 Three of the brown clusters are very close to each other (just visible in B.1), and all brown clusters are elliptical.
@@
 By its nature a tour is most effective to analyst when combined with interactivity; section \@ref(interaction) reviews the components of user interfaces for manipulating tour views.
+Section \@ref(tours) shows manual tours, these are useful to test structure of selected features.
 The implementation of tour paths in statistical software is reviewed in section \@ref(software).
@@
 
-# Manual Tours
+# Manual Tours {#tours}
 

### 2021-04-13 data/clusters.R

@@ -42,7 +42,7 @@ tsne_df <- tibble(tsneX = tsne_y$Y[,1],
                     tsneY = tsne_y$Y[,2], 
                     cluster = X_tbl$cluster)
 ggplot(tsne_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
-  geom_point() + coord_equal()
+  geom_point() + coord_equal() 
 
 # multichallenge data set
 url <- "http://ifs.tuwien.ac.at/dm/download/multiChallenge-matrix.txt"
@@ -66,7 +66,14 @@ tsne_multi_df <- tibble(tsneX = tsne_multi$Y[,1],
                   cluster = multi$group)
 ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
   geom_point() + coord_equal() +
-  scale_colour_brewer("", palette="Dark2")
+  scale_colour_brewer("", palette="Dark2") +
+  ggtitle("A. t-SNE") + 
+  theme_solid() + 
+  theme(plot.title = element_text(colour = "black", size = 18),
+        legend.text = element_text(colour="black", size = 18),
+        panel.background = element_rect(fill = "white"),
+        panel.border = element_rect(fill = NA, colour="black", size=1,
+                                        linetype = 1)) 
 
 # tourr
 library(tourr)
@@ -81,7 +88,7 @@ render_gif(multi[,2:11],
 
 render_gif(multi[,2:11], 
            grand_tour(),
-           display_xy(col=clrs, axes="bottomleft", half_range = 0.5, center = TRUE), 
+           display_xy(col=clrs, axes="off", half_range = 0.7, center = TRUE), 
            frames = 200,
            "figures/cluster_example_zoom.gif")
 

### 2021-04-13 paper.Rmd
added reference to the python implementation

@@
 The tour shows that the clusters do not look like this in the full data space.
-The clusters are at various distances apart and very different sizes.
-In the four projections from the tour this can be seen.
-The two green clusters are large and spherical, and far from the brown clusters.
+The clusters are at various distances apart and very different sizes, as can be seen from the four projections from a tour.
+The two green clusters are large and almost spherical, and far from the brown clusters.
 The brown clusters have one larger one, and three smaller, very close to each other.
@@
 This gives some deeper perspective to what is learned from t-SNE, and illustrates what the t-SNE dimension reduction has done: it has found small gaps between points and expanded these gaps to yield the representation.
-It should be noted, though, that the t-SNE view clearly reveals the presence of six distinct clusters.
+It should be noted, though, that the methods (t-SNE and tour) complement each other.
+The t-SNE view gives a clear indication of six clusters, which may have been overlooked on initial viewing with a tour.
 With this information is the invitation to look closer at the data in the tour, to see, that yes, indeed, there are three, tiny, tiny clusters very close to each other.
@@
 \begin{figure*}[!t]
-         \includegraphics[width=0.5\textwidth]{figures/cluster_example_r.pdf}
+         \includegraphics[width=0.371\textwidth]{figures/cluster_example.pdf}
      \vfill
@@
 
-- other work (Di)
+A python implementation of the tour written by @Nguyen2020 is also available, and would benefit from extension by other developers.
 

### 2021-04-13 data/clusters.R

@@ -69,8 +69,8 @@ ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
   scale_colour_brewer("", palette="Dark2") +
   ggtitle("A. t-SNE") + 
   theme_solid() + 
-  theme(plot.title = element_text(colour = "black", size = 18),
-        legend.text = element_text(colour="black", size = 18),
+  theme(plot.title = element_text(colour = "black", size = 12),
+        legend.text = element_text(colour="black", size = 12),
         panel.background = element_rect(fill = "white"),
         panel.border = element_rect(fill = NA, colour="black", size=1,
                                         linetype = 1)) 

### 2021-04-14 paper.Rmd
consistency of ...-d vs ...D
fix figure path
remove duplicate reference
moving display section down
moving back into interaction

@@
 It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (see for e.g.
-@Swayne2003-cc or @Tierney1991).
+@Swayne2003 or @Tierney1991).
 To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
@@
 \begin{figure*}[!t]
-         \includegraphics[width=0.371\textwidth]{figures/cluster_example.pdf}
+         \includegraphics[width=0.371\textwidth]{figures/cluster_example_r.pdf}
      \vfill
@@
    \end{minipage}%
-\caption{Comparison of structure perception between nonlinear dimension reduction from 10D, using t-SNE (A) and a tour (B).
+\caption{Comparison of structure perception between nonlinear dimension reduction from 10-d, using t-SNE (A) and a tour (B).
 There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied.
@@
 
-The rest of the review is structured as follows: section \@ref(notation) defines the notation and components of a tour displays.
-By its nature a tour is most effective to analyst when combined with interactivity; section \@ref(interaction) reviews the components of user interfaces for manipulating tour views.
-Section \@ref(tours) shows manual tours, these are useful to test structure of selected features.
-The implementation of tour paths in statistical software is reviewed in section \@ref(software).
+The rest of the review is structured as follows: Section \@ref(notation) defines the notation and components of a tour displays.
+By its nature a tour is most effective to analyst when combined with interactivity; Section \@ref(interaction) reviews the components of user interfaces for manipulating tour views, including manual tours (Section \@ref(tours)) which are useful to test structure of selected features.
+The implementation of tour paths in statistical software is reviewed in Section \@ref(software).
 Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning and applied statistics.
-Finally, section \@ref(discussion) discusses future research directions for tours.
+Finally, Section \@ref(discussion) discusses future research directions for tours.
 
@@
 \end{figure*}
-
-## Display
-
-<!-- SL: I think this paragraph could moved down to the bottom of this section -->
-Various displays are available to show projections in 1-$d$, 2-$d$ and higher dimensions.
-A 1-$d$ projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2-$d$ projections are displayed as scatterplots.
-Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates.
-Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.3\textwidth]{figures/notation-display-1d.pdf}
-\includegraphics[width=0.3\textwidth]{figures/notation-display-2d.pdf}
-\includegraphics[width=0.3\textwidth]{figures/notation-display-5d.pdf}
-\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5d display in parallel coordinate plot (right).}
-\label{fig:display}
-\end{figure}
 
@@
 \label{fig:interpolation-2d}
+\end{figure}
+
+## Display
+
+<!-- SL: I think this paragraph could moved down to the bottom of this section -->
+Various displays are available to show projections in 1-$d$, 2-$d$ and higher dimensions.
+A 1-$d$ projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2-$d$ projections are displayed as scatterplots.
+Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates.
+Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.3\textwidth]{figures/notation-display-1d.pdf}
+\includegraphics[width=0.3\textwidth]{figures/notation-display-2d.pdf}
+\includegraphics[width=0.3\textwidth]{figures/notation-display-5d.pdf}
+\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5d display in parallel coordinate plot (right).}
+\label{fig:display}
 \end{figure}
@@
 
-# Manual Tours {#tours}
+## Manual Tours {#tours}
 

### 2021-04-14 paper.Rmd
Comment in the Intro about UMAP

@@
 With this information is the invitation to look closer at the data in the tour, to see, that yes, indeed, there are three, tiny, tiny clusters very close to each other.
+
+<!-- Reading the introduction I noticed maybe can be included UMAP, I'm not an expert (Uniform Manifold approximation and projection for dimension reduction ) I saw this before but maybe some of you know about it to evaluated if worth it to be include it here.
+I have an arXiv paper https://arxiv.org/abs/1802.03426 maybe some of you can identify if it is relevant or not or at least you know about the authors ....just a comment (Di, or ALL )-->
 

### 2021-04-14 paper.Rmd
Comment about Figure 3, color description to change (dark2 palette)

@@
 \end{figure*}
+
+<!-- FIGURE 3, I think you are using Dark2 palette but in the figure caption you mention green and brown, I'm not colorblind but I check the palette description and the color is #D95F02 (https://www.colorhexa.com/d95f02) color description: Vivid orange.
+If this is correct, you should  replace  "brown" by orange in  Figure 3  caption and in  Figure 3 comments in page 4 (two brown each)-->
 

### 2021-04-14 paper.Rmd
add umap citation

--- 2021-04-14 paper.Rmd
@@
 
-Nonlinear dimension reduction techniques such as t-SNE have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary.
+Nonlinear dimension reduction techniques such as t-SNE and Uniform Manifold Alignment and Projection (UMAP) have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary [@Van_der_Maaten2008-qa;@mcinnes2020umap].
 However, it is only a summary, and it likely involves substantial warping of the original data space.
@@
 Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
-@li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
+@li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP [@mcinnes2020umap] and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.
 

### 2021-04-14 paper.Rmd
add reference to huber plot

@@
 (Not sure this adds anything) --> With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows an example of lower dimensional projections of the palmer penguins data in a Huber plot and a histogram [@penguins].
+Figure \ref{fig:notation-huber} shows an example of lower dimensional projections of the palmer penguins data in a Huber plot and a histogram [@huber1990;@penguins].
 

### 2021-04-14 paper.Rmd
new cluster image
tweaks to notation section

@@
   fig.show = 'hold',
-  dpi = 300 #,
+  dpi = 300,
+  cache = FALSE
   # out.width = "\\textwidth",
@@
 \begin{figure*}[!t]
-         \includegraphics[width=0.371\textwidth]{figures/cluster_example_r.pdf}
+         \includegraphics[width=0.371\textwidth]{figures/cluster_example.pdf}
      \vfill
@@
 
+\begin{figure*}
+   \centering
+         \includegraphics[width=\textwidth]{figures/tour_path.png}
+\caption{Illustration of a tour path, indicating target planes, and interpolation planes. (Adapted from Buja et al. 2005, Figure 1.) }
+\label{fig:tour-path}
+\end{figure*}
+
 ## Finding targets
@@
 
-A sequence of projections can be generated via different tour paths, i.e.
-grand tour, guided tour, manual tour, little tour, local tour, and most recently, slice tour.
-The focus of this section will be on the grand and guided tour, with manual and slice tour being introduced in the later sections of the review.
+A tour path, that is a sequence of projections, can be generated by geodesic interpolation between a set of target planes (Figure \ref{fig:tour-path}).
+Different methods for choosing target planes provide different tour paths.
+The grand tour is generated using randomly selected target.
+A guided tour is generated by choosing particularly structured projection from a projection pursuit optimisation.
+The little tour uses all variable bases as targets, and a local tour rocks back and forth from a particular plane to randomly chosen targets in a small neighbourhood.
 
-Grand tour, coined by @Asimov1985-xr, randomly samples target bases in the space and constructs an interpolation between them to produce a sequence of projections as in Figure  \ref{fig:notation-target-gen}.
-This tour path is suitable for an initial view of the data, however, as the dimension increases, it could take a long time before the grand tour finds a projection.
+Figure \ref{fig:notation-target-gen} shows a representation of 1D tour paths of 5D data, drawn on a PCA space.
+A grand tour path is on the left and a guided tour path is on the right.
+The grand tour path should be more wide-spread because it is attempting to show all possible projections, if left to run long enough.
+The guided tour path should be short, as the projection pursuit optimisation zeroes on the optimally interesting projection.
 
-Instead of randomly generating targets, guided tour [@Cook1995-ae] finds its targets with a technique called projection pursuit [@Friedman1974-ck].
-In projection pursuit, an index function is defined that is a subjective mapping from the projection $Y$ in $\mathbb{R}^{n \times d}$ to a scalar in $\mathbb{R}$.
-The guided tour iteratively optimizes the index function in the projection pursuit and produces the most interesting projection according to the index measure.
-The right panel of Figure \ref{fig:notation-target-gen} shows an example of a guided tour path.
+There are other tour types which don't fit this style of tour.
+The manual tour [@cook_manual_1997] allows the user to change the projection coefficients manually, to rotate a variable into and out of a project, and is discussed later.
+The recently developed slice tour [@Laa2020-js] can be applied to any of the above tour types.
+It displays a slice through the orthogonal space, as opposed to a projection, and is explained in more details later.
 
@@
 
-The smooth animation between the selected target frames is generated via a step-wise rotation in the high-dimensional space.
-There are two different methods to derive this rotation, first described in @Buja2005-cx., When interpolating between frames we rotate the starting basis onto the exact basis representation selected as a basis, and @Buja2005-cx describes three different methods to achieve this (based on decompositions of orthogonal matrices, Givens decompositions, or Householder decompositions) <!--SL : this sentence is confusing -->.
-Alternatively, when interpolating between planes the rotation is constructed towards any basis that represents the plane corresponding to the target frame, such that it follows the shortest path, i.e.
-the geodesic, avoiding any in-plane rotation.
-The geodesic interpolation is derived using singular value decomposition and is implemented in the `tourr` package.
+The smooth progression in the tour path is due to geodesic interpolation between the target planes.
+This takes into consideration two important aspects: (1) maintains the orthonormality of the projection bases, (2) contains all the rotation to be between planes, not a particular basis in any plane.
+The first is clearly important because it ensures that we a looking at low-dimensional projection of the data always.
+The latter is harder to explain, but really important from a visual perspective.
+It stops any within-plane spin, and could be considered stabilising the view.
+More details explanation of this can be found in @Buja2005-cx.
 
-<!-- SL: The above paragraph is very confusing to me, what is a plane and frame? what is a geodesic? why should it be a geodesic? -->
-
-Typically the geodesic interpolation is preferred, since within-plane rotation is distracting and does not provide new information.
-An exception might be the case of a projection pursuit guided tour with an index function that is sensitive to these rotations [@Laa2020-zj].
-
-For any method we need to ensure that each projection along the interpolated path is still orthonormal.
-As explained above this means a tour path is a trace along a Stiefel manifold.
-<!-- I still don't know if this adds anything, without it being defined --> In the case of one dimensional projections, the set of bases traces a path along the surface of a hypersphere of radius one, as illustrated in Figure \ref{fig:notation-target-gen}, and for two dimensional projections projections fall onto a torus shape.
-We can use a tour to visualize such a path, as shown in Figure \ref{fig:interpolation-2d}.
+Figure \ref{fig:interpolation-2d} shows tour paths of 2D projections of 6D data.
+The space of all tour paths is a high-dimensional torus, as represented by the grey points.
+The tour paths are shown in green and brown, and each dot indicates a projection in the sequence.
+The plots in this figure could be considered to be a tour looking at itself, because each plot is a selected view from a tour of the torus with the paths overlaid.
 
@@
 
-Interactive systems like GGobi provide the option of adding sectioning information via linked brushing.
+Interactive systems like GGobi [@Swayne2003] provide the option of adding sectioning information via linked brushing.
 Sectioning means we select points that fall in a section of the full parameter space, for example for the purpose of highlighting them in a tour display.
@@
 
-This review highlights the history of the grand tour and its modern extensions for the visualization of high dimensional numerical data.
-A rich ecosystem of tour methods are being actively developed and are finding applications in a broad range of areas such as machine learning and physics.
-Tours are being combined with interactive graphics to assist with cluster analysis (the spinifex and liminal packages), displays of projected data are augmented to allow the identification of local structures (the sage algorithm, slicing and section pursuit), and  new visualizations based on tour methods are used to better diagnose optimization procedures for projection pursuit (ferrn).
+This review highlights the history of tour methods, and modern extensions for the visualization of high dimensional numerical data.
+The rich ecosystem of tours is being actively developed and applied to a broad range of problems including machine learning.
+When tours are be combined with some interactivity it can assist with analyses like clustering (the spinifex and liminal packages).
+Modified displays of projected data can be used to expose local structures (the sage algorithm, slicing and section pursuit).
+New algorithm data collection inside the tour code is being used to better diagnose optimization procedures for projection pursuit (ferrn).
 
-There are many possible future directions for tour research, one big limitation
-of the current software ecosystem is that analysts are no longer able to perform interactions that were possible in software packages from the 80s and 90s.
-An interactive graphics system that integrated the tour methodology directly would be a big step forward.
-Such an interface could also be combined with methods for conditional model visualization, for example as implemented in the condviz package [@JSSv081i05], allowing the visual exploration of multivariate fitted models.
+There are many possible future directions for tour research.
+One big challenge for software engineering is to seamlessly embed the tour with interactivity.
+Currently, there is no implementation that analysts can use to perform tasks that were easy in software available in the 80s and 90s.
+More contributions in sectioning as an alternative to projection, could be broadly useful.
+An example is conditional model visualization, as implemented in the condviz package [@JSSv081i05], which assists with visual exploration of multivariate fitted models.
+It could be interesting to define tour paths for different types of nonlinear subspaces, or non-Euclidean space.
 

### 2021-04-14 paper.Rmd
minor fixes

@@
 The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.
-(Adapted from E.-K.
-Lee et al.
+(Adapted from Lee et al.
 2005, Figure 1.) }
@@
 <!-- SL: I think this paragraph could moved down to the bottom of this section -->
-Various displays are available to show projections in 1-$d$, 2-$d$ and higher dimensions.
-A 1-$d$ projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2-$d$ projections are displayed as scatterplots.
+Various displays are available to show projections in 1D, 2D and higher dimensions.
+A 1D projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2D projections are displayed as scatterplots.
 Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates.

### 2021-04-14 paper.Rmd
redo interpolation figure

@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/notation-2d-space.pdf}
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus001.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus017.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus064.png}}
+\centerline{
+\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus068.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus075.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus100.png}
+}
 \caption{Still views of a tour showing the space of possible two-dimensional projections as a torus shape in grey, and the trace of two short paths obtained via geodesic interpolation traced in color.}
 \label{fig:interpolation-2d}
-\end{figure}
+\end{figure*}
 
@@
 Modified displays of projected data can be used to expose local structures (the sage algorithm, slicing and section pursuit).
-New algorithm data collection inside the tour code is being used to better diagnose optimization procedures for projection pursuit (ferrn).
+New algorithm data collection inside the tour code is being used to better diagnose optimization procedures for projection pursuit [@zhang2021visual].
 

### 2021-04-14 paper.Rmd
Acknowledgements section.
small changes using American English  spelling (as Stuart Lee ISSUE#2)

@@
 With multiple variables, especially if there is some association between variables, this would be called multivariate or high-dimensional data.
-When the variables are all numeric, or quantitative, visualisation often relies on some form of dimension reduction.
+When the variables are all numeric, or quantitative, visualization often relies on some form of dimension reduction.
 This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936].
@@
 The grand tour is generated using randomly selected target.
-A guided tour is generated by choosing particularly structured projection from a projection pursuit optimisation.
-The little tour uses all variable bases as targets, and a local tour rocks back and forth from a particular plane to randomly chosen targets in a small neighbourhood.
+A guided tour is generated by choosing particularly structured projection from a projection pursuit optimization.
+The little tour uses all variable bases as targets, and a local tour rocks back and forth from a particular plane to randomly chosen targets in a small neighborhood.
 
@@
 The grand tour path should be more wide-spread because it is attempting to show all possible projections, if left to run long enough.
-The guided tour path should be short, as the projection pursuit optimisation zeroes on the optimally interesting projection.
+The guided tour path should be short, as the projection pursuit optimization zeroes on the optimally interesting projection.
 
@@
 The latter is harder to explain, but really important from a visual perspective.
-It stops any within-plane spin, and could be considered stabilising the view.
+It stops any within-plane spin, and could be considered stabilizing the view.
 More details explanation of this can be found in @Buja2005-cx.
@@
 Figure \ref{fig:interpolation-2d} shows tour paths of 2D projections of 6D data.
-The space of all tour paths is a high-dimensional torus, as represented by the grey points.
+The space of all tour paths is a high-dimensional torus, as represented by the gray points.
 The tour paths are shown in green and brown, and each dot indicates a projection in the sequence.
@@
 
-Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualisation.
+Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualization.
 Often one will want to pause on a particularly interesting projection and return it for use in a downstream analysis or static plot.
@@
 The comparison typically relies on numeric computation of the model predictions for all observables, potentially obscuring the nature of their dependence on the model parameters.
-Here, multivariate visualisation methods can provide new insights.
+Here, multivariate visualization methods can provide new insights.
 This was demonstrated in @Cook2018-jm.
@@
 
+# Acknowledgements {-}
+
+The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@Xie2018-bg] using `knitr` [@Xie2017-uh].
+
 # References {-}

### 2021-04-15 paper.Rmd
slowly fixing references

--- 2021-04-14 paper.Rmd
@@
     toc: false
-bibliography: references.bib
-biblio-style: apalike
+bibliography: "references.bib"
 ---

### 2021-04-15 paper.Rmd
replace color per suggestion

@@
 The clusters are at various distances apart and very different sizes, as can be seen from the four projections from a tour.
-The two green clusters are large and almost spherical, and far from the brown clusters.
-The brown clusters have one larger one, and three smaller, very close to each other.
+The two green clusters are large and almost spherical, and far from the orange clusters.
+The orange clusters have one larger one, and three smaller, very close to each other.
 All of these are elliptical, which means that they have very little variability, actually no variability if you watch the full tour, in some of the 10 dimensions.
@@
 This can be seen in sample of tour projections shown.
-The two green clusters are (almost) spherical in shape, and very distance from the brown clusters.
-Three of the brown clusters are very close to each other (just visible in B.1), and all brown clusters are elliptical.
+The two green clusters are (almost) spherical in shape, and very distance from the orange clusters.
+Three of the orange clusters are very close to each other (just visible in B.1), and all orange clusters are elliptical.
 The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the dimension reduction.}
@@
 \end{figure*}
-
-<!-- FIGURE 3, I think you are using Dark2 palette but in the figure caption you mention green and brown, I'm not colorblind but I check the palette description and the color is #D95F02 (https://www.colorhexa.com/d95f02) color description: Vivid orange.
-If this is correct, you should  replace  "brown" by orange in  Figure 3  caption and in  Figure 3 comments in page 4 (two brown each)-->
 
@@
 The space of all tour paths is a high-dimensional torus, as represented by the gray points.
-The tour paths are shown in green and brown, and each dot indicates a projection in the sequence.
+The tour paths are shown in green and orange, and each dot indicates a projection in the sequence.
 The plots in this figure could be considered to be a tour looking at itself, because each plot is a selected view from a tour of the torus with the paths overlaid.

### 2021-04-15 paper.Rmd
use apa style to fix abbreviations

@@
 bibliography: "references.bib"
+csl: apa.csl
 ---

### 2021-04-15 paper.Rmd
figure -> Figure

@@
 Often one will want to pause on a particularly interesting projection and return it for use in a downstream analysis or static plot.
-Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown in figure \ref{fig:tour-controls}.
+Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown in Figure \ref{fig:tour-controls}.
 
@@
 Manual tours can then be employed to test the structure of the feature, with respect to a selected variable.
-For instance, figure \ref{fig:manual-tour}, starts from the orthonormal linear discriminant and explores the sensitivity of class separation as the contribution of a single variable is altered.
+For instance, Figure \ref{fig:manual-tour}, starts from the orthonormal linear discriminant and explores the sensitivity of class separation as the contribution of a single variable is altered.
 

### 2021-04-15 paper.Rmd
Link to github added to acknowledgements.
reshape figures

@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus001.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus017.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus064.png}}
+\centerline{\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus001.png}
+\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus017.png}
+\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus064.png}}
 \centerline{
-\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus068.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus075.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/notation-torus-frames/torus100.png}
+\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus068.png}
+\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus075.png}
+\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus100.png}
 }
@@
 \includegraphics[width=0.3\textwidth]{figures/notation-display-5d.pdf}
-\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5d display in parallel coordinate plot (right).}
+\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5D display in parallel coordinate plot (right).}
 \label{fig:display}
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{figures/roman-surface-centered-010.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/torus-4-centered-015.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/cube-6-centered-001.png}}
+\centerline{\includegraphics[width=0.3\textwidth]{figures/roman-surface-centered-010.png}
+\includegraphics[width=0.3\textwidth]{figures/torus-4-centered-015.png}
+\includegraphics[width=0.3\textwidth]{figures/cube-6-centered-001.png}}
 \centerline{
-\includegraphics[width=0.2\textwidth]{figures/roman-surface-centered-032.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/torus-4-centered-033.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{figures/cube-6-centered-063.png}
+\includegraphics[width=0.3\textwidth]{figures/roman-surface-centered-032.png}
+\includegraphics[width=0.3\textwidth]{figures/torus-4-centered-033.png}
+\includegraphics[width=0.3\textwidth]{figures/cube-6-centered-063.png}
 }

### 2021-04-15 R/tour-notation.R
trying to get minimum index value computed too

@@ -2,6 +2,7 @@
 library(ggplot2)
 library(palmerpenguins)
 library(dplyr)
+library(PPtreeViz)
 penguins_small <- penguins %>% filter(!is.na(bill_length_mm))
 p <- PPtreeViz::Huberplot(penguins_small[,3:4],penguins_small$species,PPmethod="LDA") 
 
@@ -80,4 +81,107 @@ p <- ggplot() +
 ggsave(p, filename = "notation-target-gen.pdf", 
        path = here::here("figures"),
        width = 10, height = 5, units = "in")
+
+# This is a hack to get minimum index value in the Huber plot, too
+Huberplot_min <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRUE, 
+          r = 1, lambda = 0.5, opt.proj = TRUE, UserDefFtn = NULL, 
+          ...) 
+{
+  index <- NULL
+  best.proj <- NULL
+  best.index <- 0
+  origdata2D <- as.matrix(origdata2D)
+  for (i in 0:360) {
+    theta <- pi/180 * i
+    proj.data <- matrix(cos(theta) * origdata2D[, 1] + sin(theta) * 
+                          origdata2D[, 2])
+    proj <- matrix(c(cos(theta), sin(theta)), ncol = 1)
+    if (PPmethod == "LDA") {
+      newindex <- LDAindex(origclass, origdata2D, proj = proj, 
+                           weight = weight)
+    }
+    else if (PPmethod == "PDA") {
+      newindex <- PDAindex(origclass, origdata2D, proj, 
+                           weight = weight, lambda = lambda)
+    }
+    else if (PPmethod == "Lr") {
+      newindex <- Lrindex(origclass, origdata2D, proj, 
+                          weight = weight, r = r)
+    }
+    else if (PPmethod == "GINI") {
+      newindex <- GINIindex1D(origclass, origdata2D, proj)
+    }
+    else if (PPmethod == "ENTROPY") {
+      newindex <- ENTROPYindex1D(origclass, origdata2D, 
+                                 proj)
+    }
+    else if (PPmethod == "UserDef") {
+      newindex <- UserDefFtn(proj.data, ...)
+    }
+    index <- c(index, newindex)
+  }
+  sel.index <- which(index[1:360] < signif(min(index), 6) - 
+                       1e-06)
+  theta.best.all <- pi/180 * (sel.index - 1)
+  theta.best <- theta.best.all[1]
+  proj.data.best <- matrix(cos(theta.best) * origdata2D[, 1] + 
[…]

### 2021-04-15 R/tour-notation.R
Huber plots modified

@@ -4,11 +4,6 @@ library(palmerpenguins)
 library(dplyr)
 library(PPtreeViz)
 penguins_small <- penguins %>% filter(!is.na(bill_length_mm))
-p <- PPtreeViz::Huberplot(penguins_small[,3:4],penguins_small$species,PPmethod="LDA") 
-
-ggsave(p, filename = "notation-huber.pdf", 
-       path = here::here("figures"),
-       width = 12, height = 5, units = "in")
 
 # figure 3 diagram illustrating different displays for different d
 library(tourr)
@@ -83,8 +78,9 @@ ggsave(p, filename = "notation-target-gen.pdf",
        width = 10, height = 5, units = "in")
 
 # This is a hack to get minimum index value in the Huber plot, too
-Huberplot_min <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRUE, 
+Huberplot2 <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRUE, 
           r = 1, lambda = 0.5, opt.proj = TRUE, UserDefFtn = NULL, 
+          min = FALSE, scale_data = 1,
           ...) 
 {
   index <- NULL
@@ -120,8 +116,12 @@ Huberplot_min <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRU
     }
     index <- c(index, newindex)
   }
-  sel.index <- which(index[1:360] < signif(min(index), 6) - 
+  if (!min) {
+    sel.index <- which(index[1:360] > signif(max(index), 6) - 
                        1e-06)
+  } else
+    sel.index <- which(index[1:360] < signif(min(index), 6) + 
+                         1e-06)
   theta.best.all <- pi/180 * (sel.index - 1)
   theta.best <- theta.best.all[1]
   proj.data.best <- matrix(cos(theta.best) * origdata2D[, 1] + 
@@ -160,8 +160,8 @@ Huberplot_min <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRU
   max.X <- max(unlist(plot.data))
   P1 <- ggplot(data = plot.data, aes(x = data.X, y = data.Y)) + 
     geom_path() + geom_path(aes(x = data.cX, y = data.cY), 
-                            linetype = "dashed") + geom_point(data = point.data, 
-                                                              aes(x = x, y = y, color = group, shape = group)) + scale_x_continuous(breaks = NULL) + 
+                            linetype = "dashed") + 
[…]

### 2021-04-15 paper.Rmd
typo

@@
    \centering
-         \includegraphics[width=\textwidth]{figures/notation-huber.pdf}
-\caption{Huber's plot (left) of the penguins data (bill length and depth only).
+         \includegraphics[width=\textwidth]{figures/notation-huber1.pdf}
+         \includegraphics[width=\textwidth]{figures/notation-huber2.pdf}
+\caption{Huber's plot (left plots) of the penguins data (bill length and depth only).
 The solid ellipse represents how well the projection, onto different directions, separates the three species and the dashed circle is a reference guide.
-The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the huber's plot.
+The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
 (Adapted from Lee et al.
@@
 
-The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@Xie2018-bg] using `knitr` [@Xie2017-uh].
+The authors gratefully acknowledge the support of the Australian Research Council.
+The paper was written in `rmarkdown` [@Xie2018-bg] using `knitr` [@Xie2017-uh].
+The source material for this paper is available at \url{https://github.com/dicook/wiley-isghdd}.
 

### 2021-04-15 paper.Rmd
Some comments about Figures, included in Issues too

@@
 
+<!-- Figure 4:   Huber's plot (left plots) of the penguins data....Maybe will be better to  capital letter for axis labels (Count, Group and Best projected data).
+And maybe use the same Dark2 palette.
+-->
+
 \begin{figure*}
@@
 \end{figure}
+
+<!-- Figure 6: "A comparison of guided....." and Figure 7:"Still views of a tour showing the space...." maybe change the color (brown) because are difficult to distinguish.
+Similar problem in Figure 12: Three  projections  from  a  manual  tour  where  variable  "bl" ....
+"  Not sure if you have selected a colorblind friendly palette but I check these Figures with a colorblind person (small sample but...) and was very  difficult  for him to distinguish the mixed colors.
+-->
 

### 2021-04-15 prepare_arxiv.R
script for zipping up materials required for arxiv

@@ -1,0 +1,21 @@
+prepare_for_arxiv <- function() {
+  
+  # assumes you have done the incantation after generating the tex files of
+  # pdflatex paper; biber paper; pdflatex paper
+  
+  if (!dir.exists("arxiv")) dir.create("arxiv")
+  
+  zip_name <- "arxiv/arxiv-submission.zip"
+  tex_files <-list.files(pattern = "tex$|bbl$")
+  figures <- list.files("./figures/", recursive = TRUE, full.names = TRUE)
+  
+  
+  files_to_zip <-c(tex_files, figures)
+  
+  if (file.exists(zip_name)) {
+    file.remove(zip_name)
+  }
+  
+  zip(zipfile = zip_name, files = files_to_zip)
+  
+}

### 2021-04-15 paper.Rmd
Figure 5 comment
tidy up

--- 2021-04-15 paper.Rmd
@@
 
-<!-- Reading the introduction I noticed maybe can be included UMAP, I'm not an expert (Uniform Manifold approximation and projection for dimension reduction ) I saw this before but maybe some of you know about it to evaluated if worth it to be include it here.
-I have an arXiv paper https://arxiv.org/abs/1802.03426 maybe some of you can identify if it is relevant or not or at least you know about the authors ....just a comment (Di, or ALL )-->
-
 \begin{figure*}[!t]
@@
 A projection basis, $A_{p \times d}$, is matrix that characterizes the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires that each column in $A$ has unit length and are perpendicular.
-<!-- This condition gives the $p$-dimensional basis space a geometrical structure of a Stiefel manifold.
-(Not sure this adds anything) --> With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
+With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
 Figure \ref{fig:notation-huber} shows an example of lower dimensional projections of the palmer penguins data in a Huber plot and a histogram [@huber1990;@penguins].
@@
 \end{figure*}
+
+<!-- Figure 4:   Huber's plot (left plots) of the penguins data....Maybe will be better to  capital letter for axis labels (Count, Group and Best projected data).
+And maybe use the same Dark2 palette.
+-->
 
@@
 
-<!-- SL: I think this paragraph could moved down to the bottom of this section -->
 Various displays are available to show projections in 1D, 2D and higher dimensions.
@@
 
+<!-- Figure 5: "Various displays of the projections..." Maybe increasing the axis text size will be good to see the axis labels -->
 \begin{figure}
@@
 
-A python implementation of the tour written by @Nguyen2020 is also available, and would benefit from extension by other developers.
+A Python implementation of the tour written by @Nguyen2020 is also available, and would benefit from extension by other developers.
 
@@
 
-<!--
-* highlighting the recent developments, spinifex, liminal, sage, slicing and sectioning, ferrn etc. 
-* what's next - conditional model visualisation (Catherine Hurley's work), a interactive graphics system that easily integrates tours that provides the direct manipulations we want, pre-processing / transformations for high dimensional data, penalised projection pursuit, different kinds of data and spaces
--->
-
 # Acknowledgements {-}

### 2021-04-16 paper.Rmd
fig 4 caption done

@@
 With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows an example of lower dimensional projections of the palmer penguins data in a Huber plot and a histogram [@huber1990;@penguins].
+Figure \ref{fig:notation-huber} shows two examples of low-dimensional projections of the palmer penguins data [@penguins] in a Huber plot [@huber1990] and a histogram.
 
@@
 \caption{Huber's plot (left plots) of the penguins data (bill length and depth only).
-The solid ellipse represents how well the projection, onto different directions, separates the three species and the dashed circle is a reference guide.
-The histogram (right) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
+The solid line represents how well the projection onto different directions, separates the three species, and the dashed circle is a reference guide.
+The histogram (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
 (Adapted from Lee et al.

### 2021-04-16 paper.Rmd
further reference fixes and typo edits

@@
     toc: false
+    citation_package: biblatex
 bibliography: "references.bib"
-csl: apa.csl
 ---
@@
 The view on the right responds by highlighting the corresponding points.
-(Adapted from Cook and Swayne, 2007 Figure 2.12)}
+Adapted from \cite{Cook2007-be} Figure 2.12.}
 \label{fig:brushing}
@@
 The histogram (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
-(Adapted from Lee et al.
-2005, Figure 1.) }
+Adapted from \cite{Lee2005-gl}, Figure 1.}
 \label{fig:notation-huber}
@@
          \includegraphics[width=\textwidth]{figures/tour_path.png}
-\caption{Illustration of a tour path, indicating target planes, and interpolation planes. (Adapted from Buja et al. 2005, Figure 1.) }
+\caption{Illustration of a tour path, indicating target planes, and interpolation planes. Adapted from \cite{Buja2005-cx}, Figure 1.}
 \label{fig:tour-path}
@@
 }
-\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). (Adapted from Laa, Cook and Valencia, 2020, Figure 4.)}
+\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
 \label{fig:tour-slice}
@@
 Right: using the $\gamma$ tuning parameter can also reveal the structure.
-(Adapted from Laa, Cook and Lee, 2020, Figure 9.)}
+Adapted from \cite{Laa2020-uv}, Figure 9.}
 \label{fig:pollen}
@@
 Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation.
-(Adapted from Lee, Laa and Cook, 2020, Figure 2.)}
+Adapted from \cite{Lee2020-kg}, Figure 2.}
 \label{fig:tour-controls}
@@
 Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.
-(Adapted from Cook and Swayne, 2007, Figure 5.3.)}
+Adapted from \cite{Cook2007-be}, Figure 5.3.}
 \label{fig:prim7-tour}
@@
 (Right) The complete prediction region shows the points classified as purple are embedded within the green slice.
-(Adapted from Wickham, Hofmann and Cook, 2015, Figure 6.)}
+Adapted from \cite{Wickham2015-do}, Figure 6.}
   \label{fig:svm-tour}
@@
 The axes correspond to the probabilities that the model of an image belonging to a certain class.
-(Adapted from Li, Zhao and Scheidigger, 2020.)}
+Adapted from \cite{Li2020-kg}.}
 \label{fig:tour-nn}
@@
 Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.
-(Adapted from Cook, Laa and Valencia, 2018, Figures 6 and 7.)}
+Adapted from \cite{Cook2018-jm}, Figures 6 and 7.}
 \label{fig:physics}
@@
 \caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation.
-The sage display was used to identify and verify cluster sepearation between the three highlighted clusters estimated from a clustering algorithm.
-(Adapted from Laa, Lee and Cook, 2020 Figure 6.)}
+The sage display was used to identify and verify cluster separation between the three highlighted clusters estimated from a clustering algorithm.
+Adapted from \cite{Laa2020-uv}, Figure 6.}
 \label{fig:mouse-cluster}
@@
 High-dimensional data is on many analysts' minds in recent years, and a way to build understanding of high-dimensional spaces is to examine common high-dimensional shapes using a tour.
-Many different shapes can be simulated using the R package geozoo.
+Many different shapes can be simulated using the R package geozoo [@RJ-2016-044].
 Figure \ref{fig:geozoo} shows points on the surface of two different types of high-dimensional torii.
@@
 In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes.
-(Adapted from Schloerke, et al., 2016, Figures 10 and 13.)}
+Adapted from \cite{RJ-2016-044}, Figures 10 and 13.}
 \label{fig:geozoo}

### 2021-04-16 paper.Rmd
typo in title

@@
 title:
-  formatted: "A Review of the State-of-the-Art on Tours for Dynamic Visualisation of High-dimensional Data"
+  formatted: "A Review of the State-of-the-Art on Tours for Dynamic Visualization of High-dimensional Data"
 abstract: > 

### 2021-04-20 paper.Rmd
corresponding author and ORCID as requested

@@
     email: stuart.a.lee@monash.edu
-  - name: Dianne Cook
+    orcid: 0000-0003-1179-8436
+  - name: Dianne Cook*
     address: |
@@
     email: dicook@monash.edu
+    orcid: 000-0002-3813-7155
   - name: Natalia da Silva
@@
     email: natalia@iesta.edu.uy
+    orcid: 0000-0002-6031-7451
   - name: Ursula Laa 
@@
     email:  ursula.laa@boku.ac.at
+    orcid: 0000-0002-0249-6439
   - name: Nicholas Spyrison
@@
     email: earo.wang@auckland.ac.nz
+    orcid: 0000-0001-6448-5260
   - name: H. Sherry Zhang
@@
     email: huize.zhang@monash.edu
+    orcid: 0000-0002-7122-1463
 title:
-  formatted: "A Review of the State-of-the-Art on Tours for Dynamic Visualization of High-dimensional Data"
+  formatted: "Advanced Review: The State-of-the-Art on Tours for Dynamic Visualization of High-dimensional Data"
 abstract: > 

### 2021-06-15 paper.Rmd
Nick's Orcid, (Di's email from 20/4)

@@
     email: nicholas.spyrison@monash.edu
+    orcid: 0000-0002-8417-0212
   - name: Earo Wang

### 2021-06-15 paper.Rmd
typo

--- 2021-04-20 paper.Rmd
@@
     address: |
-      | Faculty of Information and Technology, Monash University
+      | Faculty of Information Technology, Monash University
     affiliation: Monash University
     email: nicholas.spyrison@monash.edu
+    orcid: 0000-0002-8417-0212
   - name: Earo Wang

### 2021-08-27 response_to_review.Rmd
started response to review

---
title: "Response to review"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thank you for the detailed and helpful reviews. These are the ways that we have changed the article in response to the comments.

# Co-editor

- I am fond of this reference, but perhaps Debby's contains it:

Salch, J.D. and D.W. Scott, (1997),
``Data Exploration with the Density Grand Tour,''
 Statistical Graphics and Computing Newsletter,
ASA, Alexandria, VA, 8: 7-11, April, 1997.

# Reviewer: 1

- Overall, there is nothing wrong in the article, but lots of language issues, problems with references, and a few aspects of tours (including software and references) that have not been explored in detail. Please see the attached markup for details.

- Before submitting the revision, I would encourage one of the senior native English speaking authors to carefully do a final language review of the article. I have marked many language problems, but there may be more.

# Reviewer 2

## Summary

- This work summarizes the state-of-the-art tour methods for visualizing high-dimensional data. The tour works by linearly projecting the set of data points down to low dimensions through a sequence of smoothly changing orthonormal projections. The work summarizes the development, components for the display, interaction modes, software and latest applications of the tour method. Overall, the work is well-written and well-covered.  I would recommend a minor revision to authors sometime to revise the content and text (see recommended revisions).

## Strengths

- The literature is well covered and described, and up to date. There are abundant number of examples with figures which covers a large variety of use cases of the tour method (e.g. kinds of display, structure of measurements in physics, geometric shapes, a tour looking at itself, ...)

## Recommended Revisions

- I highly appreciate the specification of notations in Section 2.1, but few sections use the notation. 
Instead, different tour methods are only described in text. 
I anticipate that the readers would benefit from a highlight of key differences between different tour variants and interactions using the unified notations.
For example, (as far as I understand) guided tour finds a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$; local tour rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$; manual tour manipulates the contribution of any variable by editing the corresponding row $A[i]$ in the projection frame $A$, $A[i] \leftarrow A[i] + \Delta[i]$ and re-orthonormalizing the columns.

- Section 2.2: "A guided tour ... The little tour ... and a local tour ..." --- cite?

- Section 2.3: "contains all the rotation to be between planes, not a particular basis in any plane". It maybe worth mentioning the difference between "interpolating frames" (interpolating projection frame A and B directly, which involves particular bases) vs. "interpolating planes" (between span(A) and span(B) ), and briefly discuss how the motion can be decomposed to motion of planes and within plane spin.

- Section 2.3: "The space of all tour *paths* is a high-dimensional torus..." --- All paths, all k-planes or all orthogonal k-frames? 
From the context it's not clear to me whether the space of all tour paths is a high-D torus. 
Consider giving some contexts/explanations to this statement?

- Section 3: Li et al. 2020 seems to propose a different way to control the projection, via brushing and dragging data points as opposed to moving individual variables in manual tour (what they call ``the data point mode'' direct manipulation). 
Perhaps it is worth mentioning it as a way of manual control in Section 3.

### 2021-08-27 paper.Rmd
changes in response to review

@@
 Different methods for choosing target planes provide different tour paths.
-The grand tour is generated using randomly selected target.
-A guided tour is generated by choosing particularly structured projection from a projection pursuit optimization.
-The little tour uses all variable bases as targets, and a local tour rocks back and forth from a particular plane to randomly chosen targets in a small neighborhood.
+The grand tour [@Asimov1985-xr] is generated using randomly selected target.
+A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$.
+A local tour (available originally in [@Buja1986-la] rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$.
+A little tour (available originally in [@Buja1986-la] navigates through all of the coordinate planes, for example, for 2D projections $\{A_{ik} = 1, A_{lj} = 1, \forall i, j, k, l = 1, ..., p; i\neq j\}$.
+
+A manual tour [@cook_manual_1997] is the most different, because it effectively allows rotating from a projection where variable, $j$, is fully extended in one direction to another where it contributes nothing, i.e.
+$A[j] \in [0,1]; j=1, ..., p$.
+The contribution of all other variables is restricted by the initial projection used, and by the contribution of $A[j]$.
+The initial projection is typically the optimal projection found by the guided tour, and the manual tour is thus used to determine whether which variables are most important to the interesting structure seen.
 
@@
 
-There are other tour types which don't fit this style of tour.
-The manual tour [@cook_manual_1997] allows the user to change the projection coefficients manually, to rotate a variable into and out of a project, and is discussed later.
 The recently developed slice tour [@Laa2020-js] can be applied to any of the above tour types.

### 2021-08-29 response_to_review.Rmd
changes in response to review
working on reviewer 2 suggestions

@@
 
+## Marked up comments
+
+- check your citation settings; I would have expected that clicking on a citation brings you to that reference in the reference list, but this does not seem to be the case here
+
+    - **This might be a problem with the Wiley template**
+
+- "means" sounds strange; perhaps "refers to the number of dimensions in ..." ?
+
+    - **Changed**
+
+- first introduce "one-dimensional (1D) through to five-dimensional (5D)"; then you can use 1D, ..., 5D
+
+    - **Changed**
+
+- perhaps "As the dimension increases, ..."
+
+    - **Changed**
+
+- change to  "a 2D cube consists of ..." and "a 3D cube consists of ..."
+
+    - **Changed**
+
+- fig 1: explain how the cube doubles: by connecting the vertexes of the two cubes.
+can you use 3 colors: two colors for the two original cubes and a third color for the connecting edges of the vertexes; perhaps draw those connectors as dashed lines
+
+- p2 l49: perhaps "seen that as the dimension increases by one, ..."
+
+- p3 l14 just "e.g.," or "see for example"
+
+- p3 l14 mention the underlying software as well (and then cite these references)
+
+- p3 l16 what are geodesic interpolations? explain in 1 or2 sentences and add a reference
+
+- p3 l17 "a multivariate ..." or "distributions"
+
 # Reviewer 2
@@
 
+    - **Thanks for the positive comments**
+
 ## Strengths
@@
 
+    - **Thanks for the positive comments**
+
 ## Recommended Revisions
 
-- I highly appreciate the specification of notations in Section 2.1, but few sections use the notation. 
-Instead, different tour methods are only described in text. 
+- I highly appreciate the specification of notations in Section 2.1, but few sections use the notation.  Instead, different tour methods are only described in text. 
 I anticipate that the readers would benefit from a highlight of key differences between different tour variants and interactions using the unified notations.
@@
 
+    - **Done**
+
 - Section 2.2: "A guided tour ... The little tour ... and a local tour ..." --- cite?
+
+    - **Done, cited as much as is possible. There isn't a good citation for little tour or local tour, so we have referred to the original Dataviewer software.**
 

### 2021-08-29 paper.Rmd
working on reviewer 2 suggestions

@@
 
-The term "high-dimensional" here means Euclidean space.
+The term "high-dimensional" here refers to the  dimensionality of the Euclidean space.
 Figure \ref{fig:cubes} shows a way to imagine this.
-It shows a sequence of cube wireframes, ranging from 1D through to 5D, where beyond 2D is a linear projection of the cube.
-As dimension increase, a new orthogonal axis is added.
-For cubes, this is achieved by doubling the cube: a 2D is two 1D cubes, a 3D is two 2D cubes, and so forth.
+It shows a sequence of cube wireframes, ranging from one-dimensional (1D) through to five-dimensional (5D), where beyond 2D is a linear projection of the cube.
+As the dimension increases, a new orthogonal axis is added.
+For cubes, this is achieved by doubling the cube: a 2D cube consists of two 1D cubes, a 3D cube consists of two 2D cubes, and so forth.
 

### 2021-08-29 R/cubes.R
cibes code in R directory

@@ -1,0 +1,132 @@
+# wire frame cubes
+library(tidyverse)
+library(ggthemes)
+library(geozoo)
+library(tourr)
+library(patchwork)
+set.seed(5)
+c1 <- cube.iterate(p = 1)
+c1$points <- as_tibble(c1$points)
+c1$edges <- as_tibble(c1$edges)
+c2 <- cube.iterate(p = 2)
+c2$points <- as_tibble(c2$points)
+c2$edges <- as_tibble(c2$edges)
+c3 <- cube.iterate(p = 3)
+proj <- basis_random(3,2) 
+c3$points <- c3$points %*% proj
+colnames(c3$points) <- c("Var1", "Var2")
+c3$points <- as_tibble(c3$points)
+c3$edges <- as_tibble(c3$edges)
+c4 <- cube.iterate(p = 4)
+proj <- basis_random(4,2) 
+c4$points <- c4$points %*% proj
+colnames(c4$points) <- c("Var1", "Var2")
+c4$points <- as_tibble(c4$points)
+c4$edges <- as_tibble(c4$edges) 
+c4$edges.sub <- tibble(from = c(1,1,1,2,2,3,3,4,5,5,6,7), 
+                       to = c(2,3,5,4,6,4,7,8,6,7,8,8))
+c5 <- cube.iterate(p = 5)
+proj <- basis_random(5, 2) 
+c5$points <- c5$points %*% proj
+colnames(c5$points) <- c("Var1", "Var2")
+c5$points <- as_tibble(c5$points)
+c5$edges <- as_tibble(c5$edges) 
+c5$edges.sub <- tibble(from = c(1,1,1,1,2,2,2,3,3,3,4,4,5,5,5,6,6,7,7,8,9,9,9,10,10,11,11,12,13,13,14,15), 
+                       to = c(2,3,5,9,4,6,10,4,7,11,8,12,6,7,13,8,14,8,15,16,10,11,13,12,14,12,15,16,14,15,16,16))
+
+# plot
+# 1D
+p1 <- ggplot() +
+  geom_point(data=c1$points, aes(x=Var1, y=1)) +
+  geom_segment(data=c1$edges, 
+               aes(x=c1$points$Var1[c1$edges$from], 
+                   xend=c1$points$Var1[c1$edges$to],
+                   y=1, yend=1)) + 
+  geom_point(data=c1$points[1,], aes(x=Var1, y=1), colour = "#E7298A") +
[…]

### 2021-08-30 response_to_review.Rmd
detailed all of reviewer 1 pdf annotations - phew

@@
 
+    - **Done**
+
 - p2 l49: perhaps "seen that as the dimension increases by one, ..."
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p3 l14 mention the underlying software as well (and then cite these references)
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p3 l17 "a multivariate ..." or "distributions"
+
+    - **Done**
+
+- p3 l28 this sounds like a limitation of PCPs on your side, but Hurley & Oldford found a possible solution, e.g., https://link.springer.com/article/10.1007/s00180-011-0229-5
+
+- p3 l28 you mention Andrews curves later on when you speak about software; mention them here as well and clarify that they are some kind of tour, but not a grand tour; Wegman & Solka (2002) describe Andrews plot as a pseudo-grand tour that is not space filling - see
+On Some Mathematics for Visualizing High Dimensional Data
+Author(s): Edward J. Wegman and Jeffrey L. Solka
+Source: Sankhyā: The Indian Journal of Statistics, Series A (1961-2002) , Jun., 2002, Vol.
+64, No. 2, Selected Articles from San Antonio Conference in Honour of C. R. Rao (Jun.,
+2002), pp. 429-452
+Published by: Indian Statistical Institute
+Stable URL: https://www.jstor.org/stable/25051404
+
+- p3 l28 Wegman & Solka (2002) also provide mathematical details and further references; while it is not necessary for this article to repeat all of that information, it should at least summarize those mathematical references in one paragraph.
+
+- p3 l32 "analyst" (singular)
+
+- p3 l43 be specific which reference is for which of the 2 techniques
+
+- p3 l51 add "have" ?
+
+- p4 l27 do not just show some graphs; provide at least some meaningful background information: 2 species of crabs where 0 represents xxx and 1 represents yyy; 2 sexes (which is M & F); the data in the left plot seem to be jittered , etc.; what are dimensions in the right plot - cm?; also cite the original article where these data have first been reported and analyzed
+
+- p4 l41 add a "," here
+
+- p5 fig 3 what does flat and hierarchical mean here?
+
+- p5 fig3 be consistent and write as "10D"
+
+- p5 fig3 perhaps "can be seen" or "shown"
+
+- p5 fig3 perhaps "a sample" or "a selection" (as you picked specific projections)
+
+- p5 fig3 distance does not seem to fit here - perhaps "distant"
+
+- p5 l38 add "a"
+
+- p5 l40 perhaps "all columns are..."
+
+- p5 2.2 header these are rarely used; briefly explain what a Huber plot is and how it can be interpreted
+
+- p5 l43 as for the two previous data examples, briefly explain the data set, source, variables, and other information; what are the dimensions of this data set?
+
+- p5 l48 "targets" (plural)
+
+- p5 l49 "projections"
+
+- p5 l50 cite 1 or 2 references for projection pursuit, including the WIREs article - see
+https://wires.onlinelibrary.wiley.com/doi/abs/10.1002/wics.23
+
+- p5 l53 define PCA and include references; one should be the WIREs article - see
+https://onlinelibrary.wiley.com/doi/10.1002/wics.101
+
+- fig 4 what does "best.proj.data" represent - and what does the scale for this variable mean?
+
+- p7 l35 perhaps "these two tour types." ?
+
+- p7 l37 be specific: "in Section xxx"
+
+- p7 l39 same: "in Section yyy"
+
+- p7 l42 perhaps "it maintains" or "to maintain"; same for (2) below
+
+- p7 l44 "rotations"
+
+- p7 l45 "are" not "a"
+
+- p7 l46 "projections{
+
+- p7 l47 "A more detailed ..."
+
+- fig 6 caption "paths", "The guided...", "the grand..."
+
+- p8 l30 Cook worked with 3D tours in VRGobi; ideally cite one of those original references where the 3D tour is introduced or cite this chapter (in particular Section 10.5.2) from a book chapter by Symanzik as a secondary reference:
+http://sfb649.wiwi.hu-berlin.de/fedc_homepage/xplore/ebooks/html/csa/node83.html
+This is from 
+Handbook of Computational Statistics by	 
+J.E. Gentle, W. Härdle, Y. Mori, Editors, ISBN-10: 3540404643
+
+- fig 7 caption remind the reader that small points are identical start points while large points are different end points AND emphasize that each of these 6 views shows exactly the same tour, just under a different projection in 2D space
+
+- p9 l36 change order; perhaps "Transformations of data (such as ..) are often useful prior to touring.
+
+- p9 l45 as before, briefly summarize this data set (dimensions, number of points, and the text that magically occurs with the proper methods are being applied)
+
+- fig 9 caption how many dimensions does this have?; add ","; "6D"
+
+- fig 10 caption what do R and gamma tune? can you verbally explain their purpose?
+
+- section 3 header be consistent with capitalizations of headings, e.g., lower case "i" here but upper case in the subheading below (check and adjust elsewhere)
+
+- p12 l11 add "to"
+
+- p12 l12 which software is this? and which options for tour interactions are commonly available in other software with tour functionality such as XGobi, RGobi, tourr ?
+
+- p12 l25 "or it can be used"
+
+- p12 l30 add "set"
+
+- fig 12 caption perhaps "variable "bl" (shown as an orange line) is being..."; as before, describe the data set, source, dimensions, variables (what is bl), etc.?; add ","
+
+- fig 13 what is hidden (and why)?; what is the purpose of "connect dots"? how are they connected (this does not look like the convex hull in the projection plane); which software has been used here?
+
+- p15 l11 add ","
+
+- p15 l11 add "that"
+
+- p15 l21 indicate the names of the software, i.e., XplorN and (X)Lisp-Stat; in particular, XplorN deserves some further description as it was one of the first software implementations that provided a PCP grand tour; here is a suitable technical report (follow the cite references for more):
+https://apps.dtic.mil/sti/pdfs/ADA313999.pdf
+
+- p15 l22 this paragraph doesn't seem to belong here as it doesn't seem to relate to software; I suggested earlier on that Andrews curves should be introduced earlier in this review; this paragraph can be used there
+
+- p15 l26 explain in one sentence what the image grand tour is and cite one original article for it (likely something from Wegman et al); by the way, here, you cite the WIREs review article, but an original article is missing; for several of my previous comments, you cited the original article, but not the WIREs review article; ideally, both (original and review article where it exists) should be cited in a review like this
+
+- p15 l28 also cite rggobi as it seemed to support some tour as well - see https://vita.had.co.nz/papers/rggobi-rnews.pdf
+
+- p15 l33 write as "three"
+
+- p15 l36 sentence structure is too complex; break at least into 2 sentences and avoid the double parentheses for (software (author))
+
+- p15 l44 remove "a"
+
+- p15 l46 perhaps "structures such as .. separations"
+
+- p15 l48 perhaps "from further extensions" ?
+
+- fig 14 make sure to indicate which software has been used for each tour figure; as before, provide more details about the underlying data; Be consistent how to introduce multiple plots in the same figures; you use (left), (Left), Left: etc.
+
+- p16 l31 3D (for consistency)
+
+- p16 l32 add ","
+
+- p16 l32 period here; then start a new sentence: "By touring..."
+
+- p16 l35 you mix present tense and past tense when you summarize the content of references; here, you use past tense, elsewhere, you use present tense; at least, be consistent; personally, I prefer past tense as authors did something in the past (and not in the present)
+
+- p16 l37 perhaps "running the" ?
+
+- p16 l38 add "the"
+
+- p16 l47 do you need the O notation here? what about dozens and hundreds instead?
+
+- p16 l50 "we" or "they" (several times)? [and note that this is past tense again]
+
+- fig 15 as always, much more details about the underlying data needed
+
+- p17 l36 add ","
+
+- p17 l41 again "we"; perhaps "the authors"
+
+- p17 l41 mention Figure 17 here in the text
+
+- p17 add "5.5 Historical applications" and briefly review other examples from the literature here; e.g., Koschat and Swayne seem to have used the tour for telecommunication data (see below)
+
+- fig 16 as always. more details about the data needed
+
+- fig 16 what is ATLAS7new facet ? explain in main text
+
+- fig 18 perhaps "subset... illustrates ... shape"
+
+- p19 l10 remove "be"
+
+- p19 l12 add "R" (as you mentioned Python and other software before)
+
+- p19 l12 "it" ? - perhaps "supporting software" ?
+
+- p19 l13 add ","
+
+- p19 add a paragraph about success of the various tour methods; Cook et al (1995) stated this (p.163): 
+"but it is important to statisticians for the reason that the inherent structure has never been completely described.
+The combination of the grand tour and projection pursuit contributes significantly to revealing the nature of the variable relationships in seven dimensions.
+Recently, Koschat and Swayne (1992, in press) have used the projection pursuit guided tour in XGobi to explore telecommunications data, and indeed found previously undetected structure.
+"
+
+- p19 l18 add "to perform"
+
+- p19 l19 add ","
+
+- p19 l19 be specific what is missing in current software, i.e., continue with "(such as ...)."
+
+- p19 l23 perhaps "spaces" (there could be more than just one?)
+
+- References
+    - video library url needed 
+    - buja et al in wegman & solka add city
+    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal
+    - carr add page numbers (or is this the previous ref)
+    - chambers et al list the editors
+    - coleman add when accessed
+    - cook et al (1995) adjust style (compare with previous reference)
+    - cook et al (2018) remove the ArXiv information for an article that has been published
+    - cook, swayne (2007) reference is wrong add add city of publication (New York, Berlin, Heidelberg?)
+    - diaconis as above - either full journal names or abbreviations (adjust other references in a similar way, according to WIREs specifications)
+    - fisher remove the first URL and just keep the one with the DOI
+    - friedman is there a DOI?
+    - horst capitalize both "A" as these seem to be geographic names; just "R package" once
+    - hotelling should be "417-441"
+    - huber is there a URL where this can be obtained?
+    - laa et al (2020) section pursuit incomplete
+    - laa et al (2021) burning sage incomplete
+    - laa et al (2020) slice tour update reference (it is published now) and remove the arXiv information
+    - lee at al DOI ?
+    - li incomplete; use lower case "c" and "t" here
+    - li where does the 5(3) come from? - check the web page again
+    - martinez remove the first URL
+    - mcinnes capitalize "UMAP"
+    - moustafa incomplete; Remove the first URL
+    - moustafa and hadi remove the first URL
+    - nguyen capitalize P (in Python), R (in the next reference), etc.
+    - pedersen just once; check gganimate capitalisation
+    - sievert city?; should be "Chapman and Hall/CRC" (several times)
+    - spyrison change to lower case (except R); 243-257 (again!)
+    - swayne DOI
+    - swayne et al DOI; XGobi (as a software name); issue ?
+    - tierney DOI
+    - tukey editors?
+    - tukey pages, city, publisher ?; city?
+    - van der maaten they list this as 9(86) on their web page
+    - wegman list editors only once; page numbers; is there a volume ? - these are Interface (conference proceedings)
+    - wilkinson capital "S"
+    - xie lower case
+    - you "R"
+    - zhang extend reference as for other R packages; incomplete
+
+- I am curious: Why do you not cite https://doi.org/10.1051/epjconf/202024506018
+
+- as you cite R packages, reference R as well
 

### 2021-08-30 paper.Rmd
detailed all of reviewer 1 pdf annotations - phew

@@
 \caption{Illustrating what is meant by "high-dimensional" in this paper, and a "linear projection".
-From a sequence of increasing dimension cubes, from 1D to 5D, as wireframes, it can be seen that the as dimension increase by 1, the cube doubles.
+From a sequence of increasing dimension cubes, from 1D to 5D, as wireframes, it can be seen that as the dimension increase by one, the cube doubles.
 }
@@
 A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
-It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (see for e.g.
-@Swayne2003 or @Tierney1991).
+It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (e.g.
+GGobi @Swayne2003 or XLispStat @Tierney1991).
 To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
-It allows the viewer to extrapolate from the low-dimensional to  shapes corresponding to multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
+A geodesic interpolation produces movement along the shortest path between two planes.
+It allows the viewer to extrapolate from the low-dimensional shapes to a corresponding multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.
 

### 2021-08-30 R/cubes.R

@@ -4,6 +4,8 @@ library(ggthemes)
 library(geozoo)
 library(tourr)
 library(patchwork)
+library(ggrepel)
+
 set.seed(5)
 c1 <- cube.iterate(p = 1)
 c1$points <- as_tibble(c1$points)
@@ -41,7 +43,8 @@ p1 <- ggplot() +
   geom_segment(data=c1$edges, 
                aes(x=c1$points$Var1[c1$edges$from], 
                    xend=c1$points$Var1[c1$edges$to],
-                   y=1, yend=1)) + 
+                   y=1, yend=1), 
+               linetype=3, colour = "#607848") + 
   geom_point(data=c1$points[1,], aes(x=Var1, y=1), colour = "#E7298A") +
   ggtitle("1D") +
   theme_void() +
@@ -52,17 +55,25 @@ p1 <- ggplot() +
 # 2D
 p2 <- ggplot() +
   geom_point(data=c2$points, aes(x=Var1, y=Var2)) +
-  geom_segment(data=c2$edges, 
-               aes(x=c2$points$Var1[c2$edges$from], 
-                   xend=c2$points$Var1[c2$edges$to],
-                   y=c2$points$Var2[c2$edges$from], 
-                   yend=c2$points$Var2[c2$edges$to])) + 
-  geom_point(data=c2$points[1:2,], aes(x=Var1, y=Var2), colour = "#E7298A") +
+  geom_segment(data=c2$edges[c(1,4),], 
+               aes(x=c2$points$Var1[from], 
+                   xend=c2$points$Var1[to],
+                   y=c2$points$Var2[from], 
+                   yend=c2$points$Var2[to])) + 
+  geom_segment(data=c2$edges[c(2,3),], 
+               aes(x=c2$points$Var1[from], 
+                   xend=c2$points$Var1[to],
+                   y=c2$points$Var2[from], 
+                   yend=c2$points$Var2[to]), 
+               linetype = 3, colour = "#607848") + # dashed connectors
+  geom_point(data=c2$points[1:2,], aes(x=Var1, y=Var2), 
+             colour = "#E7298A") +
   geom_segment(data=c2$edges[1,], 
                aes(x=c2$points$Var1[from], 
                    xend=c2$points$Var1[to],
                    y=c2$points$Var2[from], 
-                   yend=c2$points$Var2[to]), colour = "#E7298A") + 
[…]

### 2021-08-30 paper.Rmd
added hurley and oldford ref

@@
 Similarly, parallel coordinates plots (PCP) can be used to explore correlation and collinearity [@Inselberg1985-kf; @Wegman1990-dy].
-By placing multiple variables side by side in a PCP higher order structure like clustering or lower dimensional embeddings are revealed, however the ordering of variables along the axis changes what can be learned.
+By placing multiple variables side-by-side in a PCP higher order, structure like clustering or lower dimensional embeddings are revealed.
+However the ordering of variables along the axis changes what can be learned, and @hurley-oldford suggest repeating variables to alleviate this.
 Another display that relies on variable ordering is the heatmap, which is widely used to visualize cluster structure in bioinformatics.

### 2021-08-30 paper.Rmd
addes the andrews, and wegman references

@@
 While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualizing structure in high dimensional data.
+Very similar to a tour was the Andrews plot [@andrews] which uses a fourier series to create a smooth series of 1-D projections.
+But as @wegman-solka point out this is a pseudo-grand tour because it is not space-filling.
 An early technique proposed for assessing pairwise relationships between variables is the scatterplot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
@@
 @Wilkinson2009-pj provides a comprehensive history of this display in the social and natural sciences.
+
+Recently, some review papers have summarized different aspects of tour methods.
+@Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve.
+@Moustafa2010 describes a computational shortcut for a tour.
+@Martinez2013 explains the image grand tour.
 
@@
 
-Recently, some other review papers have summarized different aspects of tour methods.
-@Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve.
-@Moustafa2010 describes a computational shortcut for a tour.
-@Martinez2013 explains the image grand tour.
-
 The most accessible current software is the R package called tourr [@Wickham2011-uz].

### 2021-08-30 paper.Rmd
brushing

@@
 All of the aforementioned techniques can be enhanced through the use of interactivity.
-By combining views with interactive elements like tool tips or highlighting the analysts is able to quickly interrogate interesting features of the data.
+By combining views with interactive elements like tool tips or highlighting, the analyst is able to quickly interrogate interesting features of the data.
 One particularly important interaction technique in the history of statistical graphics is called brushing [@Becker1987-dw].
-When brushing an analyst drags their mouse over the view which results in a region being drawn onto the canvas.
+When brushing an analyst drags their mouse over the view, which results in a region (the brush) being drawn onto the canvas.
 When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used to highlight or filter data on adjacent views (Figure \ref{fig:brushing}).
-This technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).
+This brushing technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).
 

### 2021-08-30 paper.Rmd
have
crabs reference

@@
 The view on the right responds by highlighting the corresponding points.
-Adapted from \cite{Cook2007-be} Figure 2.12.}
+Adapted from \cite{Cook2007-be} Figure 2.12.
+The left plot shows a jittered scatterplot of two species and two sexes of crabs from \cite{CM74}, and the right plot shows two of their five physical measurements.}
 \label{fig:brushing}
@@
 The 10-dimensional data comes from @Rauber2009-kk.
-The t-SNE view (plot A) shows six clearly separated clusters, all spherical with different sizes.
+The t-SNE view (plot A) shows six clearly separated clusters, all spherical, which have different sizes.
 The tour shows that the clusters do not look like this in the full data space.

### 2021-08-30 R/clusters.R
flat/hierarchical removed

@@ -69,11 +69,12 @@ ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
   scale_colour_brewer("", palette="Dark2") +
   ggtitle("A. t-SNE") + 
   theme_solid() + 
-  theme(plot.title = element_text(colour = "black", size = 12),
-        legend.text = element_text(colour="black", size = 12),
+  theme(legend.position = "none",
+        plot.title = element_text(colour = "black", size = 12),
+#        legend.text = element_text(colour="black", size = 12),
         panel.background = element_rect(fill = "white"),
         panel.border = element_rect(fill = NA, colour="black", size=1,
-                                        linetype = 1)) 
+                                       linetype = 1)) 
 
 # tourr
 library(tourr)

### 2021-08-30 paper.Rmd
through page 4
flat/hierarchical removed
done with fig 3

@@
 \begin{figure*}[!t]
-         \includegraphics[width=0.371\textwidth]{figures/cluster_example.pdf}
+         \includegraphics[width=0.255\textwidth]{figures/cluster_example.pdf}
      \vfill
@@
    \end{minipage}%
-\caption{Comparison of structure perception between nonlinear dimension reduction from 10-d, using t-SNE (A) and a tour (B).
-There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied.
-This can be seen in sample of tour projections shown.
-The two green clusters are (almost) spherical in shape, and very distance from the orange clusters.
+\caption{Comparison of structure perception between nonlinear dimension reduction from 10D, using t-SNE (A) and a tour (B).
+There are six clusters, visible in the t-SNE view, but the relative distance between the clusters is extremely varied.
+This can be seen in a selection of tour projections shown.
+The two green clusters are (almost) spherical in shape, and very distant from the orange clusters.
 Three of the orange clusters are very close to each other (just visible in B.1), and all orange clusters are elliptical.
@@
 The implementation of tour paths in statistical software is reviewed in Section \@ref(software).
-Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning and applied statistics.
+Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning, and applied statistics.
 Finally, Section \@ref(discussion) discusses future research directions for tours.

### 2021-08-30 response_to_review.Rmd
addes the andrews, and wegman references
brushing
have
through page 4
flat/hierarchical removed
done with fig 3
Fixing bibliography based on review

@@
 
+    - **added a sentence and the reference**
+
 - p3 l28 you mention Andrews curves later on when you speak about software; mention them here as well and clarify that they are some kind of tour, but not a grand tour; Wegman & Solka (2002) describe Andrews plot as a pseudo-grand tour that is not space filling - see
@@
 
+    - "Added, and explained"
+
 - p3 l28 Wegman & Solka (2002) also provide mathematical details and further references; while it is not necessary for this article to repeat all of that information, it should at least summarize those mathematical references in one paragraph.
+
+    - "Not done. We have reference the paper but the Buja et al is the better reference for mathematics."
 
@@
 
+    - **Done**
+
 - p3 l43 be specific which reference is for which of the 2 techniques
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p4 l27 do not just show some graphs; provide at least some meaningful background information: 2 species of crabs where 0 represents xxx and 1 represents yyy; 2 sexes (which is M & F); the data in the left plot seem to be jittered , etc.; what are dimensions in the right plot - cm?; also cite the original article where these data have first been reported and analyzed
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p5 fig 3 what does flat and hierarchical mean here?
+
+    - **removed legend because these labels are not really sensible or useful**
 
@@
 
+    - **Done**
+
 - p5 fig3 perhaps "can be seen" or "shown"
+
+    - **used "visible"**
 
@@
 
+    - **Done**
+
 - p5 fig3 distance does not seem to fit here - perhaps "distant"
+
+    - **Done**
 
@@
 
+    - **Moved to the intro with other discussion of lietrature**
+
 - p15 l26 explain in one sentence what the image grand tour is and cite one original article for it (likely something from Wegman et al); by the way, here, you cite the WIREs review article, but an original article is missing; for several of my previous comments, you cited the original article, but not the WIREs review article; ideally, both (original and review article where it exists) should be cited in a review like this
@@
 - References
-    - video library url needed 
-    - buja et al in wegman & solka add city
-    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal
-    - carr add page numbers (or is this the previous ref)
-    - chambers et al list the editors
-    - coleman add when accessed
-    - cook et al (1995) adjust style (compare with previous reference)
-    - cook et al (2018) remove the ArXiv information for an article that has been published
-    - cook, swayne (2007) reference is wrong add add city of publication (New York, Berlin, Heidelberg?)
-    - diaconis as above - either full journal names or abbreviations (adjust other references in a similar way, according to WIREs specifications)
-    - fisher remove the first URL and just keep the one with the DOI
-    - friedman is there a DOI?
-    - horst capitalize both "A" as these seem to be geographic names; just "R package" once
-    - hotelling should be "417-441"
-    - huber is there a URL where this can be obtained?
-    - laa et al (2020) section pursuit incomplete
-    - laa et al (2021) burning sage incomplete
-    - laa et al (2020) slice tour update reference (it is published now) and remove the arXiv information
-    - lee at al DOI ?
-    - li incomplete; use lower case "c" and "t" here
-    - li where does the 5(3) come from? - check the web page again
-    - martinez remove the first URL
-    - mcinnes capitalize "UMAP"
-    - moustafa incomplete; Remove the first URL
-    - moustafa and hadi remove the first URL
-    - nguyen capitalize P (in Python), R (in the next reference), etc.
-    - pedersen just once; check gganimate capitalisation
-    - sievert city?; should be "Chapman and Hall/CRC" (several times)
-    - spyrison change to lower case (except R); 243-257 (again!)
-    - swayne DOI
-    - swayne et al DOI; XGobi (as a software name); issue ?
-    - tierney DOI
-    - tukey editors?
-    - tukey pages, city, publisher ?; city?
-    - van der maaten they list this as 9(86) on their web page
-    - wegman list editors only once; page numbers; is there a volume ? - these are Interface (conference proceedings)
-    - wilkinson capital "S"
-    - xie lower case
-    - you "R"
-    - zhang extend reference as for other R packages; incomplete
+    - video library url needed - done
+    - buja et al in wegman & solka add city - done
+    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal - I think this needs to be fixed via bibstyle
+    - carr add page numbers (or is this the previous ref) - I think this is for the previous reference, again it is there in the bib file, to show it the bibstyle should be formatted accordingly for proceeding entries
+    - chambers et al list the editors - I don't think there was any 'scientific' editor on this book, didn't find any relevant person to list here ...
+    - coleman add when accessed - this was taken from animation package so I got the date when the rda file was added to its github repo and added it here...
+    - cook et al (1995) adjust style (compare with previous reference) - done
+    - cook et al (2018) remove the ArXiv information for an article that has been published - done
+    - cook, swayne (2007) reference is wrong add add city of publication (New York, Berlin, Heidelberg?) - done
+    - diaconis as above - either full journal names or abbreviations (adjust other references in a similar way, according to WIREs specifications) - done
+    - fisher remove the first URL and just keep the one with the DOI - done
+    - friedman is there a DOI? - added
+    - horst capitalize both "A" as these seem to be geographic names; just "R package" once - fixed capitalization, but repetition needs to be fixed in bibstyle
+    - hotelling should be "417-441" - fixed (plus full journal name instead of abbreviation)
+    - huber is there a URL where this can be obtained? - could not find it via google at least...
+    - laa et al (2020) section pursuit incomplete - fixed
+    - laa et al (2021) burning sage incomplete - fixed
+    - laa et al (2020) slice tour update reference (it is published now) and remove the arXiv information - fixed
+    - lee at al DOI ? - added
+    - li incomplete; use lower case "c" and "t" here - fixed
+    - li where does the 5(3) come from? - check the web page again - fixed
+    - martinez remove the first URL - fixed
+    - mcinnes capitalize "UMAP" - fixed
+    - moustafa incomplete; Remove the first URL - fixed
+    - moustafa and hadi remove the first URL - fixed
+    - nguyen capitalize P (in Python), R (in the next reference), etc. - fixed
+    - pedersen just once; check gganimate capitalisation - fixed capitalisation, but repeating version number is bibstyle issue
+    - sievert city?; should be "Chapman and Hall/CRC" (several times) - fixed
+    - spyrison change to lower case (except R); 243-257 (again!) - fixed
+    - sutherland et al DOI - added
+    - swayne DOI - I have no idea why it is not printing the DOI for the XGobi reference...
+    - swayne et al DOI; XGobi (as a software name); issue ? -fixed
+    - tukey editors? - added something based on web search, but not sure I got the right names..
+    - tukey pages, city, publisher ?; city? - as above ..
+    - van der maaten they list this as 9(86) on their web page - fixed
+    - wegman list editors only once; page numbers; is there a volume ?
+- these are Interface (conference proceedings) - problem with showing these information for proceedings entry, again bibstyle issue?
+    - wilkinson capital "S" - fixed
+    - xie lower case - fixed
+    - you "R" - fixed
+    - zhang extend reference as for other R packages; incomplete - fixed
 

### 2021-08-31 paper.Rmd
fixing tsne plot and cleaning up intro

@@
 When the variables are all numeric, or quantitative, visualization often relies on some form of dimension reduction.
-This can be done by taking linear projections, for example, principal component analysis [@Hotelling1933-of] or linear discriminant analysis [@Fisher1936].
-It is also common to reduce dimension with nonlinear techniques like multidimensional scaling (MDS) [@Kruskal1964-do] or t-Distributed Stochastic Neighbor Embedding (t-SNE) [@Van_der_Maaten2008-qa].
+This can be done by taking linear projections, for example, principal component analysis (PCA) [@Hotelling1933-of], linear discriminant analysis (LDA) [@Fisher1936] or projection pursuit (PP) [@Friedman1974-ck].
+It is also common to reduce dimension with nonlinear techniques like multidimensional scaling (MDS) [@Kruskal1964-do],  t-Distributed Stochastic Neighbor Embedding (t-SNE) [@Van_der_Maaten2008-qa] or Uniform Manifold Alignment and Projection (UMAP) [@mcinnes2020umap].
 
@@
 
-The focus of this review will be on visualizing high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx].
+The focus of this review will be on visualizing high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx] and its recent derivatives.
 The reason being that it is not feasible to adequately review the very large area of visualizing high-dimensions, and there have been numerous developments in tours recently.
@@
 A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again.
-It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted (e.g.
-GGobi @Swayne2003 or XLispStat @Tierney1991).
+It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted e.g.
+GGobi [@Swayne2003] or XLispStat [@Tierney1991].
 To create the smooth sequence, a geodesic interpolation is computed between consecutive frames.
@@
 
-Nonlinear dimension reduction techniques such as t-SNE and Uniform Manifold Alignment and Projection (UMAP) have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary [@Van_der_Maaten2008-qa;@mcinnes2020umap].
+<!--Tours can be considered to be a dimension reduction technique.
+Visualizing high-dimensions often starts with dimension reduction, for example, principal component analysis (PCA) [@pearson-pca;@hotelling-pca;@joliffe-cadima-pca], which is a linear dimension reduction technique.
+PCA finds the linear combination (without the normality constraint of a projection vector) of maximum variance, and can be considered to be a special case of projection pursuit (PP) [@Friedman1974-ck] which maximizes an arbitrary function.-->
+
+Nonlinear dimension reduction with the techniques t-SNE and UMAP, have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary.
 However, it is only a summary, and it likely involves substantial warping of the original data space.
@@
 \begin{figure*}[!t]
-         \includegraphics[width=0.255\textwidth]{figures/cluster_example.pdf}
-     \vfill
-   \begin{minipage}{\linewidth}     
    \centering
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour1.png}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour2.png}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour3.png}
-         \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour4.png}
-   \end{minipage}%
-\caption{Comparison of structure perception between nonlinear dimension reduction from 10D, using t-SNE (A) and a tour (B).
-There are six clusters, visible in the t-SNE view, but the relative distance between the clusters is extremely varied.
-This can be seen in a selection of tour projections shown.
-The two green clusters are (almost) spherical in shape, and very distant from the orange clusters.
-Three of the orange clusters are very close to each other (just visible in B.1), and all orange clusters are elliptical.
-The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the dimension reduction.}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example.pdf}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/1.jpg}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/5.jpg}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/3.jpg}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom3/1.jpg}
+         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom3/2.jpg}
+     %\vfill
+   %\begin{minipage}{\linewidth}     
+   %\centering
+   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour1.png}
+   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour2.png}
+  %       \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour3.png}
+   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour4.png}
+  % \end{minipage}%
+\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and linear dimension reduction using a tour (B) from 10D simulated data.
+There are six clusters visible in the t-SNE view, three large and three small, roughly equidistant.
+The relative distance between the clusters in the original space is extremely varied, as can be seen from the selection of tour projections shown.
+The two green clusters are (almost) spherical in shape, because they are spherical in every projection viewed.
+They are also far from all four orange clusters.
+Three of the orange clusters are tiny, and very close to each other (just visible in B.1, and clearer when zoomed in B.5).
+All orange clusters are elliptical.
+The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the nonlinear dimension reduction.}
 \label{fig:tsne-tour}
@@
 Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest.
-A projection basis, $A_{p \times d}$, is matrix that characterizes the direction from which the data are projected and needs to satisfy an orthonormality condition, which requires that each column in $A$ has unit length and are perpendicular.
+A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.
+It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other.
 With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$.
-Figure \ref{fig:notation-huber} shows two examples of low-dimensional projections of the palmer penguins data [@penguins] in a Huber plot [@huber1990] and a histogram.
+The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure.
+Two variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
+In the Huber plots (left) a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of illustrates the projection shows two examples of low-dimensional projections of the palmer penguins data [@penguins] in a Huber plot [@huber1990] and a histogram.
+(The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.)
 
@@
 The histogram (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
-Adapted from \cite{Lee2005-gl}, Figure 1.}
+These plots are made using the R package PPtreeViz.}
 \label{fig:notation-huber}
@@
 
-## Finding targets
+## Different target plane methods produce different tour types
 

### 2021-08-31 R/clusters.R

@@ -1,10 +1,11 @@
 # adapted from Holmes and Ngyuen, 2019
-# Simple Gaussian 10-d
+# Simple Gaussian 10D
 library(ggplot2)
 library(tourr)
 library(Rtsne)
 library(dplyr)
 library(readr)
+library(ggthemes)
 
 tidy_tsne <- function(model, data) {
   enframe <- as.data.frame(model[["Y"]])
@@ -67,10 +68,11 @@ tsne_multi_df <- tibble(tsneX = tsne_multi$Y[,1],
 ggplot(tsne_multi_df, aes(x=tsneX, y=tsneY, colour=cluster)) +
   geom_point() + coord_equal() +
   scale_colour_brewer("", palette="Dark2") +
-  ggtitle("A. t-SNE") + 
+  #ggtitle("A. t-SNE") + 
+  annotate("text", x=-17, y=18.5, label="A. t-SNE", size=5) +
   theme_solid() + 
   theme(legend.position = "none",
-        plot.title = element_text(colour = "black", size = 12),
+        #plot.title = element_text(colour = "black", size = 12),
 #        legend.text = element_text(colour="black", size = 12),
         panel.background = element_rect(fill = "white"),
         panel.border = element_rect(fill = NA, colour="black", size=1,
@@ -93,5 +95,17 @@ render_gif(multi[,2:11],
            frames = 200,
            "figures/cluster_example_zoom.gif")
 
+render_gif(multi[,2:11], 
+           grand_tour(),
+           display_xy(col=clrs, axes="off", half_range = 1, center = TRUE), 
+           frames = 500,
+           width = 200, height = 200,
+           "figures/cluster_example_zoom2.gif")
 
+render_gif(multi[,2:11], 
+           grand_tour(),
+           display_xy(col=clrs, axes="off", half_range = 0.3, center = FALSE), 
+           frames = 500,
+           width = 200, height = 200,
+           "figures/cluster_example_zoom3.gif")
 

### 2021-08-31 paper.Rmd
huber plots

@@
 The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure.
-Two variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
-In the Huber plots (left) a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of illustrates the projection shows two examples of low-dimensional projections of the palmer penguins data [@penguins] in a Huber plot [@huber1990] and a histogram.
-(The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.)
+(The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
+In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections.
+(The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D projection which is plotted in the histogram to the right.
+The histogram exhibits in the top row clustering of the three species, and correspond to the 1D projection with a high PP index value.
+The histogram in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have no interesting pattern.
 

### 2021-08-31 paper.Rmd
made histograms density plots

@@
 The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure.
-(The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
+(The figures produced with a modified function from the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
 In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections.
-(The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D projection which is plotted in the histogram to the right.
-The histogram exhibits in the top row clustering of the three species, and correspond to the 1D projection with a high PP index value.
-The histogram in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have no interesting pattern.
+(The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D data projection which is plotted in the density plot to the right.
+The density plot in the top row exhibits clustering of the three species, and corresponds to a 1D projection with a high PP index value.
+The density plot in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have a less interesting pattern.
 
@@
 The solid line represents how well the projection onto different directions, separates the three species, and the dashed circle is a reference guide.
-The histogram (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
+The density plots (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot.
 These plots are made using the R package PPtreeViz.}
@@
 Different methods for choosing target planes provide different tour paths.
-The grand tour [@Asimov1985-xr] is generated using randomly selected target.
+The grand tour [@Asimov1985-xr] is generated using randomly selected targets.
 A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$.
@@
 Figure \ref{fig:notation-target-gen} shows a representation of 1D tour paths of 5D data, drawn on a PCA space.
+Here PCA provides a convenient 2D representation of the 5D.
 A grand tour path is on the left and a guided tour path is on the right.
@@
 The guided tour path should be short, as the projection pursuit optimization zeroes on the optimally interesting projection.
+(An interesting aside is that although PCA is used here for creating a summary plot, it can also be considered to be a PP method, where the search finds projections where the data is most varied.
+There are many different PP indexes that have been proposed over the last few decades.
+@abdi-williams and @joliffe-cadima-pca provide recent overviews of PCA and @jee2009 provides an overview of PP.)
 

### 2021-08-31 R/tour-notation.R

@@ -175,7 +175,10 @@ Huberplot2 <- function (origdata2D, origclass, PPmethod = "LDA", weight = TRUE,
   group <- origclass
   hist.data <- data.frame(best.proj.data, group)
   P2 <- ggplot(data = hist.data, aes(x = best.proj.data, group = group)) + 
-    geom_histogram(aes(fill = group), position = "stack")
+    #geom_histogram(aes(fill = group), position = "stack") +
+    geom_density(aes(fill = group, colour = group), alpha=0.7) +
+    xlab("Projection") +
+    theme(axis.text.x = element_blank())
   gridExtra::grid.arrange(P1, P2, nrow = 1)
 }
 

### 2021-08-31 response_to_review.Rmd
fixing tsne plot and cleaning up intro
made histograms density plots
done to page 7

@@
  Statistical Graphics and Computing Newsletter,
-ASA, Alexandria, VA, 8: 7-11, April, 1997.
+ASA, Alexandria, VA, 8: 7-11, April, 1997. https://community.amstat.org/jointscsg-section/newsletter
+
+    - **Added**
 
@@
 Please see the attached markup for details.
+
+    - **We are extremely grateful for the detailed read and fixes!**
 
@@
 
+    - **Done**
+
 - p5 l40 perhaps "all columns are..."
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p5 l43 as for the two previous data examples, briefly explain the data set, source, variables, and other information; what are the dimensions of this data set?
+
+    - **Done**
 
@@
 
+    - **Done**
+
 - p5 l49 "projections"
+
+    - **Sentence changed**
 
@@
 
+    - **Introduced in section 1 where PCA and LDA is first mentioned**
+
 - p5 l53 define PCA and include references; one should be the WIREs article - see
@@
 
+    - **added more details on both PCA and PP here**
+
 - fig 4 what does "best.proj.data" represent - and what does the scale for this variable mean?
+
+    - **changed**
 
@@
 
+    - **Paragraph is now changed**
+
 - p7 l37 be specific: "in Section xxx"
+
+    - **done**
 
@@
 
+    - **Paragraph changed and it doesn't fit to forward reference here**
+
 - p7 l42 perhaps "it maintains" or "to maintain"; same for (2) below
+
+    - **done**
 
@@
 
+    - **singular is correct here**
+
 - p7 l45 "are" not "a"
+
+    - **done**
 
@@
 
+    - **done**
+
 - p7 l47 "A more detailed ..."
+
+    - **done**
 
@@
 
+    - **This was a conference presentation and summarised work in other papers. It looks like a paper, which is the physics approach, but it cannot be considered a paper in its own right.**
+
 - as you cite R packages, reference R as well

### 2021-08-31 paper.Rmd
done to page 7

@@
 But as @wegman-solka point out this is a pseudo-grand tour because it is not space-filling.
+@Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve.
 An early technique proposed for assessing pairwise relationships between variables is the scatterplot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw].
@@
 @Wilkinson2009-pj provides a comprehensive history of this display in the social and natural sciences.
-
-Recently, some review papers have summarized different aspects of tour methods.
-@Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve.
-@Moustafa2010 describes a computational shortcut for a tour.
-@Martinez2013 explains the image grand tour.
 
@@
 The recently developed slice tour [@Laa2020-js] can be applied to any of the above tour types.
-It displays a slice through the orthogonal space, as opposed to a projection, and is explained in more details later.
+It displays a slice through the orthogonal space, as opposed to a projection, and is explained in Section \@ref(slices).
 
@@
 The smooth progression in the tour path is due to geodesic interpolation between the target planes.
-This takes into consideration two important aspects: (1) maintains the orthonormality of the projection bases, (2) contains all the rotation to be between planes, not a particular basis in any plane.
-The first is clearly important because it ensures that we a looking at low-dimensional projection of the data always.
+This takes into consideration two important aspects: (1) it maintains the orthonormality of the projection bases, (2) contains all the rotation to be between planes, not a particular basis in any plane.
+The first is clearly important because it ensures that we are looking at low-dimensional projections of the data always.
 The latter is harder to explain, but really important from a visual perspective.
 It stops any within-plane spin, and could be considered stabilizing the view.
-More details explanation of this can be found in @Buja2005-cx.
+A more detailed explanation of this can be found in @Buja2005-cx.
 
@@
 
-Various displays are available to show projections in 1D, 2D and higher dimensions.
+Various displays can be used to show projections in 1D, 2D and higher dimensions.
 A 1D projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2D projections are displayed as scatterplots.
@@
 Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables.
+It is possible to define tours for more complex data also.
+For example, @Martinez2013 proposed the image grand tour for multivariate spatial raster data, which displays a linear combination of variables at each spatial pixel.
 
@@
 
-## Slices, sections and projections
+## Slices, sections and projections {#slices}
 
@@
 The videos titled "XGobi: Dynamic Graphics for Data Analysis" (Swayne et al, 1991) and "Grand Tour and Projection Pursuit" (Cook et al, 1993) show the tour tools in the XGobi software system.
-Some work on tours, for example @Wegman92 and @Tierney1991 is not documented in the video collection.
+Some important work on tours, for example @Wegman92 and @Tierney1991, @salch-scott and @Moustafa2010, is not documented in the video collection.
 

### 2021-08-31 paper.Rmd
through page 8

@@
 \includegraphics[width=\textwidth]{figures/notation-target-gen.pdf}
-\caption{A comparison of guided and grand tour path in the PCA reduced 2D space.
+\caption{A comparison of a guided tour path and grand tour path in a PCA-reduced 2D space.
 The start point of tour paths is smaller and lighter than the end point.
@@
 }
-\caption{Still views of a tour showing the space of possible two-dimensional projections as a torus shape in grey, and the trace of two short paths obtained via geodesic interpolation traced in color.}
+\caption{A selection of projections from a tour showing the space of possible 2D projections from 6D.
+Grey points indicate possible 2D projections.
+The traces of two short paths obtained via geodesic interpolation are shown in color, with starting and ending projections being slightly larger.
+The space of all 2D projections is a torus, as noted in \cite{Buja1986-sw}.
+}
 \label{fig:interpolation-2d}
@@
 
-## Display
+## Displaying data projections
 
 Various displays can be used to show projections in 1D, 2D and higher dimensions.
-A 1D projection displays the data in a histogram analogous to a shadow puppet being projected onto a wall, while 2D projections are displayed as scatterplots.
+A 1D projection displays the data in a histogram, while 2D projections are displayed as scatterplots,  analogous to a shadow puppet being projected onto a wall.
+Cook implemented 3D tours in VRGobi [@vrgobi;@Nelson1999XGobiVT].
 Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates.
@@
 }
-\caption{Still views of the slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
+\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
 \label{fig:tour-slice}

### 2021-08-31 paper.Rmd
fig 9, 10 captions

@@
 }
-\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface (left), 4D torus (middle) and a 6d cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
+\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface, 3D (left), 4D torus (middle) and a 6D cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
 \label{fig:tour-slice}
@@
 
-Transformations of data are often useful prior to touring such as scaling, sphering, or a logarithmic transformation for skewed distributions.
+Transformations of data such as scaling, sphering, or a logarithmic transformation for skewed distributions are often useful prior to touring.
 However, we may also want to transform the data after projecting onto lower dimensions, to correct for unwanted effects of the projections.
@@
 This is illustrated in Figure \ref{fig:pollen} showing the application of the sage transformation to the classical pollen dataset [@Coleman1986-so].
-The projection without any rescaling looks similar to what is found when rescaling with default options (left), and we can use either of the two tuning parameters to better resolve the distribution near the center and reveal the hidden structure (middle and right).
+(This data has five variables and was simulated as a challenge for the 1986 American Statistical Association Data Expo.) The projection without any rescaling looks similar to what is found when rescaling with default options (left), and we can use either of the two tuning parameters to better resolve the distribution near the center and reveal the hidden structure (middle and right).
 
@@
 \includegraphics[width=0.3\textwidth]{figures/pollen_sage_gam20-088.png}
-\caption{Snapshots of the same projection from the sage tour on the pollen data, illustrating the different tuning options.
-Left: using the default settings for the transformation results in a similar view as found in projections without the sage transformation.
-Middle: using the tuning parameter $R$ for better resolution near the center reveals the word "EUREKA".
-Right: using the $\gamma$ tuning parameter can also reveal the structure.
+\caption{Snapshots of the same projection of the classical pollen data during a sage tour, where two parameters control a radial transformation of the projected data.
+Three tuning options are used: (left) default resulting in a similar view as seen in a regular tour; (middle, right) changing the radius and scale parameters, respectively provides better resolution for seeing the hidden word "EUREKA".
 Adapted from \cite{Laa2020-uv}, Figure 9.}
@@
 (Left) The prediction region for points classified as orange produces a slice of a cube.
-(Middle) The predicition region for points classified as green produces another slice of the cube.
+(Middle) The prediction region for points classified as green produces another slice of the cube.
 (Right) The complete prediction region shows the points classified as purple are embedded within the green slice.

### 2021-08-31 paper.Rmd
notation explanation of geodesic added

@@
 A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$.
-A local tour (available originally in [@Buja1986-la] rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$.
-A little tour (available originally in [@Buja1986-la] navigates through all of the coordinate planes, for example, for 2D projections $\{A_{ik} = 1, A_{lj} = 1, \forall i, j, k, l = 1, ..., p; i\neq j\}$.
+A local tour (available originally in @Buja1986-la rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$.
+A little tour (available originally in @McDonald1982-ht navigates through all of the coordinate planes,  e.g.
+in 2D this would correspond to all elements of $A$ are 0 except for one element in each column, $(A_{i1} = A_{j2} = 1, A_{k1} = A_{k2} = 0, ~~i, j, k = 1, ..., p, ~~i\neq j \neq k)$.
 
@@
 It stops any within-plane spin, and could be considered stabilizing the view.
+A geodesic interpolation between two planes is obtained by (1) using singular value decomposition of $A_1^T A_2$ to compute the $d$ smallest angles between the two planes, (2) dividing these angles into small intervals, (3) generating intermediate projection planes by rotating the initial basis $A_1$ by these angle increments.
 A more detailed explanation of this can be found in @Buja2005-cx.
@@
 
-## Basic Interactions
+## Basic interactions
 
@@
 
-## Manual Tours {#tours}
+## Manual tours {#tours}
 
@@
 
-## Model Visualization
+## Model visualization
 

### 2021-08-31 paper.Rmd
through page 15

@@
 The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure.
-(The figures produced with a modified function from the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
+(The figures produced with a modified function from the R [@R_Core_Team2020-lr] package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration.
 In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections.
@@
 It is possible to define tours for more complex data also.
-For example, @Martinez2013 proposed the image grand tour for multivariate spatial raster data, which displays a linear combination of variables at each spatial pixel.
+For example, @Wegman98 proposed the image grand tour for multivariate spatial raster data, which displays a linear combination of variables at each spatial pixel.
+(This is also described in the review paper by @Martinez2013.)
 
@@
 }
-\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface, 3D (left), 4D torus (middle) and a 6D cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
+\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface, 3D (left), 4D torus (middle), and a 6D cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
 \label{fig:tour-slice}
@@
 Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualization.
-Often one will want to pause on a particularly interesting projection and return it for use in a downstream analysis or static plot.
-Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown in Figure \ref{fig:tour-controls}.
+Often one will want to pause on a particularly interesting projection and return to use in a downstream analysis or static plot.
+Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown using the liminal [@Lee2020-ut] R package in Figure \ref{fig:tour-controls}.
+These basic interactions were available in many historical interactive graphics software, but are available using  liminal or plotly [@sievert_interactive_2020] using projection sequence provided by the tourr [@Wickham2011-uz] R package.
 
@@
          \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl4_r.pdf}
-\caption{Three projections from a manual tour where variable "bl" is being rotated out of the projection (orange line).
-When this variable is removed the two light green clusters merge, which informs us that "bl" is an important variable for distinguishing between these two groups.}
+\caption{Three projections from a manual tour where variable bill length (bl) (shown in orange) from the palmer penguins data is being rotated out of the projection.
+When this variable is removed, the two light green clusters merge, which informs us that bill length is an important variable for distinguishing between these two species of penguins.}
 \label{fig:manual-tour}
@@
   {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp13.pdf}}}
-\caption{Stepwise identification of clusters in the physics data.
-Each time a cluster is clearely separated we stop the tour and brush the points, followed by another spin of the data, until we can capture the full distribution.
+\caption{Stepwise identification of clusters in the physics data using the XGobi software.
+Each time a cluster is clearly separated we stop the tour and brush the points, followed by another spin of the data, until all visible clusters are colored.
+The shape of this 7D data is very interesting: a 2D triangle with two linear clusters extending from each vertex.
+Drawing lines in a projection helps to outline the triangle shape.
 Adapted from \cite{Cook2007-be}, Figure 5.3.}
@@
 
-Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or used to section lower dimensional views.
+Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or it can be used to section lower dimensional views.
 This aids tasks such as the identification of outlying points or visual cluster analysis when combined with the tour.
@@
 As an example we briefly summarize an analysis from @Cook2007-be that explores clustering of the physics data previously used to in the context of projection pursuit [@Friedman1974-ck; @Cook1995-ae].
-This data is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
+This data set is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex.
 These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Figure \ref{fig:prim7-tour}.
@@
 The American Statistical Association Statistical Graphics Video Library web site [@ASASGVL] hosts videos documenting the history.
-It is fabulous watching the videos titled "Multidimensional Scaling" (Kruskal, 1962), "Real-Time Rotation" (Chang, 1970) and "Prim-9" (Tukey, 1973) show some preliminary methods leading to the development of the grand tour.
+It is fabulous watching the videos titled "Multidimensional Scaling" (Kruskal, 1962), "Real-Time Rotation" (Chang, 1970), and "Prim-9" (Tukey, 1973) that some preliminary methods leading to the development of the grand tour.
 The video titled "Use of the Grand Tour in Remote Sensing" (McDonald and Willis, 1987) is the first to show a tour, and the video titled "Dataviewer: A Program for Looking at Data in Several Dimensions" (Buja and Tukey, 1987) demonstrates the tour as part of a larger data analysis system.
 The videos titled "XGobi: Dynamic Graphics for Data Analysis" (Swayne et al, 1991) and "Grand Tour and Projection Pursuit" (Cook et al, 1993) show the tour tools in the XGobi software system.
-Some important work on tours, for example @Wegman92 and @Tierney1991, @salch-scott and @Moustafa2010, is not documented in the video collection.
+Some important works on tours, for example XplorN [@Wegman92] and XLispStat [@Tierney1991], density grand tours [@salch-scott] and computationally efficient grand tours [@Moustafa2010], are not documented in the video collection.
 
-The most accessible current software is the R package called tourr [@Wickham2011-uz].
-This software evolved from the Dataviewer [@Buja1986-la], XGobi [@Swayne1998], GGobi [@Swayne2003], Orca [@Sutherland2000] and cranvas [@Xie2014] ancestry.
+The most accessible current software is the R [@R_Core_Team2020-lr] package called tourr [@Wickham2011-uz].
+This software evolved from the Dataviewer [@Buja1986-la], XGobi [@Swayne1998], GGobi [@Swayne2003], Orca [@Sutherland2000], rggobi [@rggobi] and cranvas [@Xie2014] ancestry.
 The tourr package has a wide range of display types, for different projection dimensions, and a selection of target generating methods including grand, guided, little, local, section, sage and frozen.
 
-The R package spinifex [@spyrison_spinifex_2020] has 3 primary features: it produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package, renders (manual or other) tours as animations exportable to static .gif (gganimate [@pedersen_gganimate_2020] package) or interactive .html widgets (plotly [@sievert_interactive_2020] package), lastly it offers interactive shiny application that offers an graphical user interface to quickly sample tour features.
+The R package spinifex [@spyrison_spinifex_2020] has three primary features.
+It produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package.
+It renders (manual or other) tours as animations exportable to static .gif (gganimate @pedersen_gganimate_2020 package) or interactive .html widgets (plotly @sievert_interactive_2020 package).
+Lastly it offers an interactive shiny application providing a graphical user interface to quickly explore different variables contribution to structure in the projection.
 
-The liminal R package uses the tour to explore the quality of non-linear dimensionality reduction algorithms [@Lee2020-kg].
-The interface consists of two side by side views consisting of a scatter plot displaying a reduced form of the data, and an interactive tour.
+The liminal R package uses the tour to explore the quality of non-linear dimension reduction algorithms [@Lee2020-kg].
+The interface consists of two side-by-side views consisting of a scatter plot displaying a reduced form of the data, and an interactive tour.
 Controls such as play/stop/restart are implemented allowing a user to pause on interesting projections and return them to their R session for further analysis.
-Linked brushing is implemented on both views - if a users brush on the scatter plot view they can see if and how an algorithm like t-SNE has distorted distances in higher dimensional space, while if they brush on the tour view, the tour is paused and structure like cluster separation can be ascertained.
+Linked brushing is implemented on both views - if  users brush on the scatter plot view they can see if and how an algorithm like t-SNE has distorted distances in higher dimensional space, while if they brush on the tour view, the tour is paused and structures such as cluster separation can be explored.
 
-A Python implementation of the tour written by @Nguyen2020 is also available, and would benefit from extension by other developers.
+A Python implementation of the tour written by @Nguyen2020 is also available, and would benefit from further extensions by other developers.
 
@@
 For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance (Figure \ref{fig:tour-nn}).
-Similarly, the difference between training and test sets can be assessed through side by side linked tours of layer outputs.
+Similarly, the difference between training and test sets can be assessed through side-by-side linked tours of layer outputs.
 @li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP [@mcinnes2020umap] and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.

### 2021-08-31 paper.Rmd
done with NN example

@@
     \includegraphics[width=0.32\textwidth]{figures/svm-radial-purple.png}
-  \caption{A grand tour run on the prediction boundaries of a radial support vector machine.
-(Left) The prediction region for points classified as orange produces a slice of a cube.
-(Middle) The prediction region for points classified as green produces another slice of the cube.
-(Right) The complete prediction region shows the points classified as purple are embedded within the green slice.
+  \caption{A grand tour using GGobi software to explore the prediction boundaries of a radial support vector machine.
+The prediction region for points classified as orange produces a slice of a cube (left).
+The prediction region for points classified as green produces another slice of the cube (middle).
+The complete prediction region shows the points classified as purple are embedded within the green slice (right).
 Adapted from \cite{Wickham2015-do}, Figure 6.}
@@
 In Figure \ref{fig:svm-tour} samples have been taken along the prediction region (i.e.
-the value the model predicts over a grid of observations) of a 3-$d$ support vector machine used to classify three different classes (green, orange purple), by touring over the prediction region we can see how the classifier splits the decision boundaries.
+the value the model predicts over a grid of observations) of a 3D support vector machine used to classify three different classes (green, orange purple).
+By touring over the prediction region, we can see how the classifier splits the decision boundaries.
 
@@
 The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg].
-They proposed a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
+They propose a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
 They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
-For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance (Figure \ref{fig:tour-nn}).
+For example, by running the tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of the model performance (Figure \ref{fig:tour-nn}).
 Similarly, the difference between training and test sets can be assessed through side-by-side linked tours of layer outputs.

### 2021-08-31 paper.Rmd
done to page 16

@@
 
-Physics models often consider $\mathcal{O}(10)$ free parameters and are compared to a much larger number of experimental observables ($\mathcal{O}(100)$).
+Physics models often consider dozens of free parameters and are compared to hundreds of experimental observables.
 The comparison typically relies on numeric computation of the model predictions for all observables, potentially obscuring the nature of their dependence on the model parameters.
@@
 This was demonstrated in @Cook2018-jm.
-For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point.
-For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the "outlyingness" of the data point marked with an asterisk symbol in Figure \ref{fig:physics}.
+For example, they explore grouping of experiments based on how they constrain the parameter space.
+They also identify an interesting multivariate outlier that points to a potential error with that observation.
+Figure \ref{fig:physics} shows these findings using a selection of projections from a tour.
+You can see the orthogonal nature of the three groups (indicated by color), and the outlier is marked by an asterisk (grey).
 
@@
          \includegraphics[width=0.45\textwidth]{figures/jetCluster-027.png}
-\caption{Selected frames from using a grand tour of the physics data.
-Left: the three different types of measurements (shown in different colors) are aligned along different directions in parameter space.
-Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet.
+\caption{Selected frames from using a grand tour of the physics experiments data.
+The three different types of measurements (shown in different colors) are aligned along different directions in parameter space (left).
+An outlying point is marked with an asterisk symbol (grey) in the ATLAS7new facet (right).
 Adapted from \cite{Cook2018-jm}, Figures 6 and 7.}

### 2021-08-31 paper.Rmd
done with reviewer 2

@@
          \includegraphics[width=0.9\textwidth]{figures/scheidigger_nn_tour.png}
-\caption{A grand tour run on the output of a deep neural network model for classification.
+\caption{A grand tour run on the output of a deep neural network model for classification the benchmark clothing image data, fasion-MNIST.
 The axes correspond to the probabilities that the model of an image belonging to a certain class.
@@
 The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg].
-They propose a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
+Using the benchmark data fashion-MNIST [@fashion-MNIST], they describe a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour.
 They also apply the grand tour to explore the changes in single layers of the network for each training epoch.
@@
          \includegraphics[width=0.45\textwidth]{figures/jetCluster-027.png}
-\caption{Selected frames from using a grand tour of the physics experiments data.
+\caption{Selected frames from using a grand tour of the first six PCs of data on particle physics experiments.
+Each point is a physical observable and measures how the fit prediction changes in a 56D parameter space.
 The three different types of measurements (shown in different colors) are aligned along different directions in parameter space (left).
@@
 In single cell RNA sequencing, scientists are interested in identifying novel cell types and understanding the relationships between cells or their developmental trajectory.
-To achieve this they perform cluster analysis on a counts matrix or principal components and embed the results via t-SNE and label points according the cluster label.
+To achieve this, they perform cluster analysis on a counts matrix or principal components and embed the results via t-SNE and label points according the cluster label.
 One of the main advantages of t-SNE is the avoidance of over-plotting so clusters can be clearly identified on a scatter plot, however, this can come at the cost of interpretability as global distances are distorted.
-In @Laa2020-uv, we used radial transformations of the tour projections as an alternative to t-SNE that better preserves global structure while still retaining cluster topology.
+@Laa2020-uv used radial transformations of the tour projections as an alternative to t-SNE that better preserves global structure while still retaining cluster topology (Figure \ref{fig:mouse-cluster}).
 
@@
 \caption{Several views of torii: (two at left) 4D ring torus, and (two at right) 6D flat torus.
-In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes.
+In each plot, the highlighted points illustrate the low-dimensional donut shapes that build the high-dimensional torus.
 Adapted from \cite{RJ-2016-044}, Figures 10 and 13.}
@@
 \end{figure*}
+
+## Historical applications {#history}
+
+Over the years there have been numerous applications of tours for problems solving.
+Here are a few examples.
+@Koschat-Swayne used the tour for telecommunication data to explore clustering of customers.
+@Symanzik2002NewAO used the image grand tour to detect locations of mines from satellite images, and traces of rock art beneath graffiti from photos.
+Census data is explored in @huh-kim.
+The multicollinearity between predictors in a large regression model on political preferences based on demographics is check with a tour by @forbes-cook-hyndman.
+Clustering is a common application for tours, and @newell-cook-hofmann-jannink discusses exploring clusters in population genetics of food crops.
+@macedo-cook-brown apply the grand tour to examine multiple climate variables in relation to their spatial coordinates.
 
@@
 The rich ecosystem of tours is being actively developed and applied to a broad range of problems including machine learning.
-When tours are be combined with some interactivity it can assist with analyses like clustering (the spinifex and liminal packages).
-Modified displays of projected data can be used to expose local structures (the sage algorithm, slicing and section pursuit).
+When tours are combined with some interactivity supporting software can assist with analyses like clustering (the spinifex and liminal R packages).
+Modified displays of projected data can be used to expose local structures (the sage algorithm, slicing, and section pursuit).
 New algorithm data collection inside the tour code is being used to better diagnose optimization procedures for projection pursuit [@zhang2021visual].
+
+It needs to be stated that using tours can enhance our understanding of relationships between many variables, and typically leads to detecting previously unknown structure.
+Illustrations supporting this statement can be found in many of the applications cited in Section \@ref(history).
+A purpose of this review paper is to encourage broader use of tour methods.
 
@@
 One big challenge for software engineering is to seamlessly embed the tour with interactivity.
-Currently, there is no implementation that analysts can use to perform tasks that were easy in software available in the 80s and 90s.
-More contributions in sectioning as an alternative to projection, could be broadly useful.
+Currently, there is no implementation that analysts can use to perform tasks that were easy to perform in software available in the 80s and 90s, particularly linking between dynamic tours and other displays (histograms, scatterplots or maps).
+More contributions in sectioning, as an alternative to projection, could be broadly useful.
 An example is conditional model visualization, as implemented in the condviz package [@JSSv081i05], which assists with visual exploration of multivariate fitted models.
-It could be interesting to define tour paths for different types of nonlinear subspaces, or non-Euclidean space.
+It could be interesting to define tour paths for different types of nonlinear subspaces, or non-Euclidean spaces.
 

### 2021-08-31 response_to_review.Rmd
through page 8
fig 9, 10 captions
notation explanation of geodesic added
through page 15
done with NN example
done to page 16
done with reviewer 2
completely done with revision

@@
 
+    - **done**
+
 - p8 l30 Cook worked with 3D tours in VRGobi; ideally cite one of those original references where the 3D tour is introduced or cite this chapter (in particular Section 10.5.2) from a book chapter by Symanzik as a secondary reference:
@@
 
+    - **done**
+
 - fig 7 caption remind the reader that small points are identical start points while large points are different end points AND emphasize that each of these 6 views shows exactly the same tour, just under a different projection in 2D space
+
+    - **done**
 
@@
 
+    - **done**
+
 - p9 l45 as before, briefly summarize this data set (dimensions, number of points, and the text that magically occurs with the proper methods are being applied)
+
+    - **done**
 
@@
 
+    - **done**
+
 - fig 10 caption what do R and gamma tune? can you verbally explain their purpose?
+
+    - **done**
 
@@
 
+    - **done**
+
 - p12 l11 add "to"
+
+    - **done**
 
@@
 
+    - **This has been done as much as possible, and the reference to rggobi is added**
+
 - p12 l25 "or it can be used"
+
+    - **done**
 
@@
 
+    - **done**
+
 - fig 12 caption perhaps "variable "bl" (shown as an orange line) is being..."; as before, describe the data set, source, dimensions, variables (what is bl), etc.?; add ","
+
+    - **done**
 
@@
 
+    - **done**
+
 - p15 l11 add ","
+
+    - **done**
 
@@
 
+    - **done**
+
 - p15 l21 indicate the names of the software, i.e., XplorN and (X)Lisp-Stat; in particular, XplorN deserves some further description as it was one of the first software implementations that provided a PCP grand tour; here is a suitable technical report (follow the cite references for more):
 https://apps.dtic.mil/sti/pdfs/ADA313999.pdf
+
+    - **done**
 
@@
 
+    - **done**
+
 - p15 l28 also cite rggobi as it seemed to support some tour as well - see https://vita.had.co.nz/papers/rggobi-rnews.pdf
+
+    - **done**
 
@@
 
+    - **done**
+
 - p15 l36 sentence structure is too complex; break at least into 2 sentences and avoid the double parentheses for (software (author))
+
+    - **done**
 
@@
 
+    - **done**
+
 - p15 l46 perhaps "structures such as .. separations"
+
+    - **done**
 
@@
 
+    - **done**
+
 - fig 14 make sure to indicate which software has been used for each tour figure; as before, provide more details about the underlying data; Be consistent how to introduce multiple plots in the same figures; you use (left), (Left), Left: etc.
+
+   - **done**
 
@@
 
+   - **done**
+
 - p16 l32 add ","
+
+   - **done**
 
@@
 
+   - **done**
+
 - p16 l35 you mix present tense and past tense when you summarize the content of references; here, you use past tense, elsewhere, you use present tense; at least, be consistent; personally, I prefer past tense as authors did something in the past (and not in the present)
+
+   - **fixed**
 
@@
 
+   - **done**
+
 - p16 l38 add "the"
+
+   - **done**
 
@@
 
+   - **done**
+
 - p16 l50 "we" or "they" (several times)? [and note that this is past tense again]
+
+   - **done**
 
@@
 
+   - **done**
+
 - p17 l36 add ","
+
+   - **done**
 
@@
 
+   - **done**
+
 - p17 l41 mention Figure 17 here in the text
+
+   - **done**
 
@@
 
+    - **Phew! Tall order. Added a few examples to illustrate the range of applicability**
+
 - fig 16 as always. more details about the data needed
+
+    - **done**
 
@@
 
+    - **this was a typo - fixed**
+
 - fig 18 perhaps "subset... illustrates ... shape"
+
+   - **done**
 
@@
 
+   - **done**
+
 - p19 l12 add "R" (as you mentioned Python and other software before)
+
+   - **done**
 
@@
 
+   - **done**
+
 - p19 l13 add ","
+
+   - **done**
 
@@
 
+   - **done**
+
 - p19 l18 add "to perform"
+
+   - **done**
 
@@
 
+   - **done**
+
 - p19 l19 be specific what is missing in current software, i.e., continue with "(such as ...)."
 
+   - **done**
+
 - p19 l23 perhaps "spaces" (there could be more than just one?)
+
+   - **done**
 
@@
 
+    - **Fixed. Some of the problems are related to the template,. The bibtex files are complete so we hope that the typesetter can suitably make the complete citations.**
+
 - I am curious: Why do you not cite https://doi.org/10.1051/epjconf/202024506018
@@
 - as you cite R packages, reference R as well
+
+    - **done**
 
@@
 
+    - **done**
+
 - Section 2.3: "The space of all tour *paths* is a high-dimensional torus..." --- All paths, all k-planes or all orthogonal k-frames? 
@@
 
+    - **cited the original buja and asimov paper that states this, and you can see it from the torus shapes in Figure 18, too.**
+
 - Section 3: Li et al.
@@
 Perhaps it is worth mentioning it as a way of manual control in Section 3.
+
+    - **done**

### 2021-08-31 paper.Rmd
completely done with revision

@@
 Figure \ref{fig:interpolation-2d} shows tour paths of 2D projections of 6D data.
-The space of all tour paths is a high-dimensional torus, as represented by the gray points.
-The tour paths are shown in green and orange, and each dot indicates a projection in the sequence.
+The space of all tour paths is a high-dimensional torus [@Buja1986-sw], as represented by the gray points.
+(Note that the shape of the grey points matches that of the torus shapes discussed later and shown in Figure \ref{fig:geozoo}.) The tour paths are shown in green and orange, and each dot indicates a projection in the sequence.
 The plots in this figure could be considered to be a tour looking at itself, because each plot is a selected view from a tour of the torus with the paths overlaid.
@@
 
+Note that, in the application of tours to understanding neural networks by @Li2020-kg} in Section \@ref(models) also has a manually controlled tour based on selecting and dragging individual points that they call the "the data point mode".
+
 ## Spin-and-brush {#sec:spin}
@@
 
-## Model visualization
+## Model visualization {#models}
 

### 2021-09-01 paper.Rmd
stuart's address + rggobi reference

@@
       | Department of Econometrics and Business Statistics, Monash University
-    email: stuart.a.lee@monash.edu
+    email: stuart.andrew.lee@gmail.com
     orcid: 0000-0003-1179-8436

### 2021-09-01 response_to_review.Rmd
Small commit so I can pull changes..

--- 2021-08-31 response_to_review.Rmd
@@
     - buja et al in wegman & solka add city - done
-    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal - I think this needs to be fixed via bibstyle
-    - carr add page numbers (or is this the previous ref) - I think this is for the previous reference, again it is there in the bib file, to show it the bibstyle should be formatted accordingly for proceeding entries
+    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal, add page numbers - this probably needs to be fixed via bibstyle (no page numbers showing for proceeding entries)
     - chambers et al list the editors - I don't think there was any 'scientific' editor on this book, didn't find any relevant person to list here ...

### 2021-09-01 response_to_review.Rmd
Cleaned up comments on bibliography in response

@@
     - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal, add page numbers - this probably needs to be fixed via bibstyle (no page numbers showing for proceeding entries)
-    - chambers et al list the editors - I don't think there was any 'scientific' editor on this book, didn't find any relevant person to list here ...
-    - coleman add when accessed - this was taken from animation package so I got the date when the rda file was added to its github repo and added it here...
+    - chambers et al list the editors - it seems there was no relevant editor to list here
+    - coleman add when accessed - date added based on animation package (data was accessed through the package)
     - cook et al (1995) adjust style (compare with previous reference) - done
@@
     - friedman is there a DOI? - added
-    - horst capitalize both "A" as these seem to be geographic names; just "R package" once - fixed capitalization, but repetition needs to be fixed in bibstyle
+    - horst capitalize both "A" as these seem to be geographic names; just "R package" once - fixed capitalization, but repetition needs to be fixed via bibstyle
     - hotelling should be "417-441" - fixed (plus full journal name instead of abbreviation)
-    - huber is there a URL where this can be obtained? - could not find it via google at least...
+    - huber is there a URL where this can be obtained? - unfortunately not
     - laa et al (2020) section pursuit incomplete - fixed
@@
     - sutherland et al DOI - added
-    - swayne DOI - I have no idea why it is not printing the DOI for the XGobi reference...
+    - swayne DOI - information is in the bib file but not showing (maybe bibstyle issue?)
     - swayne et al DOI; XGobi (as a software name); issue ? -fixed
-    - tukey editors? - added something based on web search, but not sure I got the right names..
-    - tukey pages, city, publisher ?; city? - as above ..
+    - tukey editors? - added
+    - tukey pages, city, publisher ?; city? - added
     - van der maaten they list this as 9(86) on their web page - fixed
     - wegman list editors only once; page numbers; is there a volume ?
-- these are Interface (conference proceedings) - problem with showing these information for proceedings entry, again bibstyle issue?
+- these are Interface (conference proceedings) - problem with showing these information for proceedings entry, again this appears to be a bibstyle issue
     - wilkinson capital "S" - fixed

### 2021-09-02 paper.Rmd
Fixed typos

@@
 A local tour (available originally in @Buja1986-la rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$.
-A little tour (available originally in @McDonald1982-ht navigates through all of the coordinate planes,  e.g.
+A little tour (available originally in @McDonald1982-ht) navigates through all of the coordinate planes,  e.g.
 in 2D this would correspond to all elements of $A$ are 0 except for one element in each column, $(A_{i1} = A_{j2} = 1, A_{k1} = A_{k2} = 0, ~~i, j, k = 1, ..., p, ~~i\neq j \neq k)$.
@@
 The contribution of all other variables is restricted by the initial projection used, and by the contribution of $A[j]$.
-The initial projection is typically the optimal projection found by the guided tour, and the manual tour is thus used to determine whether which variables are most important to the interesting structure seen.
+The initial projection is typically the optimal projection found by the guided tour, and the manual tour is thus used to determine which variables are most important to the interesting structure seen.
 
@@
 
-Note that, in the application of tours to understanding neural networks by @Li2020-kg} in Section \@ref(models) also has a manually controlled tour based on selecting and dragging individual points that they call the "the data point mode".
+Note that, in the application of tours to understanding neural networks by @Li2020-kg in Section \@ref(models) also has a manually controlled tour based on selecting and dragging individual points that they call the "the data point mode".
 
@@
 The R package spinifex [@spyrison_spinifex_2020] has three primary features.
-It produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package.
+It produces manual tours (predefined path or interactive manipulation), identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package.
 It renders (manual or other) tours as animations exportable to static .gif (gganimate @pedersen_gganimate_2020 package) or interactive .html widgets (plotly @sievert_interactive_2020 package).
@@
 Census data is explored in @huh-kim.
-The multicollinearity between predictors in a large regression model on political preferences based on demographics is check with a tour by @forbes-cook-hyndman.
+The multicollinearity between predictors in a large regression model on political preferences based on demographics is checked with a tour by @forbes-cook-hyndman.
 Clustering is a common application for tours, and @newell-cook-hofmann-jannink discusses exploring clusters in population genetics of food crops.
